### 2024-04-29 homework1/paper.tex
import latex configuration from a pervious project

\documentclass[runningheads]{paper}

\usepackage{hyperref}   % hyperlinks
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{multirow}
\usepackage{longtable}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

% ------------------------------------------------------------------------------

\begin{document}
%
\title{Class Scheduling Problem}
%
\author{Cărămidă Iustina-Andreea - 332CA}
%
\institute{Facultatea de Automatică și Calculatoare \\
Universitatea Politehnica București \\
\email{iustina.caramida@stud.acs.upb.ro}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
    Problema colorării nodurilor unui graf (\textit{graph coloring - GCP} sau \textit{colorarea grafurilor})
    este una dintre cele mai importante probleme, fapt datorat numărului mare de situații din viața reală
    care se pot rezuma la o problemă de colorare a unui graf. În această temă, voi prezenta câteva comparări 
    între diferiți algoritmi dedicați acestei probleme, precum Greedy, Dsatur și RLF din 
    punct de vedere al timpului de execuție, al memoriei folosite și al performanței în raport cu numărul 
    de noduri și muchii.

\keywords{Brute-Force \and Greedy \and DSatur \and RLF \and Colorarea grafurilor.}
\end{abstract}

% ------------------------------------------------------------------------------

\section{Introducere}
\subsection{Descrierea problemei}

\quad Colorarea grafurilor este o problemă centrală în teoria grafurilor \cite{1}. Ea 
constă în alegerea unui set de culori pentru nodurile unui graf, astfel încât 
niciun nod adiacent să nu primească aceeași culoare.

Problema colorării graficului a început cu încercarea lui 
Francis Guthries de a colora toate țările pe harta Angliei. 
Acesta a presupus inițial că patru culori sunt suficiente pentru a procesa orice 
hartă, astfel încât să nu fie asociate două țări vecine
cu aceeasi culoare. Această sarcină este doar una dintre cele peste 200 de probleme
\cite{2}, legate de aria analizei grafice cromatice,
iar această situație poate fi tradusă prin colorarea fiecărui vârf
dintr-un graf în care marginile sale ar reprezenta vecinătatea
dintre două regiuni \cite{3}.

Colorarea grafurilor este o problemă de tip NP-hard bine studiată cu aplicații 
importante în optimizarea combinatorie și într-un domeniu de cercetare activ, cu 
multe aplicații practice \cite{4} în inginerie, cum ar fi \textit{alocarea registrelor}, 
\textit{atribuirea frecvenței}, \textit{potrivirea șablonului} și \textit{programări}. 
În consecință, colorarea grafurilor a fost subiectul unor cercetări intense \cite{5} \cite{2}.

Un exemplu de o problemă de colorare a grafurilor este planificarea examenelor, unde
fiecare examen este reprezentat de un nod și fiecare legătură dintre nodiri reprezintă
o interzicere de a planifica examenele. Astfel, colorarea grafului reprezintă planificarea
examenelelor, astfel încât două examene care nu pot fi planificate în același timp să fie
colorate diferit. În acest caz, numărul minim de culori necesare pentru a colora graful
este egal cu numărul minim de zile necesare pentru a planifica examenele. \cite{6}

Colorarea grafurilor este asociată cu două tipuri de colorare: colorarea vârfurilor 
și colorarea marginilor. Scopul ambelor tipuri de colorare este de a colora întregul graf fără
conflicte. Prin urmare, vârfurile adiacente sau muchiile adiacente trebuie
să fie în culori diferite. Numărul cel mai mic de culori necesare 
pentru a colora graful se numește \textbf{număr cromatic}.
Pe măsură ce numărul de vârfuri sau muchii dintr-un graf crește,
complexitatea problemei crește și ea. Din această cauză, fiecare
algoritmul nu poate găsi numărul cromatic exact  
și pot fi, de asemenea, diferențe în timpul lor de execuție \cite{7}. Pentru a obține 
o soluție mai bună pentru colorarea grafurilor, mulți algoritmi euristici și meta
euristici au fost inventați \cite{8}.
% 
\subsection{Specificarea algoritmilor aleși}

\textbf{\textit{Cerință:}} Fie un graf neorientat G cu N noduri și M muchii. Problema cere să asociem o culoare 
fiecărui nod, astfel încât oricare două noduri adiacente (conectate printr-o muchie directă) 
să aibă culori diferite. Care este numărul minim de culori necesare pentru a colora 
toate nodurile conform restricției menționate anterior? \cite{9}

\subsubsection{Brute Force:}
Când încercăm să oferim o soluție la această problemă, primul instinct este de a
folosi o abordare ”Brute Force”. Acest lucru ar duce la o soluție, care ar fi 
din punct de vedere al timpului de execuție exponențială, făcând această soluție 
inutilă pentru cazuri mari. A spune că nu putem găsi un algoritm eficient 
deoarece acesta nu există ar fi la fel ca și când am spune că problema nu are 
nicio soluție eficientă \cite{10}.
În prezent, există algoritmi care se ocupă să rezolve problema colorării grafurilor, 
deși obțin un număr cromatic apropiat de ale grafului în schimbul unui timp
rezonabil sau rezultate rapide care sunt suficient de utile \cite{11}.

\subsubsection{Greedy Algorithm:}
Logica algoritmului ia vârfurile grafului unul câte unul, urmând o ordine 
(care poate fi aleatorie) și atribuie prima culoare disponibilă fiecărui vârf \cite{12}. 
Deoarece este un algoritm euristic, soluția oferită de acesta poate să nu fie optimă.
Cu toate acestea, o alegere corectă a ordinii vârfurilor pentru
colorarea lor poate oferi o soluție optimă pentru orice graf. În
practică, algoritmul Greedy produce soluții rapid practicabile, deși aceste soluții 
pot fi “sărace” pe baza numărului de culori pe care algoritmul le cere, în comparație cu
numărul cromatic al grafului.

\subsubsection{DSatur Algorithm:}
Algoritmul DSatur (abreviere din engleză pentru “Degree Saturation”), 
propus de Brelaz (1979), se comportă foarte asemănător cu
algoritmul Greedy, cu excepția că, în acest caz,
ordonarea vârfurilor este generată de algoritmul însuși.
La fel ca în algoritmul Greedy, ordonarea a fost decisă
înainte ca orice vârf să fie colorat, ordinea vârfurilor fiind decisă euristic pe baza
caracteristicilor colorării parțiale a grafului la
momentul în care se selectează fiecare dintre vârfuri \cite{13}. 
În cel mai rău caz, complexitatea sa are aceeași situație ca și în
Algoritmul Greedy, deși în practică poate fi luat în considerare
și faptul că monitorizare saturaţiei vârfurilor necolorate
produce o complexitate puțin mai mare. Este de reținut că
Algoritmul DSatur este \textbf{exact} pentru grafurile bipartite \cite{14}.

\subsubsection{RLF Algorithm:}
Algoritmul RLF (abreviere din engleză pentru “Recursive Largest First”), propus
de Leighton (1979), lucrează prin colorarea unui graf cu o singură culoare
pentru fiecare iterație a algoritmului, în loc de un vârf per
repetare. În fiecare iterație, algoritmul caută seturi de
vârfuri independente din graf, care vor fi asociate
cu aceeași culoare. Acel set independent a fost eliminat
din graf, iar subgraful rămas va continua
în același mod, până când subgraful menționat este gol, caz în care
toate vârfurile vor fi atribuite unei culori, producând astfel o
soluție ce satisface toate cerințele \cite{12}.

\subsection{Evaluarea soluțiilor}
Sursele vor fi testate pe grafuri de dimensiuni diferite, de la 10 la 2 000 de noduri,
cu un număr de muchii de la 10 la 1 000 000. Pentru fiecare graf se va genera o
configurație aleatoare de noduri și muchii, iar apoi se va testa performanța
algoritmilor pe aceste grafuri. Pentru fiecare algoritm, se va calcula timpul de
execuție și numărul de culori folosite pentru a colora graful.

\section{Prezentarea soluțiilor}
\subsection{Algoritmul Brute Force}
Algoritmul Brute Force este un algoritm care se bazează pe forță bruta, adică
parcurge toate posibilitățile de colorare a grafului și alege cea mai bună
soluție. Complexitatea acestui algoritm este de $O(n^m)$, unde $n$ este numărul de noduri și $m$ este
numărul de culori. Pentru grafuri cu un număr mare de noduri, complexitatea
acestui algoritm este foarte mare, de aceea nu este folosit în practică. Pe de altă parte,
algoritmul este exact, adică oferă întotdeauna o soluție optimă. Un pseudocod al
algoritmului este prezentat în Pseudocodul~\ref{alg1}.

\begin{algorithm}
\caption{Brute Force Algorithm}
\label{alg1}
\begin{algorithmic}[1]
\Procedure{BruteForce}{$G$} \Return $List\langle int\rangle $ 
\State $List\langle int\rangle $ $colors = new List\langle int\rangle $ \Comment{List of colors}
\State $int$ $maxColors = 0$ \Comment{Number of colors}
\State $int$ $maxColorsIndex = 0$ \Comment{Index of the best solution}
\algstore{bkbreak}
\end{algorithmic}
\end{algorithm}

\addtocounter{algorithm}{-1}
\begin{algorithm}
\caption{Brute Force Algorithm}
\begin{algorithmic}[1]
\algrestore{bkbreak}
\State $int$ $i = 0$ \Comment{Index of the current solution}
\State $int$ $n = G.GetNumberOfNodes()$ \Comment{Number of nodes}
\State $int$ $m = G.GetNumberOfColors()$ \Comment{Number of colors}
\State $int$ $max = pow(m, n)$ \Comment{Number of possible solutions}
\While{$i < max$}
\State $colors = G.GetColors()$
\State $int$ $currentColors = colors.Count()$
\If{$currentColors > maxColors$}
\State $maxColors = currentColors$
\State $maxColorsIndex = i$
\EndIf
\State $i++$
\EndWhile
\State $colors = G.GetColors(maxColorsIndex)$
\State \Return $colors$
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Algoritmul Greedy}
Algoritmul Greedy este un algoritm care alege mereu cea mai bună soluție la
momentul curent, fără a ține cont de soluțiile viitoare.
Un pseudocod al algoritmului este prezentat în
Pseudocodul~\ref{alg2}.

\begin{algorithm} 
\caption{Greedy Algorithm}
\label{alg2}
\begin{algorithmic}[1]
\Procedure{Greedy}{$G$}
\Require{$S$ = Class Set, $V$ = Colorless vertices in random order}
\Ensure{$S$ = $\emptyset$}
\For{$v \in V$}
\For{$i$ to $S.lenght$}
\If {$NonConflitiveEdges(v \cup S_i)$}
\State $AssignClass(v, S_i)$
\State $Next Vertex$
\EndIf
\EndFor
\If {$NotColored(v)$}
\State $S_{i+1} = NewClass$
\State $AssignClass(v, S_{i+1})$
\State {Next Vertex}
\EndIf
\EndFor
\EndProcedure
\end{algorithmic}
\end{algorithm}

În algoritmul Greedy, așa cum se poate vedea în pseudocod,
se folosește o permutare inițială care este generată aleatoriu și în
care este colorat fiecare vârf, comparând în fiecare caz dacă acesta
poate fi inclus într-o culoare fără a provoca conflicte. Pentru a analiza
algoritmul în cel mai rău caz, vom presupune că fiecare vârf este 
verificat cu fiecare culoare înainte de a fi colorat,
ce ne dă un polinom care este
\begin{equation} P(x) = n(n+1)/2. \end{equation} Polinomul
se obţine din faptul că fiecare vârf, înainte de a fi colorat,
trebuie să se compare cu toate culorile actuale:
\begin{equation} P = 0 + 1 + 2 + . . . + n = n(n + 1)/2 \end{equation}

În cele din urmă, vom avea o complexitate de $O(n^2)$, unde $n$ este numărul
de noduri. Este necesar să afirmăm că algoritmul Greedy, deși calculează
numărul cromatic aproximativ, acest lucru va depinde direct de permutarea
inițială pe care o avem. Cu algoritmul următor vom vedea tehnici care ne
permit să îmbunătățim modul de a alege vârfurile pentru a obține soluții
mai bune și mai apropiate de numărul cromatic.

\subsection{Algoritmul DSatur}
Algoritmul DSatur este un algoritm care alege mereu nodul cu cea mai mare
valoare de saturație, adică nodul care are cel mai mare număr de vecini cu
aceeași culoare. Un pseudocod al algoritmului este prezentat în Pseudocodul~\ref{alg3}.
\begin{algorithm}
\caption{DSatur Algorithm}
\label{alg3}
\begin{algorithmic}[1]
\Procedure{DSatur}{$G$}
\Require{$S$ = Class Set, $V$ = Non-colored vertices}
\Ensure{$S$ = $\emptyset$}
\While {$V \neq \emptyset$}
\State $v = MaxSaturation(V)$
\For {$i$ to $S.lenght$}
\If {$NonConflitiveEdges(v \cup S_i)$}
\State $AssignClass(v, S_i)$
\State {Next Vertex (break)}
\EndIf
\If {$NotColored(v)$}
\State $S_{i+1} = NewClass$
\State $AssignClass(v, S_{i+1})$
\State $Remove(v, V)$
\State {Next Vertex}
\EndIf
\EndFor
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

Se poate observa din pseudocod că acest algoritm este foarte asemănător cu algoritmul
Greedy, însă puterea algoritmului DSatur constă în prioritatea dată nodurilor cu cea mai
mare saturație. Astfel, aceste noduri mai restrânse sunt colorate înainte de restul nodurilor,
care nu au aceste “restricții”. Complexitatea algoritmului este similară cu cea a algoritmului
Greedy, deoarece este foarte asemănător în cod, însă are și calcule suplimentare pentru a
obține saturația nodurilor, însă acestea nu afectează direct complexitatea totală a algoritmului,
care este totuși $O(n^2)$.

Trebuie punctat faptul că, deși algoritmul DSatur are o performanță mai bună,
acesta are și cazuri în care obține o soluție
mai proastă decât algoritmul Greedy. Aceasta se poate observa în \cite{15} unde se arată că o
distribuția neadecvată a vârfurilor este dificil de colorat cu algoritmul DSatur.

\subsection{Algoritmul RLF}
Algoritmul RLF caută să coloreze toate vârfurile disponibile în momentul în care
acestea nu afectează niciun conflict al aceleiași culori. De asemenea, acordă
prioritate vârfurilor care au un grad mai mare, acordând astfel prioritate tipului
nostru de restricție prevăzut pentru această euristică. Algoritmul RLF are
o complexitate în cel mai rău caz mai mare decât cea a algorimului Greedy sau
DSatur, aceasta fiind o complexitate dovedită și de Leighton (1979) - $O(n^3)$.
Cu toate acestea acest algoritm are și o îmbunătățire a calității soluțiilor pe
care le oferă, deşi la un cost de calcul mai mare. Mai jos se poate observa pseudocodul~\ref{alg4} algoritmului RLF.

\begin{algorithm}
\caption{RLF Algorithm}
\label{alg4}
\begin{algorithmic}[1]
\Procedure{RLF}{$G$}
\Require{$S$ = Class Set, $V$ = Competent vertices, $W$ = = Non-Competent vertices}
\Ensure{$S$ = $\emptyset$, $V$ = Vertices of a given graph, $W$ = $\emptyset$}
\While {$V \neq \emptyset$}
\State $S_i = NewClass$
\While {$V \neq \emptyset$}
\State $v = MaxSaturation(V)$
\State $AssignClass(v, S_i)$
\State $Add($ Adjacent Vertices $(v), W)$
\State $Remove($ Adjacent Vertices $(v), W)$
\EndWhile
\State $V = W$
\State $W = \emptyset$
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

\subsection{Complexități}
Mai jos, tabelul~\ref{tabel_complexitate} conține complexitățile algoritmilor
prezentați mai sus.

\begin{table}[h]
\centering
\caption{Complexități algoritmilor}
\label{tabel_complexitate}
\begin{tabular}{|c|c|c|c|c|}
\hline

\textbf{Algoritm} & \textbf{Complexitate} & \textbf{Complexitate} & \textbf{Complexitate} & \textbf{Complexitate} \\
 & \textbf{Brute Force} & \textbf{Greedy} & \textbf{DSatur} & \textbf{RLF} \\
\hline
\hline
\textbf{Cazul bun} & $O(n^m)$ & $O(n)$ & $O(n)$ & $O(n)$ \\
\hline
\textbf{Cazul rău} & $O(n^m)$ & $O(n^2)$ & $O(n^2)$ & $O(n^3)$ \\
\hline
\end{tabular}
\end{table}

\section{Evaluare}
\subsection{Grafuri de test și execuția algoritmilor}
Pentru a evalua performanța algoritmilor de colorare, am folosit grafuri de test
generate cu ajutorul unui generator de teste de pe Github, ce îl puteți accesa aici \cite{16}.
Am generat 34 de teste prezentate în tabelul~\ref{tab1} pentru toți algoritmii
prezentați mai sus.

\begin{longtable}{@{}|c|c|c|c|@{}}
    \hline\hline
    numărul testului & tipul grafului  & numărul de noduri (N)  & numărul de muchii (M) \\
    \hline
    1 & graf gol  & 887  & 0 \\
    2 &  & 1384 & 0 \\ \hline
    3 & graf complet & 10 & 45 \\
    4 &  & 50 & 1225 \\
    5 &  & 100 & 4950 \\
    6 &  & 500 & 124750 \\
    7 &  & 1000 & 499500 \\
    8 &  & 2000 & 1999000 \\ \hline
    9 & graf bipartit & 10 & 21 \\
    10 & complet & 50 & 504 \\
    11 &  & 100 & 1771 \\
    12 &  & 500 & 35275 \\
    13 &  & 1000 & 164151 \\
    14 &  & 2000 & 557775 \\ \hline
    15 & arbore binar & 10 & 9 \\
    16 &  & 50 & 49 \\
    17 &  & 100 & 99 \\
    18 &  & 500 & 499 \\
    19 &  & 1000 & 999 \\
    20 &  & 2000 & 1999 \\ \hline
    21 & graf planar & 10 & 20 \\
    22 &  & 50 & 100 \\
    23 &  & 100 & 200 \\
    24 &  & 500 & 1000 \\
    25 &  & 1000 & 2000\\
    26 &  & 2000 & 4000 \\ \hline
    27 & graf cu număr & 1383 & 10 \\
    28 & de muchii fixat & 886 & 100 \\
    29 &  & 777 & 500 \\
    30 &  & 915 & 1000 \\
    31 &  & 1793 & 10000 \\
    32 &  & 1386 & 100000 \\
    33 &  & 1421 & 500000 \\
    34 &  & 1763 & 1000000 \\
    \hline
    \caption{Grafuri de test}
    \label{tab1}
    \end{longtable}

\subsubsection{Descrierea testelor \\}

\textbf{Grafuri goale:} Acestea sunt grafuri care nu au nicio muchie. Numărul cromatic al acestor grafuri este 1, deoarece toate nodurile se pot colora la fel.

\textbf{Grafuri complete:} Acestea sunt grafuri care au toate nodurile conectate
între ele. Numărul cromatic al acestor grafuri este egal cu numărul de noduri.

\textbf{Grafuri bipartite complete:} Acestea sunt grafuri bipartite (=  graf care
poate fi împărțit în două multimi disjuncte
de noduri astfel încât orice muchie din graf să unească un nod din prima mulțime
cu un nod din a doua mulțime) care au proprietatea
că pentru orice nod x din A şi orice nod y din B există muchia [x, y].
Numărul cromatic al acestor grafuri este egal cu 2.

\textbf{Arbori binari:} Aceștia sunt grafuri care au proprietatea că orice nod
are cel mult 2 vecini. Numărul cromatic al acestor grafuri este egal cu 2.

\textbf{Grafuri planare:} Acestea sunt grafuri care pot fi reprezentate pe o
suprafață plană fără ca muchiile acestora să se intersecteze. Numărul cromatic
al acestor grafuri este cel mult egal cu 4.

\textbf{Grafuri cu număr de muchii fixat:} Acestea sunt grafuri care au un număr
fix de muchii. Numărul cromatic al acestor grafuri nu se poate precalcula.

\subsection{Specificațiile sistemului de calcul}
Codul a fost builduit în C++, iar fiecare dintre algoritmii prezentați mai sus au fost
evaluați folosind un Laptop cu următoarele specificații:
\begin{itemize}
\item Procesor: Intel(R) Core(TM) i7-10750H CPU @2.60GHz   2.59 GHz
\item Memorie RAM: 16 GB
\item Sistem de operare: Windows 10 Pro 64-bit
\item Versiunea compilatorului: GCC 9.4.0
\item Versiunea C++: C++17
\item Versiunea IDE: Visual Studio 2019
\item Versiunea CMake: 3.19.2
\end{itemize}

\subsection{Rezultatele evaluării soluțiilor}
Pentru fiecare dintre algoritmii prezentați mai sus, am folosit \textit{hyperfine}
pentru a evalua timpul de rulare al acestora și am salvat datele în câte un fișier
MD. Excepție a făcut algoritmul Brute Force, căruia i-am calculat timpul de rulare
doar pentru testele 1-2, 3, 9-14, 15-16, deoarece pentru celelalte teste timpul de
rulare era prea mare.

Pentru fiecare set de date, am făcut un grafic cu timpul de rulare al fiecărui
algoritm, iar rezultatele sunt prezentate în Figurile \ref{fig:empty_graphs} - \ref{fig:fixed_edge_number_graphs}.

De asemenea, în tabelul \ref{tab:results} sunt prezentate rezultatele obținute
pentru fiecare set de date în raport cu numărul cromatic teoretic al grafurilor (
    acest număr a putut fi precalculat doar pentru testr 1-26
).
Pentru anumite teste care au fost prea costisitoare în timp pentru algoritmul Brute Force, am
lăsat liber câmul respectiv.
\pagebreak

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig1.png}
\caption{Graficele timpului de rulare pentru grafuri goale}
\label{fig:empty_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig2.png}
\caption{Graficele timpului de rulare pentru grafuri complete}
\label{fig:complete_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig3.png}
\caption{Graficele timpului de rulare pentru grafuri bipartite complete}
\label{fig:bipartite_complete_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig4.png}
\caption{Graficele timpului de rulare pentru arbori binari}
\label{fig:binary_tree_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig5.png}
\caption{Graficele timpului de rulare pentru grafuri planare}
\label{fig:planar_graphs}
\end{figure}

\begin{figure}[!]
\centering
\includegraphics[width=0.8\textwidth]{./graphics/fig6.png}
\caption{Graficele timpului de rulare pentru grafuri cu număr de muchii fixat}
\label{fig:fixed_edge_number_graphs}
\end{figure}

\begin{table}[!]
\centering
\caption{Rezultatele obținute pentru fiecare set de date}
\label{tab:results}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\textbf{Nr. set} & \textbf{Nr. noduri} & \textbf{Nr. muchii} & \textbf{Nr. cromatic} & \textbf{Brute Force} & \textbf{Greedy} & \textbf{DSatur} & \textbf{RLF} \\ \hline
\textbf{1} & 887 & 0 & 1 & 1 & 1 & 1 & 1 \\ \hline 
\textbf{2} & 1384 & 0 & 1 & 1 & 1 & 1 & 1 \\ \hline
\textbf{3} & 10 & 45 & 10 & 10 & 10 & 10 & 10 \\ \hline 
\textbf{4} & 50 & 1225 & 50 & 50 & 50 & 50 & 50 \\ \hline 
\textbf{5} & 100 & 4950 & 100 & - & 100 & 100 & 100 \\ \hline 
\textbf{6} & 500 & 124750 & 500 & - & 500 & 500 & 500 \\ \hline 
\textbf{7} & 1000 & 499500 & 1000 & - & 1000 & 1000 & 1000 \\ \hline 
\textbf{8} & 2000 & 1999000 & 2000 & - & 2000 & 2000 & 2000 \\ \hline 
\textbf{9} & 10 & 21 & 2 & 2 & 2 & 2 & 2 \\ \hline 
\textbf{10} & 50 & 504 & 2 & 2 & 2 & 2 & 2 \\ \hline 
\textbf{11} & 100 & 1771 & 2 & 2 & 2 & 2 & 2 \\ \hline 
\textbf{12} & 500 & 35275 & 2 & 2 & 2 & 2 & 2 \\ \hline 
\textbf{13} & 1000 & 164151 & 2 & 2 & 2 & 2 & 2 \\ \hline 
\textbf{14} & 2000 & 557775 & 2 & 2 & 2 & 2 & 2 \\ \hline 
\textbf{15} & 10 & 9 & 2 & 2 & 3 & 2 & 3 \\ \hline 
\textbf{16} & 50 & 49 & 2 & 2 & 3 & 2 & 3 \\ \hline 
\textbf{17} & 100 & 99 & 2 & 2 & 4 & 2 & 4 \\ \hline 
\textbf{18} & 500 & 499 & 2 & - & 4 & 2 & 4 \\ \hline 
\textbf{19} & 1000 & 999 & 2 & - & 4 & 2 & 4 \\ \hline 
\textbf{20} & 2000 & 1999 & 2 & - & 4 & 2 & 4 \\ \hline 
\textbf{21} & 10 & 20 & - & 3 & 3 & 3 & 3 \\ \hline 
\textbf{22} & 50 & 100 & - & 4 & 4 & 3 & 4 \\ \hline 
\textbf{23} & 100 & 200 & - & - & 4 & 4 & 4 \\ \hline 
\textbf{24} & 500 & 1000 & - & - & 4 & 3 & 4 \\ \hline 
\textbf{25} & 1000 & 2000 & - & - & 4 & 4 & 4 \\ \hline 
\textbf{26} & 2000 & 4000 & - & - & 4 & 4 & 4 \\ \hline 
\textbf{27} & 1383 & 10 & - & 2 & 2 & 2 & 2 \\ \hline 
\textbf{28} & 886 & 100 & - & 2 & 2 & 2 & 2 \\ \hline 
\textbf{29} & 777 & 500 & - & 4 & 4 & 3 & 4 \\ \hline 
\textbf{30} & 915 & 1000 & - & - & 4 & 3 & 4 \\ \hline 
\textbf{31} & 1793 & 10000 & - & - & 8 & 6 & 8 \\ \hline 
\textbf{32} & 1386 & 100000 & - & - & 40 & 35 & 40 \\ \hline 
\textbf{33} & 1421 & 500000 & - & - & 167 & 153 & 167 \\ \hline 
\textbf{34} & 1763 & 1000000 & - & - & 276 & 258 & 276 \\ \hline 
\end{tabular}
\end{table}

\subsection{Interpretarea rezultatelor}
În raport cu timpul de execuție, avem următoarele observații:
\begin{itemize}
    \item Algoritmul Brute Force este cel mai lent, \textbf{nu} se folosește
    în practică deloc.
    \item Algoritmul Greedy pe toate tipurile de grafuri este cel mai rapid.
    \item Algoritmul DSatur este mai rapid decât algoritmul Brute Force, dar
    mai lent decât algoritmul Greedy.
    \item Algoritmul RLF pe grafuri goale este mai lent și decât algoritmul
    Brute Force, pe grafuri planare și arbori binari este mai lent decât DSatur, 
    iar pe restul de grafuri este între Greedy și DSatur (aici au apărut cazuri
    care în teorie nu sunt adevărate din cauza complexității agoritmului (vezi \ref{tabel_complexitate}).
    Acest lucru se datorează faptului că nu s-au putut genera teste suficient de
    mari și rulate de un număr suficient de ori pentru a se observa diferența dintre algoritmi.)

\end{itemize}

În raport cu numărul cromatic, se poate observa cu ușurință faptul că algoritmul
Brute Force este un algoritm exact, iar restul sunt aproximări (deși în teste 
nu se observă o diferență între numărul cromatic teoretic și rezultatul de la algoritmul DSatur, dacă
am fi avut un număr mai mare de teste cu numărul cromatic precalculat, ar fi apărut diferența).
Dintre toți1 algoritmii euristici, algoritmul DSatur este cel mai exact.

\section{Concluzii}
În cadrul acestei teme am observat 4 algoritmi diferiți de colorare a grafurilor.

Se poate afirma că nu există o soluție complet generală pentru colorarea grafurilor.
Fiecare din cele prezentate mai sus vin cu avantajele și dezavantajele lor.

Algoritmul Brute Force este cel mai lent, dar este exact, algoritmul Greedy este
cel mai rapid, dar nu este exact, algoritmul DSatur deși este mai lent decât Greedy,
este cel mai exact dintre algoritmi, iar algoritmul RLF este mai lent decât Brute Force și 
DSatur, dar este aproximativ exact.

În practică:
\begin{itemize}
    \item Dacă lucrați cu grafice bipartite sau ciclice, algoritmul Greedy nu este
    suficient de eficient; pe de altă parte, algoritmul RLF este destul de exact și total
    recomandat pentru aceste cazuri.
    \item Dacă prioritar este timpul, se poate înclina spre utilizarea algoritmul Greedy.
    \item Dacă prioritatea este un număr mai apropiat de cel cromatic, trebuie
    să se utilizeze algoritmul DSatur.

\end{itemize}
\pagebreak
% ------------------------------------------------------------------------------

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}

    \bibitem{1}
    J. Bondy and U. Murty, Graph Theory - Graduate Texts in Mathematic.
    Springer, 2008.
    \bibitem{2}
    Z. \`{A}d\`{a}m Mann and A. Szajk\`{o}, “Average-case complexity of backtrack
    search for coloring sparse random graphs,” Journal of Computer and
    System Sciences, vol. 79, no. 8, pp. 1287–1301, 2013.
    \bibitem{3}
    “Backtrack: An o(1) expected time algorithm for the graph coloring
    problem,” Information Processing Letters, vol. 18, no. 3, pp. 119–121,
    1984.
    \bibitem{4}
    N. Barnier and P. Brisset, “Graph coloring for air traffic flow management,” 
    Annals of Operations Research, vol. 130, 03 2002.
    \bibitem{5}
    “The application of a graph coloring method to an examination scheduling problem,” 
    Institute for Operations Research and the Management
    Sciences (INFORMS), vol. 11, no. 5.
    \bibitem{6}
    \href{http://www.ijimt.org/papers/359-D0129.pdf}{University Exam Scheduling System Using Graph
    Coloring Algorithm and RFID Technology}
    \bibitem{7}
    A. Murat and B. Nurdan, “A performance comparison of graph coloring algorithms,” 
    International Conference on Advanced Technology
    Sciences (ICAT’16), vol. 4, pp. 1–19, 12 2016
    \bibitem{8}
    Z. Mann, “Complexity of coloring random graphs: An experimental
    study of the hardest region,” Journal of Experimental Algorithmics,
    vol. 23, pp. 1–19, 03 2018
    \bibitem{9}
    \href{https://acs-aa-challenge.github.io/acs-aa-challenge/18-np-2-colouring/}{Enunțul problemei}
    \bibitem{10}
    M. Garey and D. Johnson, Computer and Intractability: A Guide to the
    Theory of NP-Completeness, 01 1979.
    \bibitem{11}
    D. Porumbel, J.-K. Hao, and P. Kuntz, “An evolutionary approach with
    diversity guarantee and well-informed grouping recombination for graph
    coloring,” Computers Operations Research, vol. 37, pp. 1822–1832, 10
    2010.
    \bibitem{12}
    L. Ouerfelli and H. Bouziri, “Greedy algorithms for dynamic graph coloring,” 2011 International Conference on Communications, Computing
    and Control Applications, CCCA 2011, 03 2011.
    \bibitem{13}
    \`{A}. E. Eiben, J. K. Van Der Hauw, and J. I. van Hemert, “Graph coloring
    with adaptive evolutionary algorithms,” Journal of Heuristics, vol. 4,
    no. 1, pp. 25–46, 1998.
    \bibitem{14}
    D. Br\`{e}laz, “New methods to color the vertices of a graph,”  Commun.
    ACM, vol. 22, pp. 251–256, 04 1979.
    \bibitem{15}
    R. Janczewski, K. Manuszewski, and K. Piwakowski, “The smallest
    hard-to-color graph for algorithm dsatur,” Discrete Mathematics, vol.
    236, pp. 151–165, 06 2001.
    \bibitem{16}
    \href{https://github.com/ifsmirnov/jngen}{GitHub - Jngen: preparing tests made simpler}
    \bibitem{}
    \href{https://en.wikipedia.org/wiki/Graph_coloring}{Graph coloring - Wikipedia}
    \bibitem{}
    \href{https://citeseerx.ist.psu.edu/document?repid=rep1&type=pdf&doi=203a7b17267a28a06808bfb3b0b9571e32d15503}{A Comparison of Parallel Graph Coloring Algorithms}
    \bibitem{}
    \href{https://dergipark.org.tr/en/download/article-file/254140}{A Performance Comparison of Graph Coloring Algorithms}
    % \bibitem{}
    % \href{https://github.com/brrcrites/graph-coloring}{Graph coloring C++ library}
    \bibitem{}
    \href{https://curs.upb.ro/2022/mod/folder/view.php?id=77105}{Moodle - Analiza algoritmilor}
    % 
    \end{thebibliography}
\end{document}

### 2024-04-29 homework1/a_star.py

@@ -6,9 +6,9 @@ import utils
 from heapq import heappush, heappop
 from state import State
 
-def astar(size: (int, int), input_file: str, statistics_file: str, output_file: str) -> None:
+def astar(input_file: str, statistics_file: str, output_file: str) -> None:
     frontier = []
-    initial = State(input_file, size)
+    initial = State(input_file)
     heappush(frontier, (initial.g() + initial.h(), initial))
 
     students_per_subject = str(initial.students_per_subject)
@@ -20,7 +20,7 @@ def astar(size: (int, int), input_file: str, statistics_file: str, output_file: 
         if current.is_final():
             final = current
             break
-        next_states = current.get_next_states_astar()
+        next_states = current.get_next_states()
         with open(statistics_file, 'a', encoding='utf-8') as file:
             file.write(f"Current state: {current.students_per_subject}\n")
             # file.write(f"Next states: {[neighbour.students_per_subject for neighbour in next_states]}\n")
@@ -43,13 +43,12 @@ def astar(size: (int, int), input_file: str, statistics_file: str, output_file: 
     with open(output_file, 'w', encoding='utf-8') as file:
         file.write(s)
 
-def thread_function(
+def run_test(
     name_of_input_file: str, name_of_statistics_file: str, name_of_output_file: str):
     ''' Function that runs the Hill Climbing algorithm in a separate thread. '''
     with open(name_of_statistics_file, 'w', encoding='utf-8') as file:
         file.write("Statistics for the A* algorithm\n\n")
-    size = (6, 5) #TODO: Get the size of the schedule from the input file
-    astar(size, name_of_input_file, name_of_statistics_file, name_of_output_file)
+    astar(name_of_input_file, name_of_statistics_file, name_of_output_file)
 
 if __name__ == '__main__':
     threads = []
@@ -59,11 +58,11 @@ if __name__ == '__main__':
                 'orar_constrans_incalcat',
                 'orar_bonus_exact']
 
-    for nfile in nfiles:
-        input_file = f'inputs/{nfiles}.yaml'
-        statistics_file = f'statistics/astar/{nfiles}.txt'
-        output_file = f'outputs/astar/{nfiles}.txt'
-        thread = threading.Thread(target=thread_function,
+    for i, nfile in enumerate(nfiles):
+        input_file = f'inputs/{nfile}.yaml'
+        statistics_file = f'statistics/astar/{nfile}.txt'
+        output_file = f'outputs/astar/{nfile}.txt'
[…]

### 2024-04-29 homework1/check_constraints.py

@@ -200,18 +200,22 @@ def check_optional_constraints(timetable : {str : {(int, int) : {str : (str, str
 if __name__ == '__main__':
 
     
-    if len(sys.argv) == 1:
-        print('\nSe rulează de exemplu:\n\npython3 check_constraints.py orar_mic_exact\n')
+    if len(sys.argv) != 3:
+        print('\nSe rulează de exemplu:\n\npython3 check_constraints.py astar orar_mic_exact\n')
         sys.exit(0)
 
     if sys.argv[1] == '-h':
-        print('\nSe rulează de exemplu:\n\npython3 check_constraints.py orar_mic_exact\n')
+        print('\nSe rulează de exemplu:\n\npython3 check_constraints.py astar orar_mic_exact\n')
 
-    name = sys.argv[1]
+    algorithm = sys.argv[1]
+    name = sys.argv[2]
+
+    if algorithm != 'astar' and algorithm != 'hc':
+        print('\nAlgoritmul trebuie să fie "astar" sau "hc"!\n')
+        sys.exit(0)
 
     input_name = f'inputs/{name}.yaml'
-    output_name = f'outputs/empty/astar/{name}.txt'
-    # output_name = f'outputs/empty/hc/{name}.txt'
+    output_name = f'outputs/{algorithm}/{name}.txt'
 
     timetable_specs = read_yaml_file(input_name)
 

### 2024-04-29 homework1/hill_climbing.py

@@ -45,16 +45,19 @@ def random_restart_hill_climbing(
     initial_state: State,
     statistics_file: str,
     output_file: str,
-    max_restarts: int = 20,
-    run_max_iters: int = 3000,
+    seed: int,
+    max_restarts: int = 100,
+    run_max_iters: int = 3000
 ) -> Result:
     ''' Random Restart Hill Climbing algorithm with stochastic hill climbing at each restart. '''
     total_iters, total_states = 0, 0
     best_state = None
 
+    # TODO use seed
+    random.seed(seed)
+
     for index in range(max_restarts):
-        state = State(initial_state.file_name, initial_state.size,
-            seed=random.random()) if index > 0 else initial_state.clone()
+        state = State(initial_state.file_name) if index > 0 else initial_state.clone()
         with open(statistics_file, 'a', encoding='utf-8') as file:
             file.write(f"Restarting test {index + 1}...\n")
         is_final, iters, states, state = \
@@ -67,8 +70,8 @@ def random_restart_hill_climbing(
             best_state = state
 
         with open(statistics_file, 'a', encoding='utf-8') as file:
-            file.write(f"Final state evaluation: {state.eval()}\n")
-            file.write(f"Final state constraints: {state.no_constraints()}\n")
+            file.write(f"Final state vs best state evaluation: {state.eval()} / {best_state.eval()}\n")
+            file.write(f"Final state vs best state constraints: {state.no_constraints()} / {best_state.no_constraints()}\n")
 
         if is_final:
             break
@@ -76,56 +79,13 @@ def random_restart_hill_climbing(
     best_state.display(output_file)
     return is_final, total_iters, total_states, state
 
-def run_test(input_file: str, statistics_file: str, output_file: str):
-    ''' Run the test for the Hill Climbing algorithm. '''
-    size = (6, 5) #TODO: Get the size of the schedule from the input file
-    wins, fails = 0, 0
-    total_iters, total_states, distance = 0, 0, 0
-    random.seed(47)
-
-    initials = []
-    for _ in range(N_TRIALS):
-        initials.append(State(input_file, size, seed=random.random()))
-
-    for initial, step in zip(initials, range(N_TRIALS)):
-        with open(statistics_file, 'a', encoding='utf-8') as file:
-            file.write(f"Starting test {step + 1}...\n")
-        is_final, iters, states, state = \
[…]

### 2024-04-29 homework1/orar.py

@@ -1,0 +1,26 @@
+import a_star
+import hill_climbing
+import sys
+
+if __name__ == '__main__':
+	# check if the number of parameters is correct
+	if len(sys.argv) != 3:
+		print('\nSintaxă greșită! Se rulează de exemplu:\n\npython3 orar.py astar orar_mic_exact\n')
+		sys.exit(0)
+
+	# get the 2 parameters from the command line
+	algorithm = sys.argv[1]
+	name = sys.argv[2]
+
+	# check if the algorithm is A* or Hill Climbing
+	if algorithm == 'astar':
+		# run the A* algorithm
+		a_star.run_test(f'inputs/{name}.yaml', f'statistics/astar/{name}.txt', f'outputs/astar/{name}.txt')
+		print(f"\nOutput-ul a fost scris în fișierul outputs/astar/{name}.txt!\n")
+	elif algorithm == 'hc':
+		# run the Hill Climbing algorithm
+		hill_climbing.run_test(f'inputs/{name}.yaml', f'statistics/hc/{name}.txt', f'outputs/hc/{name}.txt')
+		print(f"\nOutput-ul a fost scris în fișierul outputs/hc/{name}.txt!\n")
+	else:
+		print('\nAlgoritmul trebuie să fie "astar" sau "hc"!\n')
+		sys.exit(0)

### 2024-04-29 homework1/state.py

@@ -16,19 +16,18 @@ class State:
     def __init__(
         self,
         file_name: str,
-        size: (int, int) = (6, 5),
+        size: (int, int) = None,
         yaml_dict: dict[str, dict[str, dict[str, str]]] = None,
         schedule: dict[str, dict[str, dict[str, tuple[str, str]]]] | None = None,
         count_techer_slots: dict[str, int] = None,
         students_per_subject: dict[str, int] = None,
-        cost_last_class: int = None,
-        seed: int = 47
+        cost_last_class: int = None
     ) -> None:
-        self.size = size
         self.file_name = file_name
         self.yaml_dict = utils.read_yaml_file_for_hc(file_name) if yaml_dict is None else yaml_dict
+        self.size = size if size is not None else (len(self.yaml_dict[utils.INTERVALS]), len(self.yaml_dict[utils.DAYS]))
         self.schedule = schedule if schedule is not None \
-            else self.generate_empty_schedule(size, self.yaml_dict[utils.CLASSROOMS])
+            else self.generate_empty_schedule(self.size, self.yaml_dict[utils.CLASSROOMS])
         self.students_per_subject = students_per_subject if students_per_subject is not None \
             else copy.deepcopy(self.yaml_dict[utils.SUBJECTS])
         self.count_techer_slots = count_techer_slots if count_techer_slots is not None \
@@ -239,11 +238,6 @@ class State:
         if len(new_states) == 0:
             new_states = self._get_neighbours(False)
 
-        return new_states
-
-    def get_next_states_astar(self) -> list[State]:
-        ''' Returns a list of all possible states that can be reached from the current state. '''
-        new_states = self._get_neighbours(False)
         return new_states
 
     def display(self, output_file: str = None) -> None:

### 2024-04-30 homework1/paper.tex
writing the first 2 sections

@@
 %
-\institute{Facultatea de Automatică și Calculatoare \\
-Universitatea Politehnica București \\
+\institute{Faculty of Automatic Control and Computer Science \\
+University Politehnica of Bucharest \\
 \email{iustina.caramida@stud.acs.upb.ro}
@@
 \begin{abstract}
-    Problema colorării nodurilor unui graf (\textit{graph coloring - GCP} sau \textit{colorarea grafurilor})
-    este una dintre cele mai importante probleme, fapt datorat numărului mare de situații din viața reală
-    care se pot rezuma la o problemă de colorare a unui graf. În această temă, voi prezenta câteva comparări 
-    între diferiți algoritmi dedicați acestei probleme, precum Greedy, Dsatur și RLF din 
-    punct de vedere al timpului de execuție, al memoriei folosite și al performanței în raport cu numărul 
-    de noduri și muchii.
+    With the rapid growth of student enrollment and the expansion of academic 
+    offerings in universities and colleges worldwide, the task of scheduling 
+    classes within existing timetables and facilities has become increasingly 
+    complex. Today, class scheduling requires consideration of multiple factors, 
+    including room availability, capacity, instructors' preferences, and more. This problem is considered to be 
+    NP-complete and has received some research during the past few years. 
+    Several formulations and algorithms have been proposed to solve scheduling
+    problems, most of which are based on local search techniques. In this paper, 
+    we propose to compare 2 different types of algorithms to solve the class
+    scheduling problem: the random restart Hill-Climbing algorithm and the A-Star algorithm.
 
-\keywords{Brute-Force \and Greedy \and DSatur \and RLF \and Colorarea grafurilor.}
+\keywords{NP-complete \and A Star \and Hill-Climbing \and searching algorithms \and class scheduling.}
 \end{abstract}
@@
 
-\section{Introducere}
-\subsection{Descrierea problemei}
+\section{Introduction}
+\subsection{Problem Declaration}
 
-\quad Colorarea grafurilor este o problemă centrală în teoria grafurilor \cite{1}. Ea 
-constă în alegerea unui set de culori pentru nodurile unui graf, astfel încât 
-niciun nod adiacent să nu primească aceeași culoare.
+\quad The Class Scheduling Problem represents a critical challenge in the field 
+of educational operational research.  Universities contend with an ever-growing 
+course catalog, necessitating the efficient allocation of a diverse range of 
+classes to classrooms with varying capacities. This optimization problem seeks 
+to construct a course schedule that adheres to a comprehensive set of university 
+constraints while simultaneously maximizing the effective and efficient 
+utilization of existing facilities.
 
-Problema colorării graficului a început cu încercarea lui 
-Francis Guthries de a colora toate țările pe harta Angliei. 
-Acesta a presupus inițial că patru culori sunt suficiente pentru a procesa orice 
-hartă, astfel încât să nu fie asociate două țări vecine
-cu aceeasi culoare. Această sarcină este doar una dintre cele peste 200 de probleme
-\cite{2}, legate de aria analizei grafice cromatice,
-iar această situație poate fi tradusă prin colorarea fiecărui vârf
-dintr-un graf în care marginile sale ar reprezenta vecinătatea
-dintre două regiuni \cite{3}.
+The complexity of the Class Scheduling Problem presents a multifaceted challenge due 
+to a confluence of factors. First, there is significant variety in the 
+number of students enrolled in each course.  Similarly, classroom capacities 
+exhibit a wide range of variation. These factors are further compounded by the 
+existence of course-specific classroom constraints. Additionally, faculty 
+members often express preferences regarding the day of the week, time slot, 
+and even break schedule for their assigned courses. These factors, combined 
+with the need to satisfy a multitude of university-specific regulations, render 
+the task of assigning courses to classrooms a highly intricate endeavor.
 
-Colorarea grafurilor este o problemă de tip NP-hard bine studiată cu aplicații 
-importante în optimizarea combinatorie și într-un domeniu de cercetare activ, cu 
-multe aplicații practice \cite{4} în inginerie, cum ar fi \textit{alocarea registrelor}, 
-\textit{atribuirea frecvenței}, \textit{potrivirea șablonului} și \textit{programări}. 
-În consecință, colorarea grafurilor a fost subiectul unor cercetări intense \cite{5} \cite{2}.
+Merely ensuring a classroom can accommodate a course's enrolled students is 
+insufficient. Such an approach would lead to suboptimal space utilization, 
+potentially hindering the educational experience and fostering student 
+dissatisfaction. Consider a scenario with two courses: one with six students 
+and another with nineteen. Furthermore, imagine two classrooms exist, one 
+seating twenty and another seating fifty. While technically feasible to 
+schedule either course in either room, a more strategic allocation would be to 
+assign the larger course to the larger classroom.
 
-Un exemplu de o problemă de colorare a grafurilor este planificarea examenelor, unde
-fiecare examen este reprezentat de un nod și fiecare legătură dintre nodiri reprezintă
-o interzicere de a planifica examenele. Astfel, colorarea grafului reprezintă planificarea
-examenelelor, astfel încât două examene care nu pot fi planificate în același timp să fie
-colorate diferit. În acest caz, numărul minim de culori necesare pentru a colora graful
-este egal cu numărul minim de zile necesare pentru a planifica examenele. \cite{6}
+This example exemplifies the multifaceted nature of Class Scheduling Problem. 
+Researchers continue to explore sophisticated methods to address this intricate 
+problem, with the ultimate goal of ensuring the smooth operation and optimal 
+resource allocation within universities.
 
-Colorarea grafurilor este asociată cu două tipuri de colorare: colorarea vârfurilor 
-și colorarea marginilor. Scopul ambelor tipuri de colorare este de a colora întregul graf fără
-conflicte. Prin urmare, vârfurile adiacente sau muchiile adiacente trebuie
-să fie în culori diferite. Numărul cel mai mic de culori necesare 
-pentru a colora graful se numește \textbf{număr cromatic}.
-Pe măsură ce numărul de vârfuri sau muchii dintr-un graf crește,
-complexitatea problemei crește și ea. Din această cauză, fiecare
-algoritmul nu poate găsi numărul cromatic exact  
-și pot fi, de asemenea, diferențe în timpul lor de execuție \cite{7}. Pentru a obține 
-o soluție mai bună pentru colorarea grafurilor, mulți algoritmi euristici și meta
-euristici au fost inventați \cite{8}.
+\subsection{Problem Statement}
+\textbf{\textit{Statement:}} Receiving a set of courses, each with a specific
+number of students,a set of rooms with a specific capacity, and a list of 
+instructors with their list of preferences, the goal is to assign each course to
+a room and a time slot, according to a list of constraints.
+
+There are 2 types of constraints that need to be satisfied:
+\begin{itemize}
+    \item \textbf{Hard constraints:} These constraints must be satisfied in order
+    for the solution to be valid. If any of these constraints are not satisfied, 
+    the solution is invalid.
+    \item \textbf{Soft constraints:} These constraints are not mandatory, but 
+    they are taken into account when evaluating the quality of the solution.
+\end{itemize}
+
+\subsubsection*{Hard constraints}
+\begin{itemize}
+    \item Within a designated time slot and in a given room, only one subject may be taught by a single instructor.
+    \item During any given time slot, an instructor may teach only one subject, and this instruction must occur within a singular room.
+    \item An instructor may conduct classes in a maximum of 7 time slots per week.
+    \item Within a specified time slot, a room may accommodate a number of students equal to or less than its predetermined maximum capacity.
+    \item All students enrolled in a particular subject must have designated class hours allocated for that subject.
+    \item Instructors are restricted to teaching only the subjects in which they are specialized.
+    \item All rooms are designated for classes pertaining only to the subjects for which they have been assigned
+\end{itemize}
+
+\subsubsection*{Soft constraints}
+\begin{itemize}
+    \item An instructor may express preferences regarding specific days of the week for teaching or may wish to avoid teaching on certain days.
+    \item An instructor may have preferences regarding specific time slots during the day for teaching or may wish to avoid teaching during certain time slots.
+    \item An instructor may prefer not to have a break exceeding a certain number of hours between consecutive classes.
+\end{itemize}
 %
-\subsection{Specificarea algoritmilor aleși}
 
-\textbf{\textit{Cerință:}} Fie un graf neorientat G cu N noduri și M muchii. Problema cere să asociem o culoare 
-fiecărui nod, astfel încât oricare două noduri adiacente (conectate printr-o muchie directă) 
-să aibă culori diferite. Care este numărul minim de culori necesare pentru a colora 
-toate nodurile conform restricției menționate anterior? \cite{9}
+\subsection{Methods of Approach}
+\subsubsection{Random Restart Hill Climbing}
+is a local search algorithm commonly applied in optimization 
+problems. It begins with an initial solution and iteratively explores 
+neighboring solutions, selecting the one that improves the objective function 
+the most. This process continues until a local optimum is reached, where no 
+better solution can be found in the immediate neighborhood. When the algorithm
+reaches a local optimum, it restarts the search from a random initial solution.
+The algorithm terminates when a specified number of iterations have been
+completed or when a solution that satisfies all constraints is found.
 
-\subsubsection{Brute Force:}
-Când încercăm să oferim o soluție la această problemă, primul instinct este de a
-folosi o abordare ”Brute Force”. Acest lucru ar duce la o soluție, care ar fi 
-din punct de vedere al timpului de execuție exponențială, făcând această soluție 
-inutilă pentru cazuri mari. A spune că nu putem găsi un algoritm eficient 
-deoarece acesta nu există ar fi la fel ca și când am spune că problema nu are 
-nicio soluție eficientă \cite{10}.
-În prezent, există algoritmi care se ocupă să rezolve problema colorării grafurilor, 
-deși obțin un număr cromatic apropiat de ale grafului în schimbul unui timp
-rezonabil sau rezultate rapide care sunt suficient de utile \cite{11}.
+\subsubsection{A Star}
+is an informed search algorithm widely used for pathfinding and graph 
+traversal tasks. It intelligently combines both actual path costs and heuristic 
+estimates to guide the search towards the goal efficiently. A* maintains a 
+priority queue of nodes to be explored and selects the most promising node 
+based on a combination of the cost incurred so far and the estimated cost to 
+reach the goal. This allows A* to efficiently find the optimal path while 
+intelligently pruning the search space, making it highly effective for solving 
+a wide range of optimization problems.
 
-\subsubsection{Greedy Algorithm:}
-Logica algoritmului ia vârfurile grafului unul câte unul, urmând o ordine 
-(care poate fi aleatorie) și atribuie prima culoare disponibilă fiecărui vârf \cite{12}. 
-Deoarece este un algoritm euristic, soluția oferită de acesta poate să nu fie optimă.
-Cu toate acestea, o alegere corectă a ordinii vârfurilor pentru
-colorarea lor poate oferi o soluție optimă pentru orice graf. În
-practică, algoritmul Greedy produce soluții rapid practicabile, deși aceste soluții 
-pot fi “sărace” pe baza numărului de culori pe care algoritmul le cere, în comparație cu
-numărul cromatic al grafului.
+\subsection{Evaluation}
+\quad Both algorithms will be evaluated based on the quality of the solutions they
+provide, the time required to find these solutions and the total number of
+states explored during the search. The quality of the solutions will be
+evaluated based on the number of constraints satisfied.  The time required to 
+find the solutions will be measured in milliseconds, and the total number of
+states explored will be counted during the search process.
 
-\subsubsection{DSatur Algorithm:}
-Algoritmul DSatur (abreviere din engleză pentru “Degree Saturation”), 
-propus de Brelaz (1979), se comportă foarte asemănător cu
-algoritmul Greedy, cu excepția că, în acest caz,
-ordonarea vârfurilor este generată de algoritmul însuși.
-La fel ca în algoritmul Greedy, ordonarea a fost decisă
-înainte ca orice vârf să fie colorat, ordinea vârfurilor fiind decisă euristic pe baza
-caracteristicilor colorării parțiale a grafului la
-momentul în care se selectează fiecare dintre vârfuri \cite{13}. 
-În cel mai rău caz, complexitatea sa are aceeași situație ca și în
-Algoritmul Greedy, deși în practică poate fi luat în considerare
-și faptul că monitorizare saturaţiei vârfurilor necolorate
-produce o complexitate puțin mai mare. Este de reținut că
-Algoritmul DSatur este \textbf{exact} pentru grafurile bipartite \cite{14}.
+There will be 5 types of tests, describe in \textit{Tests Description} section.
 
-\subsubsection{RLF Algorithm:}
-Algoritmul RLF (abreviere din engleză pentru “Recursive Largest First”), propus
-de Leighton (1979), lucrează prin colorarea unui graf cu o singură culoare
-pentru fiecare iterație a algoritmului, în loc de un vârf per
-repetare. În fiecare iterație, algoritmul caută seturi de
-vârfuri independente din graf, care vor fi asociate
-cu aceeași culoare. Acel set independent a fost eliminat
-din graf, iar subgraful rămas va continua
-în același mod, până când subgraful menționat este gol, caz în care
-toate vârfurile vor fi atribuite unei culori, producând astfel o
-soluție ce satisface toate cerințele \cite{12}.
+% ------------------------------------------------------------------------------
+\section{Algorithm Design}
+\subsection{State Representation}
 
-\subsection{Evaluarea soluțiilor}
-Sursele vor fi testate pe grafuri de dimensiuni diferite, de la 10 la 2 000 de noduri,
-cu un număr de muchii de la 10 la 1 000 000. Pentru fiecare graf se va genera o
-configurație aleatoare de noduri și muchii, iar apoi se va testa performanța
-algoritmilor pe aceste grafuri. Pentru fiecare algoritm, se va calcula timpul de
-execuție și numărul de culori folosite pentru a colora graful.
+The state representation for the Class Scheduling Problem consists of a
+schedule that assigns each course to a room and a time slot. The state is
+represented as a class that contains the following fields:
+\begin{itemize}
+    \item \textbf{file\_name:} The name of the file from which the data was read.
+    \item \textbf{yaml\_dict:} A dictionary containing the data read from the file.
+    \item \textbf{size:} A tuplet (days, time\_slots) representing the size of the schedule.
+    \item \textbf{schedule:} A dictionary that has the following structure: \{(day: str) : \{(time\_slot: (int, int)) : \{(classroom : str) : ((instructor : str), (subject : str))\}\}\}.
+    \item \textbf{strudents\_per\_subject:} A dictionary that contains the number of students for each subject that needs to be scheduled.
+    \item \textbf{count\_techer\_slots:} A dictionary that contains the number of scheduled slots for each instructor.
+    \item \textbf{trade\_off:} A number that represents the trade-off between the number of constraints satisfied and the the choosen classroom at each step (used in A Star).
+\end{itemize}
 
-\section{Prezentarea soluțiilor}
-\subsection{Algoritmul Brute Force}
-Algoritmul Brute Force este un algoritm care se bazează pe forță bruta, adică
-parcurge toate posibilitățile de colorare a grafului și alege cea mai bună
-soluție. Complexitatea acestui algoritm este de $O(n^m)$, unde $n$ este numărul de noduri și $m$ este
-numărul de culori. Pentru grafuri cu un număr mare de noduri, complexitatea
-acestui algoritm este foarte mare, de aceea nu este folosit în practică. Pe de altă parte,
-algoritmul este exact, adică oferă întotdeauna o soluție optimă. Un pseudocod al
-algoritmului este prezentat în Pseudocodul~\ref{alg1}.
+\subsubsection{Initial State}
+The initial state is generated as an empty schedule, which is initialized with 
+the parameters derived from the input file, including the number of days, time 
+slots, and classrooms. Additionally, the number of students per subject is 
+obtained from the input file, while the count of scheduled slots for each 
+instructor is set to zero.
+
+This initial state serves as the starting point for the search algorithms, 
+enabling them to iteratively allocate courses to rooms and time slots until 
+a valid schedule is achieved.
+
+An alternative method for generating the initial state involves randomly 
+assigning courses, instructors to rooms and time slots in a manner that 
+adheres to the hard constraints. A comparative analysis of these two approaches 
+will be conducted in the \textit{Initial State Selection} section.
+
+\subsubsection{Generating Neighbors}
+The neighbors of a state are generated by considering all possible combinations 
+of assigning a course to a room, a time slot and an instructor, while ensuring 
+adherence to the hard constraints.
+
+In the initial phase, all potential neighbors that adhere to both the hard and 
+soft constraints are generated. Subsequently, in the event that no neighbors 
+satisfying all soft constraints are found, a secondary phase ensues where only 
+neighbors satisfying the hard constraints are generated. This strategy reduces 
+the total number of neighbors generated, allowing the algorithm to prioritize 
+those that satisfy all constraints. Consequently, the algorithm minimizes time 
+wastage by avoiding the generation of neighbors that would not be utilized, 
+resulting in a reduced number of states generated.
+
+\subsubsection{Initial State Selection}
+As previously mentioned, the initial state can be generated in two methodologies: either 
+as an empty schedule or through the random assignment of courses, instructors, 
+and rooms to time slots. The former method exhibits a greater degree of 
+determinism, initializing the schedule with vacant slots, whereas the latter 
+introduces stochasticity into the initial state generation process.
+
+Upon experimentation with both approaches, it became evident that the random 
+initialization method could often yield solutions that fail to adhere to all 
+specified soft constraints. This underscores the importance of carefully 
+considering all potential subsequent states that may arise from the current 
+state when designing the random initialization method.
+
+For instance, in instances where an instructor's preferences are violated within 
+a particular time slot, the algorithm should endeavor to substitute instructor with 
+another whose preferences remain unviolated for that time slot. Moreover, the 
+algorithm should endeavor to substitute the time slot with another or substitute
+the room with anothers that have the same capacity in total. This approach
+was found more difficult to implement. When tring to implement the first two
+substitutes, the algorithm was not able to find a solution that satisfies all the
+soft constraints and it took quite a lot of time to find a partial solution due to
+a large number of constraints that should be checked while creating the neighbors.
+On the other hand, this approach will always find a partial solution that 
+satisfies all the hard constraints. Thus, this one is recommended in cases 
+when the soft constraints are not that important.
+
+The method that I used in 
+the end was the empty schedule initialization. It is acknowledged that without 
+the random initialization component, the Hill-Climbing algorithm may struggle 
+to assign all students to a room, thereby violating a hard constraint. However,
+with  the inclusion of the random restart, the algorithm will always find a 
+solution that satisfies all the hard constraints, the soft ones being satisfied 
+in most of the cases. This approach is recommended in cases when the soft 
+constraints are more important than in the previous case and it is easier to 
+generate the neighbors.
+
+\subsection{Random Restart Hill Climbing}
+
+As previously mentioned, the Random Restart Hill Climbing algorithm is the 
+most suitable for this approach. The algorithm is initialized with an empty
+schedule and generates neighbors that adhere to the hard constraints. The
+algorithm iteratively explores the neighborhood of the current state, selecting
+randomly one of the neighbors that satisfies the most constraints. This process continues until
+a local optimum is reached, at which point the algorithm restarts the search.
+The algorithm terminates when a solution that
+satisfies all constraints is found or when a specified number of iterations / 
+restarts have been completed. A pseudocode of the algorithm is presented in Algorithm
+\ref{alg1}.
 
 \begin{algorithm}
-\caption{Brute Force Algorithm}
+\caption{Random Restart Hill Climbing Algorithm}
 \label{alg1}
 \begin{algorithmic}[1]
-\Procedure{BruteForce}{$G$} \Return $List\langle int\rangle $ 
-\State $List\langle int\rangle $ $colors = new List\langle int\rangle $ \Comment{List of colors}
-\State $int$ $maxColors = 0$ \Comment{Number of colors}
-\State $int$ $maxColorsIndex = 0$ \Comment{Index of the best solution}
+\Procedure{hill\_climbing}{$max\_restarts$} \Return $[is\_final, total\_iters, total\_states, best\_state] $ 
+\State $total\_iters = 0$
+\State $total\_states = 0$
+\State $best\_state = None$
+
+\For{$index$ in $range(max\_restarts)$}
+\State $state = InitialState()$
+\State $is\_final, iters, states, state = stochastic\_hill\_climbing(state)$
+\State $total\_iters += iters$
+\State $total\_states += states$
+
+\If{$is\_final$}
+\State \Return $[is\_final, total\_iters, total\_states, state]$
+\EndIf
+
+\If{$state\ does\ not\ have\ hard\ constraints$}
+\If{$best\_state == None$ or $state\ has\ less\ soft\ constraints$
+\hspace*{6em}$unsatisfied\ than\ best\_state$}
+
+\State $best\_state = state$
+\EndIf
+\EndIf
+\EndFor
+
+\State \Return $[is\_final, total\_iters, total\_states, best\_state]$
+
 \algstore{bkbreak}
@@
 \begin{algorithm}
-\caption{Brute Force Algorithm}
+\caption{Stochastic Hill Climbing}
 \begin{algorithmic}[1]
-\algrestore{bkbreak}
-\State $int$ $i = 0$ \Comment{Index of the current solution}
-\State $int$ $n = G.GetNumberOfNodes()$ \Comment{Number of nodes}
-\State $int$ $m = G.GetNumberOfColors()$ \Comment{Number of colors}
-\State $int$ $max = pow(m, n)$ \Comment{Number of possible solutions}
-\While{$i < max$}
-\State $colors = G.GetColors()$
-\State $int$ $currentColors = colors.Count()$
-\If{$currentColors > maxColors$}
-\State $maxColors = currentColors$
-\State $maxColorsIndex = i$
+\Procedure{stochastic\_hill\_climbing}{$state, max\__iters$} \Return $[is\_final, total\_iters, total\_states, best\_state] $ 
+\State $total\_iters = 0$
+\State $total\_states = 0$
+
+\While{$total\_iters < max\_iters$}
+\State $total\_iters += 1$
+
+\If{$state\ is\ final$}
+\Return $[True, total\_iters, total\_states, state]$
 \EndIf
-\State $i++$
+
+\State $neighbors = state.generate\_neighbors()$
+\State $total\_states += len(neighbors)$
+
+\If{$neighbors$ == $None$}
+\Return $[False, total\_iters, total\_states, state]$
+\EndIf
+
+\State $state = random.choice(from\ neighbors\ one\ of\ the\ neighbors\ with$ \\
+\hspace*{2em}$\text{minimum\ number\ of\ constraints\ unsatisfied)}$
 \EndWhile
-\State $colors = G.GetColors(maxColorsIndex)$
-\State \Return $colors$
+
+\State \Return $[False, total\_iters, total\_states, state]$
+
 \EndProcedure
@@
 
-\subsection{Algoritmul Greedy}
-Algoritmul Greedy este un algoritm care alege mereu cea mai bună soluție la
-momentul curent, fără a ține cont de soluțiile viitoare.
-Un pseudocod al algoritmului este prezentat în
-Pseudocodul~\ref{alg2}.
+\subsection{A Star}
+For the A Star algorithm, the state representation is the same as for the Random
+Restart Hill Climbing algorithm. The algorithm is initialized with an empty
+schedule and generates neighbors that adhere to the hard constraints. The 
+algorithm iteratively explores the neighborhood of the current state.
+
+The frontier represents a heap that contains the states that need to be explored.
+The diccovered is a dictionary that contains as keys the number of students that
+need to be scheduled for each subject and as values the cost of the state that
+brought us to this configuration.
+
+The function used in the A Star algorithm is:
+\begin{equation}
+    f(state) = g(state) + h(state)
+\end{equation}
+
+where
+
+\begin{equation}
+    h(state) = total\ number\ of\ students\ that\ are\ not\ assigned
+\end{equation}
+\begin{equation}
+    g(state) = number\ of\ constraints\ unsatisfied * weight + trade\_off
+\end{equation}
+\begin{equation}
+    trade\_off = \frac{number\ of\ classrooms(subject)}{total\ number\ of\ classrooms} 
+\end{equation}
+
+The heuristic function is admisible, because the return value is always less than
+or equal to the actual cost of the state to reach a final one, and is
+equal to 0 in the final states. On the other hand, the function h is not consistent,
+because a state from discovered can be added to the frontier with a smaller cost.
+
+The cost function is calculated as the number of constraints unsatisfied multiplied
+by a weight and the trade off. In cases where the number of constraints unsatisfied
+is equal, multiplying it by a weight will prioritize the states that have the trade
+off smaller. Moreover, when the number of constraints unsatisfied is different, the
+prioritization will be made based on the number of constraints unsatisfied, not on
+the trade off.
+
+The trade off is calculated as the number of classrooms that are assigned to a
+subject divided by the total number of classrooms. We want to prioritize adding
+a subject to a classroom that has fewer subjects assigned to it. For instance,
+if we have 2 subjects: A and B, 2 classrooms: 1 and 2, and classroom 1 is assigned
+to subject A and classroom 2 is assigned to both subjects, while tring to assign
+a subject to classroom 2 we will choose subject B (trade off = 0.5) instead of
+subject A (trade off = 1).
+
+The pseudocode of the algorithm is presented in Algorithm \ref{alg2}.
 
 \begin{algorithm} 
-\caption{Greedy Algorithm}
+\caption{A Star Algorithm}
 \label{alg2}
 \begin{algorithmic}[1]
-\Procedure{Greedy}{$G$}
-\Require{$S$ = Class Set, $V$ = Colorless vertices in random order}
-\Ensure{$S$ = $\emptyset$}
-\For{$v \in V$}
-\For{$i$ to $S.lenght$}
-\If {$NonConflitiveEdges(v \cup S_i)$}
-\State $AssignClass(v, S_i)$
-\State $Next Vertex$
+\Procedure{astar}{} \Return $[is\_final, total\_iters, total\_states, best\_state] $
+\State{$frontier = []$}
+\State{$discovered = \{\}$}
+\State{$state = InitialState()$}
+\State{$frontier.append((f(state), state))$}
+\State{$discovered[state] = 0$}
+\State{$total\_iters = 0$}
+\State{$total\_states = 1$}
+\While{$frontier$}
+\State{$current\_state = frontier.pop(1)$}
+\State{$total\_iters += 1$}
+\If{$current\_state\ is_final == 0$}
+\Return $[True, total\_iters, total\_states, current\_state]$
+\EndIf
+
+\State{$neighbors = current\_state.generate\_neighbors()$}
+\State{$total\_states += len(neighbors)$}
+\For{$neighbor$ in $neighbors$}
+\State{$new\_cost = g(neighbor) + h(neighbor)$}
+\State{students\_per\_subject = neighbor.students\_per\_subject}
+\If{$students\_per\_subject\ not\ in\ discovered$ or $new\_cost < discovered[students\_per\_subject]$}
+\State{$discovered[neighbor] = new\_cost$}
+\State{$frontier.append((new\_cost, neighbor))$}
 \EndIf
 \EndFor
-\If {$NotColored(v)$}
-\State $S_{i+1} = NewClass$
-\State $AssignClass(v, S_{i+1})$
-\State {Next Vertex}
-\EndIf
-\EndFor
+\EndWhile
+\State \Return $[False, total\_iters, total\_states, current\_state]$
 \EndProcedure
@@
 
-În algoritmul Greedy, așa cum se poate vedea în pseudocod,
-se folosește o permutare inițială care este generată aleatoriu și în
-care este colorat fiecare vârf, comparând în fiecare caz dacă acesta
-poate fi inclus într-o culoare fără a provoca conflicte. Pentru a analiza
-algoritmul în cel mai rău caz, vom presupune că fiecare vârf este 
-verificat cu fiecare culoare înainte de a fi colorat,
-ce ne dă un polinom care este
-\begin{equation} P(x) = n(n+1)/2. \end{equation} Polinomul
-se obţine din faptul că fiecare vârf, înainte de a fi colorat,
-trebuie să se compare cu toate culorile actuale:
-\begin{equation} P = 0 + 1 + 2 + . . . + n = n(n + 1)/2 \end{equation}
+\subsection{Complexities}
+The complexity of the Random Restart Hill Climbing algorithm is $O(n)$, where
+$n$ is the total number of iterations. The complexity of the A Star algorithm is
+$O(b^d)$, where $b$ is the branching factor and $d$ is the depth of the solution.
 
-În cele din urmă, vom avea o complexitate de $O(n^2)$, unde $n$ este numărul
-de noduri. Este necesar să afirmăm că algoritmul Greedy, deși calculează
-numărul cromatic aproximativ, acest lucru va depinde direct de permutarea
-inițială pe care o avem. Cu algoritmul următor vom vedea tehnici care ne
-permit să îmbunătățim modul de a alege vârfurile pentru a obține soluții
-mai bune și mai apropiate de numărul cromatic.
+The complexity of the generate neighbors function is $O(d*t*c*i*s)$, where $d$ is
+the number of days, $t$ is the number of time slots, $c$ is the number of classrooms,
+$i$ is the number of instructors and $s$ is the number of subjects. Because the number
+of days does not exceed 7 (the worst case) and the number of time slots does not exceed 12, 
+the complexity becomes $O(c*i*s)$.
 
-\subsection{Algoritmul DSatur}
-Algoritmul DSatur este un algoritm care alege mereu nodul cu cea mai mare
-valoare de saturație, adică nodul care are cel mai mare număr de vecini cu
-aceeași culoare. Un pseudocod al algoritmului este prezentat în Pseudocodul~\ref{alg3}.
-\begin{algorithm}
-\caption{DSatur Algorithm}
-\label{alg3}
-\begin{algorithmic}[1]
-\Procedure{DSatur}{$G$}
-\Require{$S$ = Class Set, $V$ = Non-colored vertices}
-\Ensure{$S$ = $\emptyset$}
-\While {$V \neq \emptyset$}
-\State $v = MaxSaturation(V)$
-\For {$i$ to $S.lenght$}
-\If {$NonConflitiveEdges(v \cup S_i)$}
-\State $AssignClass(v, S_i)$
-\State {Next Vertex (break)}
-\EndIf
-\If {$NotColored(v)$}
-\State $S_{i+1} = NewClass$
-\State $AssignClass(v, S_{i+1})$
-\State $Remove(v, V)$
-\State {Next Vertex}
-\EndIf
-\EndFor
-\EndWhile
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
+\section{Evaluation}
+\subsection{Tests Description}
 
-Se poate observa din pseudocod că acest algoritm este foarte asemănător cu algoritmul
-Greedy, însă puterea algoritmului DSatur constă în prioritatea dată nodurilor cu cea mai
-mare saturație. Astfel, aceste noduri mai restrânse sunt colorate înainte de restul nodurilor,
-care nu au aceste “restricții”. Complexitatea algoritmului este similară cu cea a algoritmului
-Greedy, deoarece este foarte asemănător în cod, însă are și calcule suplimentare pentru a
-obține saturația nodurilor, însă acestea nu afectează direct complexitatea totală a algoritmului,
-care este totuși $O(n^2)$.
+\subsubsection{orar\_mic\_exact:} Contains 3 courses, 2 rooms, and 13 instructors. The
+number of students for each course are 300, 330, and 330. The capacity of the rooms are
+20 and 30. The instructors have few preferences.
+\subsubsection{orar\_mediu\_relaxat:} Contains 4 courses, 4 rooms, and 18 instructors. The
+number of students for each course is 660, 660, 665 and 685. The capacity of the rooms is
+25, 25, 35 and 70. The instructors have few preferences.
+\subsubsection{orar\_mare\_relaxat:} Contains 8 courses, 6 rooms, and 37 instructors. The
+number of students for each course is 470, 475, 475, 495, 500, 530, 535 and 550. The capacity of the rooms is
+25, 30, 30, 35, 85 and 85. The instructors have few preferences.
+\subsubsection{orar\_constrans\_incalcat:} Contains 4 courses, 2 rooms, and 17 instructors. The
+number of students for each course is 720, 750, 780 and 810. The capacity of the rooms is
+15 and 90. The instructors have a lot of preferences.
+\subsubsection{orar\_bonus\_exact:} Contains 5 courses, 5 rooms, and 23 instructors. The
+number of students for each course is 500, 510, 515, 520 and 545. The capacity of the rooms is
+15, 15, 15, 15 and 50. The instructors have a lot of preferences, including breaks constraints.
 
-Trebuie punctat faptul că, deși algoritmul DSatur are o performanță mai bună,
-acesta are și cazuri în care obține o soluție
-mai proastă decât algoritmul Greedy. Aceasta se poate observa în \cite{15} unde se arată că o
-distribuția neadecvată a vârfurilor este dificil de colorat cu algoritmul DSatur.
-
-\subsection{Algoritmul RLF}
-Algoritmul RLF caută să coloreze toate vârfurile disponibile în momentul în care
-acestea nu afectează niciun conflict al aceleiași culori. De asemenea, acordă
-prioritate vârfurilor care au un grad mai mare, acordând astfel prioritate tipului
-nostru de restricție prevăzut pentru această euristică. Algoritmul RLF are
-o complexitate în cel mai rău caz mai mare decât cea a algorimului Greedy sau
-DSatur, aceasta fiind o complexitate dovedită și de Leighton (1979) - $O(n^3)$.
-Cu toate acestea acest algoritm are și o îmbunătățire a calității soluțiilor pe
-care le oferă, deşi la un cost de calcul mai mare. Mai jos se poate observa pseudocodul~\ref{alg4} algoritmului RLF.
-
-\begin{algorithm}
-\caption{RLF Algorithm}
-\label{alg4}
-\begin{algorithmic}[1]
-\Procedure{RLF}{$G$}
-\Require{$S$ = Class Set, $V$ = Competent vertices, $W$ = = Non-Competent vertices}
-\Ensure{$S$ = $\emptyset$, $V$ = Vertices of a given graph, $W$ = $\emptyset$}
-\While {$V \neq \emptyset$}
-\State $S_i = NewClass$
-\While {$V \neq \emptyset$}
-\State $v = MaxSaturation(V)$
-\State $AssignClass(v, S_i)$
-\State $Add($ Adjacent Vertices $(v), W)$
-\State $Remove($ Adjacent Vertices $(v), W)$
-\EndWhile
-\State $V = W$
-\State $W = \emptyset$
-\EndWhile
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexități}
-Mai jos, tabelul~\ref{tabel_complexitate} conține complexitățile algoritmilor
-prezentați mai sus.
-
-\begin{table}[h]
-\centering
-\caption{Complexități algoritmilor}
-\label{tabel_complexitate}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-
-\textbf{Algoritm} & \textbf{Complexitate} & \textbf{Complexitate} & \textbf{Complexitate} & \textbf{Complexitate} \\
- & \textbf{Brute Force} & \textbf{Greedy} & \textbf{DSatur} & \textbf{RLF} \\
-\hline
-\hline
-\textbf{Cazul bun} & $O(n^m)$ & $O(n)$ & $O(n)$ & $O(n)$ \\
-\hline
-\textbf{Cazul rău} & $O(n^m)$ & $O(n^2)$ & $O(n^2)$ & $O(n^3)$ \\
-\hline
-\end{tabular}
-\end{table}
-
-\section{Evaluare}
-\subsection{Grafuri de test și execuția algoritmilor}
-Pentru a evalua performanța algoritmilor de colorare, am folosit grafuri de test
-generate cu ajutorul unui generator de teste de pe Github, ce îl puteți accesa aici \cite{16}.
-Am generat 34 de teste prezentate în tabelul~\ref{tab1} pentru toți algoritmii
-prezentați mai sus.
-
-\begin{longtable}{@{}|c|c|c|c|@{}}
-    \hline\hline
-    numărul testului & tipul grafului  & numărul de noduri (N)  & numărul de muchii (M) \\
-    \hline
-    1 & graf gol  & 887  & 0 \\
-    2 &  & 1384 & 0 \\ \hline
-    3 & graf complet & 10 & 45 \\
-    4 &  & 50 & 1225 \\
-    5 &  & 100 & 4950 \\
-    6 &  & 500 & 124750 \\
-    7 &  & 1000 & 499500 \\
-    8 &  & 2000 & 1999000 \\ \hline
-    9 & graf bipartit & 10 & 21 \\
-    10 & complet & 50 & 504 \\
-    11 &  & 100 & 1771 \\
-    12 &  & 500 & 35275 \\
-    13 &  & 1000 & 164151 \\
-    14 &  & 2000 & 557775 \\ \hline
-    15 & arbore binar & 10 & 9 \\
-    16 &  & 50 & 49 \\
-    17 &  & 100 & 99 \\
-    18 &  & 500 & 499 \\
-    19 &  & 1000 & 999 \\
-    20 &  & 2000 & 1999 \\ \hline
-    21 & graf planar & 10 & 20 \\
-    22 &  & 50 & 100 \\
-    23 &  & 100 & 200 \\
-    24 &  & 500 & 1000 \\
-    25 &  & 1000 & 2000\\
-    26 &  & 2000 & 4000 \\ \hline
-    27 & graf cu număr & 1383 & 10 \\
-    28 & de muchii fixat & 886 & 100 \\
-    29 &  & 777 & 500 \\
-    30 &  & 915 & 1000 \\
-    31 &  & 1793 & 10000 \\
-    32 &  & 1386 & 100000 \\
-    33 &  & 1421 & 500000 \\
-    34 &  & 1763 & 1000000 \\
-    \hline
-    \caption{Grafuri de test}
-    \label{tab1}
-    \end{longtable}
-
-\subsubsection{Descrierea testelor \\}
-
-\textbf{Grafuri goale:} Acestea sunt grafuri care nu au nicio muchie. Numărul cromatic al acestor grafuri este 1, deoarece toate nodurile se pot colora la fel.
-
-\textbf{Grafuri complete:} Acestea sunt grafuri care au toate nodurile conectate
-între ele. Numărul cromatic al acestor grafuri este egal cu numărul de noduri.
-
-\textbf{Grafuri bipartite complete:} Acestea sunt grafuri bipartite (=  graf care
-poate fi împărțit în două multimi disjuncte
-de noduri astfel încât orice muchie din graf să unească un nod din prima mulțime
-cu un nod din a doua mulțime) care au proprietatea
-că pentru orice nod x din A şi orice nod y din B există muchia [x, y].
-Numărul cromatic al acestor grafuri este egal cu 2.
-
-\textbf{Arbori binari:} Aceștia sunt grafuri care au proprietatea că orice nod
-are cel mult 2 vecini. Numărul cromatic al acestor grafuri este egal cu 2.
-
-\textbf{Grafuri planare:} Acestea sunt grafuri care pot fi reprezentate pe o
-suprafață plană fără ca muchiile acestora să se intersecteze. Numărul cromatic
-al acestor grafuri este cel mult egal cu 4.
-
-\textbf{Grafuri cu număr de muchii fixat:} Acestea sunt grafuri care au un număr
-fix de muchii. Numărul cromatic al acestor grafuri nu se poate precalcula.
-
-\subsection{Specificațiile sistemului de calcul}
-Codul a fost builduit în C++, iar fiecare dintre algoritmii prezentați mai sus au fost
-evaluați folosind un Laptop cu următoarele specificații:
-\begin{itemize}
-\item Procesor: Intel(R) Core(TM) i7-10750H CPU @2.60GHz   2.59 GHz
-\item Memorie RAM: 16 GB
-\item Sistem de operare: Windows 10 Pro 64-bit
-\item Versiunea compilatorului: GCC 9.4.0
-\item Versiunea C++: C++17
-\item Versiunea IDE: Visual Studio 2019
-\item Versiunea CMake: 3.19.2
-\end{itemize}
-
-\subsection{Rezultatele evaluării soluțiilor}
+\subsection{Results}
 Pentru fiecare dintre algoritmii prezentați mai sus, am folosit \textit{hyperfine}
@@
 \textbf{5} & 100 & 4950 & 100 & - & 100 & 100 & 100 \\ \hline 
-\textbf{6} & 500 & 124750 & 500 & - & 500 & 500 & 500 \\ \hline 
-\textbf{7} & 1000 & 499500 & 1000 & - & 1000 & 1000 & 1000 \\ \hline 
-\textbf{8} & 2000 & 1999000 & 2000 & - & 2000 & 2000 & 2000 \\ \hline 
-\textbf{9} & 10 & 21 & 2 & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{10} & 50 & 504 & 2 & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{11} & 100 & 1771 & 2 & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{12} & 500 & 35275 & 2 & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{13} & 1000 & 164151 & 2 & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{14} & 2000 & 557775 & 2 & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{15} & 10 & 9 & 2 & 2 & 3 & 2 & 3 \\ \hline 
-\textbf{16} & 50 & 49 & 2 & 2 & 3 & 2 & 3 \\ \hline 
-\textbf{17} & 100 & 99 & 2 & 2 & 4 & 2 & 4 \\ \hline 
-\textbf{18} & 500 & 499 & 2 & - & 4 & 2 & 4 \\ \hline 
-\textbf{19} & 1000 & 999 & 2 & - & 4 & 2 & 4 \\ \hline 
-\textbf{20} & 2000 & 1999 & 2 & - & 4 & 2 & 4 \\ \hline 
-\textbf{21} & 10 & 20 & - & 3 & 3 & 3 & 3 \\ \hline 
-\textbf{22} & 50 & 100 & - & 4 & 4 & 3 & 4 \\ \hline 
-\textbf{23} & 100 & 200 & - & - & 4 & 4 & 4 \\ \hline 
-\textbf{24} & 500 & 1000 & - & - & 4 & 3 & 4 \\ \hline 
-\textbf{25} & 1000 & 2000 & - & - & 4 & 4 & 4 \\ \hline 
-\textbf{26} & 2000 & 4000 & - & - & 4 & 4 & 4 \\ \hline 
-\textbf{27} & 1383 & 10 & - & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{28} & 886 & 100 & - & 2 & 2 & 2 & 2 \\ \hline 
-\textbf{29} & 777 & 500 & - & 4 & 4 & 3 & 4 \\ \hline 
-\textbf{30} & 915 & 1000 & - & - & 4 & 3 & 4 \\ \hline 
-\textbf{31} & 1793 & 10000 & - & - & 8 & 6 & 8 \\ \hline 
-\textbf{32} & 1386 & 100000 & - & - & 40 & 35 & 40 \\ \hline 
-\textbf{33} & 1421 & 500000 & - & - & 167 & 153 & 167 \\ \hline 
-\textbf{34} & 1763 & 1000000 & - & - & 276 & 258 & 276 \\ \hline 
 \end{tabular}
@@
     \href{https://dergipark.org.tr/en/download/article-file/254140}{A Performance Comparison of Graph Coloring Algorithms}
-    % \bibitem{}
-    % \href{https://github.com/brrcrites/graph-coloring}{Graph coloring C++ library}
     \bibitem{}

### 2024-04-30 homework1/a_star.py

@@ -23,7 +23,6 @@ def astar(input_file: str, statistics_file: str, output_file: str) -> None:
         next_states = current.get_next_states()
         with open(statistics_file, 'a', encoding='utf-8') as file:
             file.write(f"Current state: {current.students_per_subject}\n")
-            # file.write(f"Next states: {[neighbour.students_per_subject for neighbour in next_states]}\n")
         for neighbour in next_states:
             new_cost = neighbour.g()
             students = str(neighbour.students_per_subject)

### 2024-04-30 homework1/state.py

@@ -21,7 +21,7 @@ class State:
         schedule: dict[str, dict[str, dict[str, tuple[str, str]]]] | None = None,
         count_techer_slots: dict[str, int] = None,
         students_per_subject: dict[str, int] = None,
-        cost_last_class: int = None
+        trade_off: int = None
     ) -> None:
         self.file_name = file_name
         self.yaml_dict = utils.read_yaml_file_for_hc(file_name) if yaml_dict is None else yaml_dict
@@ -32,7 +32,7 @@ class State:
             else copy.deepcopy(self.yaml_dict[utils.SUBJECTS])
         self.count_techer_slots = count_techer_slots if count_techer_slots is not None \
             else {teacher: 0 for teacher in self.yaml_dict[utils.TEACHERS]}
-        self.cost_last_class = cost_last_class if cost_last_class is not None else 0
+        self.trade_off = trade_off if trade_off is not None else 0
 
     @staticmethod
     def generate_empty_schedule(
@@ -119,7 +119,7 @@ class State:
 
     def g(self):
         ''' Returns the cost to reach the current state. '''
-        return 100 * self.no_constraints() + self.cost_last_class
+        return 100 * self.no_constraints() + self.trade_off
 
     def get_schedule(self) -> dict[str, dict[str, dict[str, tuple[str, str]]]]:
         ''' Returns the current schedule. '''
@@ -220,7 +220,7 @@ class State:
                             no_classrooms_for_subject = len([c for c in self.yaml_dict[utils.CLASSROOMS] \
                                 if subject in self.yaml_dict[utils.CLASSROOMS][c][utils.SUBJECTS]])
                             no_classrooms = len(self.yaml_dict[utils.CLASSROOMS])
-                            new_state.cost_last_class = 1 - (no_unassigned / no_empty_capacity) + \
+                            new_state.trade_off = 1 - (no_unassigned / no_empty_capacity) + \
                                 (no_classrooms_for_subject / no_classrooms)
                             new_state.schedule[day][slot][classroom] = (teacher, subject)
                             new_state.students_per_subject[subject] -= \

### 2024-05-13 homework1/paper.tex
last modification

@@
 % ------------------------------------------------------------------------------
-
+\pagebreak
 \section{Introduction}
@@
 
-For instance, in instances where an instructor's preferences are violated within 
+For instance, in instances where an instructor\'s preferences are violated within 
 a particular time slot, the algorithm should endeavor to substitute instructor with 
@@
 satisfies all constraints is found or when a specified number of iterations / 
-restarts have been completed. A pseudocode of the algorithm is presented in Algorithm
-\ref{alg1}.
+restarts have been completed. A pseudocode of the algorithm is presented in Algorithm 1.
 
@@
 \State $state = random.choice(from\ neighbors\ one\ of\ the\ neighbors\ with$ \\
-\hspace*{2em}$\text{minimum\ number\ of\ constraints\ unsatisfied)}$
+\hspace*{2em}$minimum\ number\ of\ constraints\ unsatisfied)$
 \EndWhile
@@
 
-The pseudocode of the algorithm is presented in Algorithm \ref{alg2}.
+The pseudocode of the algorithm is presented in Algorithm 2.
 
@@
 \subsection{Results}
-Pentru fiecare dintre algoritmii prezentați mai sus, am folosit \textit{hyperfine}
-pentru a evalua timpul de rulare al acestora și am salvat datele în câte un fișier
-MD. Excepție a făcut algoritmul Brute Force, căruia i-am calculat timpul de rulare
-doar pentru testele 1-2, 3, 9-14, 15-16, deoarece pentru celelalte teste timpul de
-rulare era prea mare.
+For the Hill-Climbing algorithm, the number of restarts was established at 100,
+while the maximum number of iterations for all tests was capped at 1000. In the 
+case of the A* algorithm, a weight of 100 was applied.
 
-Pentru fiecare set de date, am făcut un grafic cu timpul de rulare al fiecărui
-algoritm, iar rezultatele sunt prezentate în Figurile \ref{fig:empty_graphs} - \ref{fig:fixed_edge_number_graphs}.
+Each test involved the execution of both algorithms. The ensuing tables depict 
+the outcomes across various categories: the count of unsatisfied soft 
+constraints, the time taken to attain a solution, the total number of states 
+explored during the search, and the iterations needed to reach a solution. 
+Additionally, for the Hill Climbing algorithm, the number of restarts executed 
+until a solution was obtained is delineated.
 
-De asemenea, în tabelul \ref{tab:results} sunt prezentate rezultatele obținute
-pentru fiecare set de date în raport cu numărul cromatic teoretic al grafurilor (
-    acest număr a putut fi precalculat doar pentru testr 1-26
-).
-Pentru anumite teste care au fost prea costisitoare în timp pentru algoritmul Brute Force, am
-lăsat liber câmul respectiv.
-\pagebreak
+The outcomes for the Hill Climbing algorithm are categorized based on the 
+number of restarts.
+
+Moreover, the individual results for each test are presented in the following
+tables 1 and 2.
 
@@
 \centering
-\includegraphics[width=0.8\textwidth]{./graphics/fig1.png}
-\caption{Graficele timpului de rulare pentru grafuri goale}
+\includegraphics[width=0.8\textwidth]{./images/iterations.png}
+\caption{Number of iterations required to find the solution}
 \label{fig:empty_graphs}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{./graphics/fig2.png}
-\caption{Graficele timpului de rulare pentru grafuri complete}
+\includegraphics[width=0.8\textwidth]{./images/states.png}
+\caption{Total number of states explored during the search}
 \label{fig:complete_graphs}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{./graphics/fig3.png}
-\caption{Graficele timpului de rulare pentru grafuri bipartite complete}
+\includegraphics[width=0.8\textwidth]{./images/times.png}
+\caption{Time required to find the solution}
 \label{fig:bipartite_complete_graphs}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{./graphics/fig4.png}
-\caption{Graficele timpului de rulare pentru arbori binari}
+\includegraphics[width=0.8\textwidth]{./images/constraints.png}
+\caption{Number of soft constraints unsatisfied}
 \label{fig:binary_tree_graphs}
@@
 \centering
-\includegraphics[width=0.8\textwidth]{./graphics/fig5.png}
-\caption{Graficele timpului de rulare pentru grafuri planare}
+\includegraphics[width=0.8\textwidth]{./images/restarts.png}
+\caption{Number of restarts until a solution was found for Hill Climbing}
 \label{fig:planar_graphs}
-\end{figure}
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.8\textwidth]{./graphics/fig6.png}
-\caption{Graficele timpului de rulare pentru grafuri cu număr de muchii fixat}
-\label{fig:fixed_edge_number_graphs}
 \end{figure}
@@
 \centering
-\caption{Rezultatele obținute pentru fiecare set de date}
-\label{tab:results}
+\caption{Results obtained for the Hill Climbing algorithm}
+\label{tab:results_hc}
 \begin{tabular}{|c|c|c|c|c|c|c|c|}
 \hline
-\textbf{Nr. set} & \textbf{Nr. noduri} & \textbf{Nr. muchii} & \textbf{Nr. cromatic} & \textbf{Brute Force} & \textbf{Greedy} & \textbf{DSatur} & \textbf{RLF} \\ \hline
-\textbf{1} & 887 & 0 & 1 & 1 & 1 & 1 & 1 \\ \hline 
-\textbf{2} & 1384 & 0 & 1 & 1 & 1 & 1 & 1 \\ \hline
-\textbf{3} & 10 & 45 & 10 & 10 & 10 & 10 & 10 \\ \hline 
-\textbf{4} & 50 & 1225 & 50 & 50 & 50 & 50 & 50 \\ \hline 
-\textbf{5} & 100 & 4950 & 100 & - & 100 & 100 & 100 \\ \hline 
+\textbf{No. set} & \textbf{No. iterations} & \textbf{No. states} & \textbf{Time mm.ss.ms} & \textbf{No. unsatisfied constraints} & \textbf{No. restars} \\ \hline
+\textbf{1} & 106 & 9989 & 0.4.317 & 0 & 3  \\ \hline 
+\textbf{2} & 71 & 53875 & 0.34.053 & 0 & 1 \\ \hline
+\textbf{3} & 76 & 72621 & 0.77.753 & 0 & 1  \\ \hline 
+\textbf{4} & 5889 & 1228027 & 10.40.491 & 6 & 100 \\ \hline 
+\textbf{5} & 2221 & 1603927 & 22.47.289 & 5 & 20 \\ \hline 
 \end{tabular}
@@
 
-\subsection{Interpretarea rezultatelor}
-În raport cu timpul de execuție, avem următoarele observații:
-\begin{itemize}
-    \item Algoritmul Brute Force este cel mai lent, \textbf{nu} se folosește
-    în practică deloc.
-    \item Algoritmul Greedy pe toate tipurile de grafuri este cel mai rapid.
-    \item Algoritmul DSatur este mai rapid decât algoritmul Brute Force, dar
-    mai lent decât algoritmul Greedy.
-    \item Algoritmul RLF pe grafuri goale este mai lent și decât algoritmul
-    Brute Force, pe grafuri planare și arbori binari este mai lent decât DSatur, 
-    iar pe restul de grafuri este între Greedy și DSatur (aici au apărut cazuri
-    care în teorie nu sunt adevărate din cauza complexității agoritmului (vezi \ref{tabel_complexitate}).
-    Acest lucru se datorează faptului că nu s-au putut genera teste suficient de
-    mari și rulate de un număr suficient de ori pentru a se observa diferența dintre algoritmi.)
+\begin{table}[!]
+\centering
+\caption{Results obtained for the A Star algorithm}
+\label{tab:results_astar}
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\textbf{No. set} & \textbf{No. iterations} & \textbf{No. states} & \textbf{Time mm.ss.ms} & \textbf{No. unsatisfied constraints} \\ \hline
+\textbf{1} & 146 & 17073 & 0.8.537 & 0  \\ \hline
+\textbf{2} & 158 & 183159 & 2.12.937 & 0  \\ \hline
+\textbf{3} & 2883 & 2171276 & 52.49.856 & 0  \\ \hline
+\textbf{4} & 91 & 12368 & 0.8.172 & 0  \\ \hline
+\textbf{5} & 2542 & 1385784 & 27.15.878 & 0  \\ \hline
+\end{tabular}
+\end{table}
 
-\end{itemize}
+\subsection{Observations}
 
-În raport cu numărul cromatic, se poate observa cu ușurință faptul că algoritmul
-Brute Force este un algoritm exact, iar restul sunt aproximări (deși în teste 
-nu se observă o diferență între numărul cromatic teoretic și rezultatul de la algoritmul DSatur, dacă
-am fi avut un număr mai mare de teste cu numărul cromatic precalculat, ar fi apărut diferența).
-Dintre toți1 algoritmii euristici, algoritmul DSatur este cel mai exact.
+The Hill Climbing algorithm exhibits faster performance compared to the A* 
+algorithm primarily due to its local search nature. The Hill Climbing algorithm
+is more efficient in terms of the number of iterations required to reach a
+solution.
 
-\section{Concluzii}
-În cadrul acestei teme am observat 4 algoritmi diferiți de colorare a grafurilor.
+The A* algorithm, while slower, is more effective in terms of the number of
+unsatisfied constraints. The A* algorithm consistently yields solutions that
+satisfy all constraints, whereas the Hill Climbing algorithm occasionally
+encounters unsatisfied constraints.
 
-Se poate afirma că nu există o soluție complet generală pentru colorarea grafurilor.
-Fiecare din cele prezentate mai sus vin cu avantajele și dezavantajele lor.
+Recognizing the prolonged duration required to attain a solution for 
+orar\_bonus\_exact using the Hill Climbing algorithm, the number of 
+restarts was reduced to 20.
 
-Algoritmul Brute Force este cel mai lent, dar este exact, algoritmul Greedy este
-cel mai rapid, dar nu este exact, algoritmul DSatur deși este mai lent decât Greedy,
-este cel mai exact dintre algoritmi, iar algoritmul RLF este mai lent decât Brute Force și 
-DSatur, dar este aproximativ exact.
+The A Star algorithm is more suitable for scenarios similar to orar\_mic\_exact,
+orar\_mediu\_relaxat, and orar\_constrans\_incalcat, where the number of
+neighbors generated is smaller. In contrast, the Hill Climbing algorithm is
+more appropriate for scenarios similar to orar\_mare\_relaxat and orar\_bonus\_exact,
+where the number of neighbors generated is larger.
 
-În practică:
-\begin{itemize}
-    \item Dacă lucrați cu grafice bipartite sau ciclice, algoritmul Greedy nu este
-    suficient de eficient; pe de altă parte, algoritmul RLF este destul de exact și total
-    recomandat pentru aceste cazuri.
-    \item Dacă prioritar este timpul, se poate înclina spre utilizarea algoritmul Greedy.
-    \item Dacă prioritatea este un număr mai apropiat de cel cromatic, trebuie
-    să se utilizeze algoritmul DSatur.
+\section{Conclusions}
 
-\end{itemize}
+In summary, within our implementation of the Class Scheduling Problem, 
+the Hill Climbing algorithm outperforms the A* algorithm in terms of efficiency.
+A* achieves slower solution discovery, but with no unsatisfied constraints. 
+Conversely, Hill Climbing is faster, but may occasionally encounter unsatisfied
+constraints. The Hill Climbing algorithm is more suitable for scenarios where
+the soft constraints are less significant and the time required to find a
+solution is a critical factor. In contrast, the A* algorithm is more appropriate
+for scenarios where the soft constraints are more significant and the quality of
+the solution is paramount.
+
 \pagebreak
@@
 \begin{thebibliography}{8}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/A*_search_algorithm#Complexity}{A* search algorithm - Wikipedia}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Hill_climbing}{Hill Climbing - Wikipedia}
+    \bibitem{}
+    \href{https://curs.upb.ro/2023/course/view.php?id=13749}{Moodle - Artifical Intelligence Course}
+    \bibitem{}
+    \href{https://www.researchgate.net/profile/Kian-Pokorny/publication/262277120_Multiple_constraint_satisfaction_problems_using_the_A-star_A_search_algorithm_classroom_scheduling_with_preferences/links/5e57f530299bf1bdb8408b03/Multiple-constraint-satisfaction-problems-using-the-A-star-A-search-algorithm-classroom-scheduling-with-preferences.pdf}{Multiple constraint satisfaction problems using the A-star search algorithm}
+    \bibitem{}
+    \href{https://d1wqtxts1xzle7.cloudfront.net/68245993/14443-libre.pdf?1626958962=&response-content-disposition=inline%3B+filename%3DCourses_timetabling_based_on_hill_climbi.pdf&Expires=1714507660&Signature=JDd-gaDdE4E7s5P66Op535FyEuysvPoA5o4fcwak4NWT1sQNtVh50ooYa9eSrHu~aHR-Ulabenz910uDjozKpgHivCFkhipTT5bJJL9MensZTSib8TBVzGtBU9lycqCVwilt21Udi0DMgfhERDSgCcR4vO8hmlFlxOLg1exD1fNLE8kxt4S0H2Qfb3nQEHu3kLsicuJNk8wfF~eWpEDUlWsjW9J9IpxuLWP0OICQ7ZReWtx37dHnPN2PTVmFs815NMNi-jyLV0gWWBTAOL7Hn7A5VSg-l~s2e8RAoCjPAYQDRnnznIzOElRy2nTsgkW3mZDzlsntJ3lFur2KquwqNg__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA}{Courses timetabling based on hill climbing algorithm}
 
-    \bibitem{1}
-    J. Bondy and U. Murty, Graph Theory - Graduate Texts in Mathematic.
-    Springer, 2008.
-    \bibitem{2}
-    Z. \`{A}d\`{a}m Mann and A. Szajk\`{o}, “Average-case complexity of backtrack
-    search for coloring sparse random graphs,” Journal of Computer and
-    System Sciences, vol. 79, no. 8, pp. 1287–1301, 2013.
-    \bibitem{3}
-    “Backtrack: An o(1) expected time algorithm for the graph coloring
-    problem,” Information Processing Letters, vol. 18, no. 3, pp. 119–121,
-    1984.
-    \bibitem{4}
-    N. Barnier and P. Brisset, “Graph coloring for air traffic flow management,” 
-    Annals of Operations Research, vol. 130, 03 2002.
-    \bibitem{5}
-    “The application of a graph coloring method to an examination scheduling problem,” 
-    Institute for Operations Research and the Management
-    Sciences (INFORMS), vol. 11, no. 5.
-    \bibitem{6}
-    \href{http://www.ijimt.org/papers/359-D0129.pdf}{University Exam Scheduling System Using Graph
-    Coloring Algorithm and RFID Technology}
-    \bibitem{7}
-    A. Murat and B. Nurdan, “A performance comparison of graph coloring algorithms,” 
-    International Conference on Advanced Technology
-    Sciences (ICAT’16), vol. 4, pp. 1–19, 12 2016
-    \bibitem{8}
-    Z. Mann, “Complexity of coloring random graphs: An experimental
-    study of the hardest region,” Journal of Experimental Algorithmics,
-    vol. 23, pp. 1–19, 03 2018
-    \bibitem{9}
-    \href{https://acs-aa-challenge.github.io/acs-aa-challenge/18-np-2-colouring/}{Enunțul problemei}
-    \bibitem{10}
-    M. Garey and D. Johnson, Computer and Intractability: A Guide to the
-    Theory of NP-Completeness, 01 1979.
-    \bibitem{11}
-    D. Porumbel, J.-K. Hao, and P. Kuntz, “An evolutionary approach with
-    diversity guarantee and well-informed grouping recombination for graph
-    coloring,” Computers Operations Research, vol. 37, pp. 1822–1832, 10
-    2010.
-    \bibitem{12}
-    L. Ouerfelli and H. Bouziri, “Greedy algorithms for dynamic graph coloring,” 2011 International Conference on Communications, Computing
-    and Control Applications, CCCA 2011, 03 2011.
-    \bibitem{13}
-    \`{A}. E. Eiben, J. K. Van Der Hauw, and J. I. van Hemert, “Graph coloring
-    with adaptive evolutionary algorithms,” Journal of Heuristics, vol. 4,
-    no. 1, pp. 25–46, 1998.
-    \bibitem{14}
-    D. Br\`{e}laz, “New methods to color the vertices of a graph,”  Commun.
-    ACM, vol. 22, pp. 251–256, 04 1979.
-    \bibitem{15}
-    R. Janczewski, K. Manuszewski, and K. Piwakowski, “The smallest
-    hard-to-color graph for algorithm dsatur,” Discrete Mathematics, vol.
-    236, pp. 151–165, 06 2001.
-    \bibitem{16}
-    \href{https://github.com/ifsmirnov/jngen}{GitHub - Jngen: preparing tests made simpler}
-    \bibitem{}
-    \href{https://en.wikipedia.org/wiki/Graph_coloring}{Graph coloring - Wikipedia}
-    \bibitem{}
-    \href{https://citeseerx.ist.psu.edu/document?repid=rep1&type=pdf&doi=203a7b17267a28a06808bfb3b0b9571e32d15503}{A Comparison of Parallel Graph Coloring Algorithms}
-    \bibitem{}
-    \href{https://dergipark.org.tr/en/download/article-file/254140}{A Performance Comparison of Graph Coloring Algorithms}
-    \bibitem{}
-    \href{https://curs.upb.ro/2022/mod/folder/view.php?id=77105}{Moodle - Analiza algoritmilor}
-    % 
     \end{thebibliography}

### 2024-05-13 homework1/a_star.py

@@ -6,48 +6,66 @@ import utils
 from heapq import heappush, heappop
 from state import State
 
-def astar(input_file: str, statistics_file: str, output_file: str) -> None:
+Result = tuple[bool, int, int, State]
+
+def astar(input_file: str, statistics_file: str, output_file: str) -> Result:
     frontier = []
     initial = State(input_file)
     heappush(frontier, (initial.g() + initial.h(), initial))
 
     students_per_subject = str(initial.students_per_subject)
-    discovered = {students_per_subject: (None, 0)}
+    discovered = {students_per_subject: 0}
     final = None
+
+    total_iters, total_states = 0, 0
 
     while frontier:
         current = heappop(frontier)[1]
+        total_iters += 1
         if current.is_final():
             final = current
             break
         next_states = current.get_next_states()
+        total_states += len(next_states)
         with open(statistics_file, 'a', encoding='utf-8') as file:
             file.write(f"Current state: {current.students_per_subject}\n")
         for neighbour in next_states:
             new_cost = neighbour.g()
             students = str(neighbour.students_per_subject)
-            if students not in discovered or new_cost < discovered[students][1]:
-                discovered[students] = (current, new_cost)
-                heappush(frontier, (new_cost + neighbour.h(), neighbour))
+            if students not in discovered or new_cost < discovered[students]:
+                discovered[students] = new_cost
+                heappush(frontier, (new_cost, neighbour))
 
     if final is None:
         with open(statistics_file, 'a', encoding='utf-8') as file:
             file.write("No solution found.\n")
         with open(output_file, 'w', encoding='utf-8') as file:
             file.write("No solution found.\n")
-        return
+        return False, total_iters, total_states, None
 
     s = utils.pretty_print_timetable(final.schedule, input_file)
 
     with open(output_file, 'w', encoding='utf-8') as file:
         file.write(s)
 
+    print(s)
+
+    return final.check_constraints(), total_iters, total_states, final
+
 def run_test(
     name_of_input_file: str, name_of_statistics_file: str, name_of_output_file: str):
     ''' Function that runs the Hill Climbing algorithm in a separate thread. '''
     with open(name_of_statistics_file, 'w', encoding='utf-8') as file:
         file.write("Statistics for the A* algorithm\n\n")
-    astar(name_of_input_file, name_of_statistics_file, name_of_output_file)
+    is_final, iters, states, final = astar(name_of_input_file,
[…]

### 2024-05-13 homework1/hill_climbing.py

@@ -27,27 +27,26 @@ def stochastic_hill_climbing(
             break
 
         next_states = state.get_next_states()
-        best_states = [s for s in next_states if s.eval() < state.eval()]
         states += len(next_states)
 
-        if len(best_states) == 0:
+        if len(next_states) == 0:
             break
 
-        best_state = random.choice(best_states)
+        best_state = random.choice(next_states)
         state = best_state
 
     with open(statistics_file, 'a') as file:
         file.write(f"Final evaluation: {state.eval()} / {intial_eval}\n")
-    
-    return state.check_constraints(), iters, states, state
+
+    return state.is_final(), iters, states, state
 
 def random_restart_hill_climbing(
     initial_state: State,
     statistics_file: str,
     output_file: str,
     seed: int,
-    max_restarts: int = 100,
-    run_max_iters: int = 3000
+    max_restarts: int = 20,
+    run_max_iters: int = 1000
 ) -> Result:
     ''' Random Restart Hill Climbing algorithm with stochastic hill climbing at each restart. '''
     total_iters, total_states = 0, 0
@@ -65,19 +64,21 @@ def random_restart_hill_climbing(
         total_iters += iters
         total_states += states
 
-        if best_state is None or (state.eval() <= best_state.eval() and \
-            state.no_constraints() < best_state.no_constraints()):
-            best_state = state
-
         with open(statistics_file, 'a', encoding='utf-8') as file:
-            file.write(f"Final state vs best state evaluation: {state.eval()} / {best_state.eval()}\n")
-            file.write(f"Final state vs best state constraints: {state.no_constraints()} / {best_state.no_constraints()}\n")
+            if best_state is not None:
+                file.write(f"Final state vs best state evaluation: {state.eval()} / {best_state.eval()}\n")
+                file.write(f"Final state vs best state constraints: {state.no_constraints()} / {best_state.no_constraints()}\n")
 
         if is_final:
-            break
+            if best_state is None or (state.eval() == 0 and \
+                state.no_constraints() < best_state.no_constraints()):
+                best_state = state
+            if state.check_constraints():
+                break
 
     best_state.display(output_file)
-    return is_final, total_iters, total_states, state
+    best_state.display()
+    return is_final, total_iters, total_states, best_state
 
 def run_test(
     name_of_input_file: str, name_of_statistics_file: str, name_of_output_file: str):
@@ -86,6 +87,13 @@ def run_test(
[…]

### 2024-05-13 homework1/state.py

@@ -207,21 +207,20 @@ class State:
                                 continue
 
                             new_state = self.clone()
-                            # number of students unassigned to the subject
-                            no_unassigned = new_state.students_per_subject[subject]
-                            # number of empty seats in all the classrooms that can host the subject
-                            no_empty_capacity = 0
-                            for d in DAYS_OF_THE_WEEK[:self.size[1]]:
-                                for s in TIME_SLOTS[:self.size[0]]:
-                                    for c in self.yaml_dict[utils.CLASSROOMS]:
-                                        if subject in self.yaml_dict[utils.CLASSROOMS][c][utils.SUBJECTS] and \
-                                            self.schedule[d][s][c] is None:
-                                            no_empty_capacity += self.yaml_dict[utils.CLASSROOMS][c][utils.CAPACITY]
+                            # # number of students unassigned to the subject
+                            # no_unassigned = new_state.students_per_subject[subject]
+                            # # number of empty seats in all the classrooms that can host the subject
+                            # no_empty_capacity = 0
+                            # for d in DAYS_OF_THE_WEEK[:self.size[1]]:
+                            #     for s in TIME_SLOTS[:self.size[0]]:
+                            #         for c in self.yaml_dict[utils.CLASSROOMS]:
+                            #             if subject in self.yaml_dict[utils.CLASSROOMS][c][utils.SUBJECTS] and \
+                            #                 self.schedule[d][s][c] is None:
+                            #                 no_empty_capacity += self.yaml_dict[utils.CLASSROOMS][c][utils.CAPACITY]
                             no_classrooms_for_subject = len([c for c in self.yaml_dict[utils.CLASSROOMS] \
                                 if subject in self.yaml_dict[utils.CLASSROOMS][c][utils.SUBJECTS]])
                             no_classrooms = len(self.yaml_dict[utils.CLASSROOMS])
-                            new_state.trade_off = 1 - (no_unassigned / no_empty_capacity) + \
-                                (no_classrooms_for_subject / no_classrooms)
+                            new_state.trade_off = no_classrooms_for_subject / no_classrooms
                             new_state.schedule[day][slot][classroom] = (teacher, subject)
                             new_state.students_per_subject[subject] -= \
                                 self.yaml_dict[utils.CLASSROOMS][classroom][utils.CAPACITY]

### 2024-05-13 homework1/statistics.py

@@ -1,0 +1,92 @@
+''' Generates images for the statistics page. '''
+import matplotlib.pyplot as plt
+import numpy as np
+import os
+
+if __name__ == '__main__':
+	# create the directories if they do not exist
+	if not os.path.exists('images'):
+		os.mkdir('images')
+
+	# create the datasets
+
+	# number of iterations
+	y_astar_iter = [146, 158, 2883, 91, 2542]
+	y_hc_iter = [106, 71, 76, 5889, 2221]
+
+	# number of states
+	y_astar_states = [17073, 183159, 2171276, 12368, 1385784]
+	y_hc_states = [9989, 53875, 72621, 1228027, 1603927]
+
+	# times
+	y_hc_times = [0.4, 0.34, 0.77, 10.40, 22.47]
+	y_astar_times = [0.8, 2.12, 52.49, 0.8, 27.15]
+
+	# restarts
+	y_hc_restarts = [3, 1, 1, 100, 20]
+
+	# constraints
+	y_astar_constraints = [0, 0, 0, 0, 0]
+	y_hc_constraints = [0, 0, 0, 6, 5]
+
+	# create the plots
+
+	# number of iterations with Bar Graph
+	plt.figure()
+	x = np.arange(len(y_astar_iter))
+	plt.bar(x, y_astar_iter, width=0.4, label='A*')
+	plt.bar(x + 0.4, y_hc_iter, width=0.4, label='Hill Climbing')
+	plt.xlabel('Test number')
+	plt.ylabel('Number of iterations')
+	plt.title('Number of iterations')
+	plt.legend()
+	plt.savefig('images/iterations.png')
+
+	# number of states Bar Graph
+	plt.figure()
+	x = np.arange(len(y_astar_states))
+	plt.bar(x, y_astar_states, width=0.4, label='A*')
+	plt.bar(x + 0.4, y_hc_states, width=0.4, label='Hill Climbing')
+	plt.xlabel('Test number')
+	plt.ylabel('Number of states')
+	plt.title('Number of states')
+	plt.legend()
+	plt.savefig('images/states.png')
+
+	# times Bar Graph
+	plt.figure()
+	x = np.arange(len(y_astar_times))
+	plt.bar(x, y_astar_times, width=0.4, label='A*')
+	plt.bar(x + 0.4, y_hc_times, width=0.4, label='Hill Climbing')
+	plt.xlabel('Test number')
+	plt.ylabel('Time (min)')
+	plt.title('Time')
[…]

### 2024-05-22 homework2/paper/paper.tex
organize paper files

--- 2024-05-13 homework1/paper.tex
@@
 %
-\title{Class Scheduling Problem}
+\title{TODO}
 %
@@
 \begin{abstract}
-    With the rapid growth of student enrollment and the expansion of academic 
-    offerings in universities and colleges worldwide, the task of scheduling 
-    classes within existing timetables and facilities has become increasingly 
-    complex. Today, class scheduling requires consideration of multiple factors, 
-    including room availability, capacity, instructors' preferences, and more. This problem is considered to be 
-    NP-complete and has received some research during the past few years. 
-    Several formulations and algorithms have been proposed to solve scheduling
-    problems, most of which are based on local search techniques. In this paper, 
-    we propose to compare 2 different types of algorithms to solve the class
-    scheduling problem: the random restart Hill-Climbing algorithm and the A-Star algorithm.
+    TODO
 
-\keywords{NP-complete \and A Star \and Hill-Climbing \and searching algorithms \and class scheduling.}
+\keywords{TO \and DO.}
 \end{abstract}
@@
 \section{Introduction}
-\subsection{Problem Declaration}
-
-\quad The Class Scheduling Problem represents a critical challenge in the field 
-of educational operational research.  Universities contend with an ever-growing 
-course catalog, necessitating the efficient allocation of a diverse range of 
-classes to classrooms with varying capacities. This optimization problem seeks 
-to construct a course schedule that adheres to a comprehensive set of university 
-constraints while simultaneously maximizing the effective and efficient 
-utilization of existing facilities.
-
-The complexity of the Class Scheduling Problem presents a multifaceted challenge due 
-to a confluence of factors. First, there is significant variety in the 
-number of students enrolled in each course.  Similarly, classroom capacities 
-exhibit a wide range of variation. These factors are further compounded by the 
-existence of course-specific classroom constraints. Additionally, faculty 
-members often express preferences regarding the day of the week, time slot, 
-and even break schedule for their assigned courses. These factors, combined 
-with the need to satisfy a multitude of university-specific regulations, render 
-the task of assigning courses to classrooms a highly intricate endeavor.
-
-Merely ensuring a classroom can accommodate a course's enrolled students is 
-insufficient. Such an approach would lead to suboptimal space utilization, 
-potentially hindering the educational experience and fostering student 
-dissatisfaction. Consider a scenario with two courses: one with six students 
-and another with nineteen. Furthermore, imagine two classrooms exist, one 
-seating twenty and another seating fifty. While technically feasible to 
-schedule either course in either room, a more strategic allocation would be to 
-assign the larger course to the larger classroom.
-
-This example exemplifies the multifaceted nature of Class Scheduling Problem. 
-Researchers continue to explore sophisticated methods to address this intricate 
-problem, with the ultimate goal of ensuring the smooth operation and optimal 
-resource allocation within universities.
-
-\subsection{Problem Statement}
-\textbf{\textit{Statement:}} Receiving a set of courses, each with a specific
-number of students,a set of rooms with a specific capacity, and a list of 
-instructors with their list of preferences, the goal is to assign each course to
-a room and a time slot, according to a list of constraints.
-
-There are 2 types of constraints that need to be satisfied:
-\begin{itemize}
-    \item \textbf{Hard constraints:} These constraints must be satisfied in order
-    for the solution to be valid. If any of these constraints are not satisfied, 
-    the solution is invalid.
-    \item \textbf{Soft constraints:} These constraints are not mandatory, but 
-    they are taken into account when evaluating the quality of the solution.
-\end{itemize}
-
-\subsubsection*{Hard constraints}
-\begin{itemize}
-    \item Within a designated time slot and in a given room, only one subject may be taught by a single instructor.
-    \item During any given time slot, an instructor may teach only one subject, and this instruction must occur within a singular room.
-    \item An instructor may conduct classes in a maximum of 7 time slots per week.
-    \item Within a specified time slot, a room may accommodate a number of students equal to or less than its predetermined maximum capacity.
-    \item All students enrolled in a particular subject must have designated class hours allocated for that subject.
-    \item Instructors are restricted to teaching only the subjects in which they are specialized.
-    \item All rooms are designated for classes pertaining only to the subjects for which they have been assigned
-\end{itemize}
-
-\subsubsection*{Soft constraints}
-\begin{itemize}
-    \item An instructor may express preferences regarding specific days of the week for teaching or may wish to avoid teaching on certain days.
-    \item An instructor may have preferences regarding specific time slots during the day for teaching or may wish to avoid teaching during certain time slots.
-    \item An instructor may prefer not to have a break exceeding a certain number of hours between consecutive classes.
-\end{itemize}
-%
-
-\subsection{Methods of Approach}
-\subsubsection{Random Restart Hill Climbing}
-is a local search algorithm commonly applied in optimization 
-problems. It begins with an initial solution and iteratively explores 
-neighboring solutions, selecting the one that improves the objective function 
-the most. This process continues until a local optimum is reached, where no 
-better solution can be found in the immediate neighborhood. When the algorithm
-reaches a local optimum, it restarts the search from a random initial solution.
-The algorithm terminates when a specified number of iterations have been
-completed or when a solution that satisfies all constraints is found.
-
-\subsubsection{A Star}
-is an informed search algorithm widely used for pathfinding and graph 
-traversal tasks. It intelligently combines both actual path costs and heuristic 
-estimates to guide the search towards the goal efficiently. A* maintains a 
-priority queue of nodes to be explored and selects the most promising node 
-based on a combination of the cost incurred so far and the estimated cost to 
-reach the goal. This allows A* to efficiently find the optimal path while 
-intelligently pruning the search space, making it highly effective for solving 
-a wide range of optimization problems.
-
-\subsection{Evaluation}
-\quad Both algorithms will be evaluated based on the quality of the solutions they
-provide, the time required to find these solutions and the total number of
-states explored during the search. The quality of the solutions will be
-evaluated based on the number of constraints satisfied.  The time required to 
-find the solutions will be measured in milliseconds, and the total number of
-states explored will be counted during the search process.
-
-There will be 5 types of tests, describe in \textit{Tests Description} section.
 
@@
 \section{Algorithm Design}
-\subsection{State Representation}
 
-The state representation for the Class Scheduling Problem consists of a
-schedule that assigns each course to a room and a time slot. The state is
-represented as a class that contains the following fields:
-\begin{itemize}
-    \item \textbf{file\_name:} The name of the file from which the data was read.
-    \item \textbf{yaml\_dict:} A dictionary containing the data read from the file.
-    \item \textbf{size:} A tuplet (days, time\_slots) representing the size of the schedule.
-    \item \textbf{schedule:} A dictionary that has the following structure: \{(day: str) : \{(time\_slot: (int, int)) : \{(classroom : str) : ((instructor : str), (subject : str))\}\}\}.
-    \item \textbf{strudents\_per\_subject:} A dictionary that contains the number of students for each subject that needs to be scheduled.
-    \item \textbf{count\_techer\_slots:} A dictionary that contains the number of scheduled slots for each instructor.
-    \item \textbf{trade\_off:} A number that represents the trade-off between the number of constraints satisfied and the the choosen classroom at each step (used in A Star).
-\end{itemize}
-
-\subsubsection{Initial State}
-The initial state is generated as an empty schedule, which is initialized with 
-the parameters derived from the input file, including the number of days, time 
-slots, and classrooms. Additionally, the number of students per subject is 
-obtained from the input file, while the count of scheduled slots for each 
-instructor is set to zero.
-
-This initial state serves as the starting point for the search algorithms, 
-enabling them to iteratively allocate courses to rooms and time slots until 
-a valid schedule is achieved.
-
-An alternative method for generating the initial state involves randomly 
-assigning courses, instructors to rooms and time slots in a manner that 
-adheres to the hard constraints. A comparative analysis of these two approaches 
-will be conducted in the \textit{Initial State Selection} section.
-
-\subsubsection{Generating Neighbors}
-The neighbors of a state are generated by considering all possible combinations 
-of assigning a course to a room, a time slot and an instructor, while ensuring 
-adherence to the hard constraints.
-
-In the initial phase, all potential neighbors that adhere to both the hard and 
-soft constraints are generated. Subsequently, in the event that no neighbors 
-satisfying all soft constraints are found, a secondary phase ensues where only 
-neighbors satisfying the hard constraints are generated. This strategy reduces 
-the total number of neighbors generated, allowing the algorithm to prioritize 
-those that satisfy all constraints. Consequently, the algorithm minimizes time 
-wastage by avoiding the generation of neighbors that would not be utilized, 
-resulting in a reduced number of states generated.
-
-\subsubsection{Initial State Selection}
-As previously mentioned, the initial state can be generated in two methodologies: either 
-as an empty schedule or through the random assignment of courses, instructors, 
-and rooms to time slots. The former method exhibits a greater degree of 
-determinism, initializing the schedule with vacant slots, whereas the latter 
-introduces stochasticity into the initial state generation process.
-
-Upon experimentation with both approaches, it became evident that the random 
-initialization method could often yield solutions that fail to adhere to all 
-specified soft constraints. This underscores the importance of carefully 
-considering all potential subsequent states that may arise from the current 
-state when designing the random initialization method.
-
-For instance, in instances where an instructor\'s preferences are violated within 
-a particular time slot, the algorithm should endeavor to substitute instructor with 
-another whose preferences remain unviolated for that time slot. Moreover, the 
-algorithm should endeavor to substitute the time slot with another or substitute
-the room with anothers that have the same capacity in total. This approach
-was found more difficult to implement. When tring to implement the first two
-substitutes, the algorithm was not able to find a solution that satisfies all the
-soft constraints and it took quite a lot of time to find a partial solution due to
-a large number of constraints that should be checked while creating the neighbors.
-On the other hand, this approach will always find a partial solution that 
-satisfies all the hard constraints. Thus, this one is recommended in cases 
-when the soft constraints are not that important.
-
-The method that I used in 
-the end was the empty schedule initialization. It is acknowledged that without 
-the random initialization component, the Hill-Climbing algorithm may struggle 
-to assign all students to a room, thereby violating a hard constraint. However,
-with  the inclusion of the random restart, the algorithm will always find a 
-solution that satisfies all the hard constraints, the soft ones being satisfied 
-in most of the cases. This approach is recommended in cases when the soft 
-constraints are more important than in the previous case and it is easier to 
-generate the neighbors.
-
-\subsection{Random Restart Hill Climbing}
-
-As previously mentioned, the Random Restart Hill Climbing algorithm is the 
-most suitable for this approach. The algorithm is initialized with an empty
-schedule and generates neighbors that adhere to the hard constraints. The
-algorithm iteratively explores the neighborhood of the current state, selecting
-randomly one of the neighbors that satisfies the most constraints. This process continues until
-a local optimum is reached, at which point the algorithm restarts the search.
-The algorithm terminates when a solution that
-satisfies all constraints is found or when a specified number of iterations / 
-restarts have been completed. A pseudocode of the algorithm is presented in Algorithm 1.
-
-\begin{algorithm}
-\caption{Random Restart Hill Climbing Algorithm}
-\label{alg1}
-\begin{algorithmic}[1]
-\Procedure{hill\_climbing}{$max\_restarts$} \Return $[is\_final, total\_iters, total\_states, best\_state] $ 
-\State $total\_iters = 0$
-\State $total\_states = 0$
-\State $best\_state = None$
-
-\For{$index$ in $range(max\_restarts)$}
-\State $state = InitialState()$
-\State $is\_final, iters, states, state = stochastic\_hill\_climbing(state)$
-\State $total\_iters += iters$
-\State $total\_states += states$
-
-\If{$is\_final$}
-\State \Return $[is\_final, total\_iters, total\_states, state]$
-\EndIf
-
-\If{$state\ does\ not\ have\ hard\ constraints$}
-\If{$best\_state == None$ or $state\ has\ less\ soft\ constraints$
-\hspace*{6em}$unsatisfied\ than\ best\_state$}
-
-\State $best\_state = state$
-\EndIf
-\EndIf
-\EndFor
-
-\State \Return $[is\_final, total\_iters, total\_states, best\_state]$
-
-\algstore{bkbreak}
-\end{algorithmic}
-\end{algorithm}
-
-\addtocounter{algorithm}{-1}
-\begin{algorithm}
-\caption{Stochastic Hill Climbing}
-\begin{algorithmic}[1]
-\Procedure{stochastic\_hill\_climbing}{$state, max\__iters$} \Return $[is\_final, total\_iters, total\_states, best\_state] $ 
-\State $total\_iters = 0$
-\State $total\_states = 0$
-
-\While{$total\_iters < max\_iters$}
-\State $total\_iters += 1$
-
-\If{$state\ is\ final$}
-\Return $[True, total\_iters, total\_states, state]$
-\EndIf
-
-\State $neighbors = state.generate\_neighbors()$
-\State $total\_states += len(neighbors)$
-
-\If{$neighbors$ == $None$}
-\Return $[False, total\_iters, total\_states, state]$
-\EndIf
-
-\State $state = random.choice(from\ neighbors\ one\ of\ the\ neighbors\ with$ \\
-\hspace*{2em}$minimum\ number\ of\ constraints\ unsatisfied)$
-\EndWhile
-
-\State \Return $[False, total\_iters, total\_states, state]$
-
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{A Star}
-For the A Star algorithm, the state representation is the same as for the Random
-Restart Hill Climbing algorithm. The algorithm is initialized with an empty
-schedule and generates neighbors that adhere to the hard constraints. The 
-algorithm iteratively explores the neighborhood of the current state.
-
-The frontier represents a heap that contains the states that need to be explored.
-The diccovered is a dictionary that contains as keys the number of students that
-need to be scheduled for each subject and as values the cost of the state that
-brought us to this configuration.
-
-The function used in the A Star algorithm is:
-\begin{equation}
-    f(state) = g(state) + h(state)
-\end{equation}
-
-where
-
-\begin{equation}
-    h(state) = total\ number\ of\ students\ that\ are\ not\ assigned
-\end{equation}
-\begin{equation}
-    g(state) = number\ of\ constraints\ unsatisfied * weight + trade\_off
-\end{equation}
-\begin{equation}
-    trade\_off = \frac{number\ of\ classrooms(subject)}{total\ number\ of\ classrooms} 
-\end{equation}
-
-The heuristic function is admisible, because the return value is always less than
-or equal to the actual cost of the state to reach a final one, and is
-equal to 0 in the final states. On the other hand, the function h is not consistent,
-because a state from discovered can be added to the frontier with a smaller cost.
-
-The cost function is calculated as the number of constraints unsatisfied multiplied
-by a weight and the trade off. In cases where the number of constraints unsatisfied
-is equal, multiplying it by a weight will prioritize the states that have the trade
-off smaller. Moreover, when the number of constraints unsatisfied is different, the
-prioritization will be made based on the number of constraints unsatisfied, not on
-the trade off.
-
-The trade off is calculated as the number of classrooms that are assigned to a
-subject divided by the total number of classrooms. We want to prioritize adding
-a subject to a classroom that has fewer subjects assigned to it. For instance,
-if we have 2 subjects: A and B, 2 classrooms: 1 and 2, and classroom 1 is assigned
-to subject A and classroom 2 is assigned to both subjects, while tring to assign
-a subject to classroom 2 we will choose subject B (trade off = 0.5) instead of
-subject A (trade off = 1).
-
-The pseudocode of the algorithm is presented in Algorithm 2.
-
-\begin{algorithm} 
-\caption{A Star Algorithm}
-\label{alg2}
-\begin{algorithmic}[1]
-\Procedure{astar}{} \Return $[is\_final, total\_iters, total\_states, best\_state] $
-\State{$frontier = []$}
-\State{$discovered = \{\}$}
-\State{$state = InitialState()$}
-\State{$frontier.append((f(state), state))$}
-\State{$discovered[state] = 0$}
-\State{$total\_iters = 0$}
-\State{$total\_states = 1$}
-\While{$frontier$}
-\State{$current\_state = frontier.pop(1)$}
-\State{$total\_iters += 1$}
-\If{$current\_state\ is_final == 0$}
-\Return $[True, total\_iters, total\_states, current\_state]$
-\EndIf
-
-\State{$neighbors = current\_state.generate\_neighbors()$}
-\State{$total\_states += len(neighbors)$}
-\For{$neighbor$ in $neighbors$}
-\State{$new\_cost = g(neighbor) + h(neighbor)$}
-\State{students\_per\_subject = neighbor.students\_per\_subject}
-\If{$students\_per\_subject\ not\ in\ discovered$ or $new\_cost < discovered[students\_per\_subject]$}
-\State{$discovered[neighbor] = new\_cost$}
-\State{$frontier.append((new\_cost, neighbor))$}
-\EndIf
-\EndFor
-\EndWhile
-\State \Return $[False, total\_iters, total\_states, current\_state]$
-\EndProcedure
-\end{algorithmic}
-\end{algorithm}
-
-\subsection{Complexities}
-The complexity of the Random Restart Hill Climbing algorithm is $O(n)$, where
-$n$ is the total number of iterations. The complexity of the A Star algorithm is
-$O(b^d)$, where $b$ is the branching factor and $d$ is the depth of the solution.
-
-The complexity of the generate neighbors function is $O(d*t*c*i*s)$, where $d$ is
-the number of days, $t$ is the number of time slots, $c$ is the number of classrooms,
-$i$ is the number of instructors and $s$ is the number of subjects. Because the number
-of days does not exceed 7 (the worst case) and the number of time slots does not exceed 12, 
-the complexity becomes $O(c*i*s)$.
+% ------------------------------------------------------------------------------
 
 \section{Evaluation}
-\subsection{Tests Description}
 
-\subsubsection{orar\_mic\_exact:} Contains 3 courses, 2 rooms, and 13 instructors. The
-number of students for each course are 300, 330, and 330. The capacity of the rooms are
-20 and 30. The instructors have few preferences.
-\subsubsection{orar\_mediu\_relaxat:} Contains 4 courses, 4 rooms, and 18 instructors. The
-number of students for each course is 660, 660, 665 and 685. The capacity of the rooms is
-25, 25, 35 and 70. The instructors have few preferences.
-\subsubsection{orar\_mare\_relaxat:} Contains 8 courses, 6 rooms, and 37 instructors. The
-number of students for each course is 470, 475, 475, 495, 500, 530, 535 and 550. The capacity of the rooms is
-25, 30, 30, 35, 85 and 85. The instructors have few preferences.
-\subsubsection{orar\_constrans\_incalcat:} Contains 4 courses, 2 rooms, and 17 instructors. The
-number of students for each course is 720, 750, 780 and 810. The capacity of the rooms is
-15 and 90. The instructors have a lot of preferences.
-\subsubsection{orar\_bonus\_exact:} Contains 5 courses, 5 rooms, and 23 instructors. The
-number of students for each course is 500, 510, 515, 520 and 545. The capacity of the rooms is
-15, 15, 15, 15 and 50. The instructors have a lot of preferences, including breaks constraints.
-
-\subsection{Results}
-For the Hill-Climbing algorithm, the number of restarts was established at 100,
-while the maximum number of iterations for all tests was capped at 1000. In the 
-case of the A* algorithm, a weight of 100 was applied.
-
-Each test involved the execution of both algorithms. The ensuing tables depict 
-the outcomes across various categories: the count of unsatisfied soft 
-constraints, the time taken to attain a solution, the total number of states 
-explored during the search, and the iterations needed to reach a solution. 
-Additionally, for the Hill Climbing algorithm, the number of restarts executed 
-until a solution was obtained is delineated.
-
-The outcomes for the Hill Climbing algorithm are categorized based on the 
-number of restarts.
-
-Moreover, the individual results for each test are presented in the following
-tables 1 and 2.
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.8\textwidth]{./images/iterations.png}
-\caption{Number of iterations required to find the solution}
-\label{fig:empty_graphs}
-\end{figure}
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.8\textwidth]{./images/states.png}
-\caption{Total number of states explored during the search}
-\label{fig:complete_graphs}
-\end{figure}
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.8\textwidth]{./images/times.png}
-\caption{Time required to find the solution}
-\label{fig:bipartite_complete_graphs}
-\end{figure}
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.8\textwidth]{./images/constraints.png}
-\caption{Number of soft constraints unsatisfied}
-\label{fig:binary_tree_graphs}
-\end{figure}
-
-\begin{figure}[!]
-\centering
-\includegraphics[width=0.8\textwidth]{./images/restarts.png}
-\caption{Number of restarts until a solution was found for Hill Climbing}
-\label{fig:planar_graphs}
-\end{figure}
-
-\begin{table}[!]
-\centering
-\caption{Results obtained for the Hill Climbing algorithm}
-\label{tab:results_hc}
-\begin{tabular}{|c|c|c|c|c|c|c|c|}
-\hline
-\textbf{No. set} & \textbf{No. iterations} & \textbf{No. states} & \textbf{Time mm.ss.ms} & \textbf{No. unsatisfied constraints} & \textbf{No. restars} \\ \hline
-\textbf{1} & 106 & 9989 & 0.4.317 & 0 & 3  \\ \hline 
-\textbf{2} & 71 & 53875 & 0.34.053 & 0 & 1 \\ \hline
-\textbf{3} & 76 & 72621 & 0.77.753 & 0 & 1  \\ \hline 
-\textbf{4} & 5889 & 1228027 & 10.40.491 & 6 & 100 \\ \hline 
-\textbf{5} & 2221 & 1603927 & 22.47.289 & 5 & 20 \\ \hline 
-\end{tabular}
-\end{table}
-
-\begin{table}[!]
-\centering
-\caption{Results obtained for the A Star algorithm}
-\label{tab:results_astar}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-\textbf{No. set} & \textbf{No. iterations} & \textbf{No. states} & \textbf{Time mm.ss.ms} & \textbf{No. unsatisfied constraints} \\ \hline
-\textbf{1} & 146 & 17073 & 0.8.537 & 0  \\ \hline
-\textbf{2} & 158 & 183159 & 2.12.937 & 0  \\ \hline
-\textbf{3} & 2883 & 2171276 & 52.49.856 & 0  \\ \hline
-\textbf{4} & 91 & 12368 & 0.8.172 & 0  \\ \hline
-\textbf{5} & 2542 & 1385784 & 27.15.878 & 0  \\ \hline
-\end{tabular}
-\end{table}
-
-\subsection{Observations}
-
-The Hill Climbing algorithm exhibits faster performance compared to the A* 
-algorithm primarily due to its local search nature. The Hill Climbing algorithm
-is more efficient in terms of the number of iterations required to reach a
-solution.
-
-The A* algorithm, while slower, is more effective in terms of the number of
-unsatisfied constraints. The A* algorithm consistently yields solutions that
-satisfy all constraints, whereas the Hill Climbing algorithm occasionally
-encounters unsatisfied constraints.
-
-Recognizing the prolonged duration required to attain a solution for 
-orar\_bonus\_exact using the Hill Climbing algorithm, the number of 
-restarts was reduced to 20.
-
-The A Star algorithm is more suitable for scenarios similar to orar\_mic\_exact,
-orar\_mediu\_relaxat, and orar\_constrans\_incalcat, where the number of
-neighbors generated is smaller. In contrast, the Hill Climbing algorithm is
-more appropriate for scenarios similar to orar\_mare\_relaxat and orar\_bonus\_exact,
-where the number of neighbors generated is larger.
-
+% ------------------------------------------------------------------------------
 \section{Conclusions}
-
-In summary, within our implementation of the Class Scheduling Problem, 
-the Hill Climbing algorithm outperforms the A* algorithm in terms of efficiency.
-A* achieves slower solution discovery, but with no unsatisfied constraints. 
-Conversely, Hill Climbing is faster, but may occasionally encounter unsatisfied
-constraints. The Hill Climbing algorithm is more suitable for scenarios where
-the soft constraints are less significant and the time required to find a
-solution is a critical factor. In contrast, the A* algorithm is more appropriate
-for scenarios where the soft constraints are more significant and the quality of
-the solution is paramount.
 
@@
 
-%
 % ---- Bibliography ----
-%
-% BibTeX users should specify bibliography style 'splncs04'.
-% References will then be sorted and formatted in the correct style.
-%
-% \bibliographystyle{splncs04}
-% \bibliography{mybibliography}
-%
 \begin{thebibliography}{8}

### 2024-05-22 homework2/paper/paper.tex
add introduction section in paper

@@
 %
-\title{TODO}
+\title{Exploring the Utility of Machine Learning Across Varied Data Formats}
 %
@@
 \begin{abstract}
-    TODO
+    This study investigates the applicability of machine learning techniques on 
+    diverse datasets. We explore the effectiveness of two algorithms, Linear 
+    Regression and Multi-Layered Perceptron (MLP), on predicting both health 
+    outcomes and financial well-being. Specifically, we utilize a stroke 
+    prediction dataset to assess the model\'s ability to identify individuals at 
+    risk of stroke. Additionally, we employ a salary prediction dataset to 
+    evaluate the model\'s capacity to classify individuals earning above a 
+    specific income threshold (e.g., \$50,000 per year). Through comparative 
+    analysis, this research aims to elucidate the strengths and limitations of 
+    each algorithm when applied to these contrasting data types, offering 
+    insights into their suitability for various prediction tasks. Furthermore, 
+    we present a framework for data analysis, outlining essential steps for data
+    cleaning, exploration, and preparation, which can be applied to enhance the 
+    effectiveness of machine learning models across diverse datasets.
 
-\keywords{TO \and DO.}
+\keywords{\textit{Machine Learning \and Heterogeneous Data \and Comparative Analysis \and
+Prediction Modeling \and Data Analysis Techniques \and Stroke Prediction \and
+Salary Prediction \and Linear Regression \and Multi-Layered Perceptron \and 
+DataPreprocessing} }
+
 \end{abstract}
+% ------------------------------------------------------------------------------
+\section{Introduction}
+
+\subsection{Motivation: The Power and Nuance of Machine Learning Data}
+Machine learning (ML) has become a cornerstone of progress in numerous 
+disciplines. Its ability to extract valuable insights from vast and complex 
+datasets has fueled breakthroughs in healthcare, finance, and social sciences. 
+However, the effectiveness of ML models is not a one-size-fits-all proposition. 
+Different data types possess unique characteristics, and understanding these 
+nuances is essential for selecting the most appropriate ML algorithms. Data can 
+be structured (organized in tables) or unstructured (text, images), numerical 
+or categorical, and may exhibit linear or non-linear relationships between 
+features.  Choosing the right algorithm depends heavily on these factors. This 
+research delves into this crucial aspect of ML application by exploring the 
+performance of two distinct algorithms on contrasting datasets.
+
+\subsection{Research Focus: Delving into Stroke Prediction and Salary Prediction}
+This study focuses on the application of ML techniques to two contrasting 
+datasets: stroke prediction and salary prediction. Stroke, a leading cause of 
+disability and death globally, poses a significant public health burden. Stroke 
+prediction models aim to identify individuals at high risk of experiencing a 
+stroke, allowing for preventive measures and early intervention. These models 
+typically analyze factors such as age, blood pressure, cholesterol levels, and 
+smoking history.
+
+Conversely, salary prediction models attempt to classify individuals based on 
+income thresholds. This information can offer valuable insights into economic 
+trends, such as income inequality, and inform policy decisions. Salary 
+prediction models might analyze factors like education level, work experience, 
+and industry sector. By investigating these two distinct datasets, this research 
+aims to gain a broader understanding of how ML algorithms perform on different 
+data types with varying underlying structures and complexities.
+
+\subsection{Methodology: Unveiling the Algorithms - Linear Regression and Multi-Layered Perceptron}
+To investigate the performance on these contrasting datasets, this study employs 
+two prominent ML algorithms: Linear Regression and Multi-Layered Perceptron (MLP).
+
+Linear Regression is a well\-established technique known for its interpretability
+and efficiency in uncovering linear relationships between features (data points) 
+and target variables (what we want to predict). This makes it a valuable tool 
+for understanding the underlying factors influencing a particular outcome, such 
+as the relationship between blood pressure and stroke risk. However, its strength 
+lies in capturing linear relationships. If the underlying relationships in the 
+data are more complex and non-linear, then Linear Regression might not be as 
+effective.
+
+On the other hand, Multi-Layered Perceptrons (MLPs) are a type of artificial 
+neural network capable of learning complex, non-linear patterns within data. 
+Unlike Linear Regression, MLPs are not limited by linearity and can potentially 
+capture more intricate relationships between features and target variables. This 
+capability makes them particularly suitable for datasets with complex underlying 
+structures, such as the factors influencing an individual's salary, which might 
+involve a combination of education, experience, industry, and other factors 
+interacting in non-linear ways.
+
+\subsection{Research Objectives: Evaluating Algorithms, Unveiling Strengths and Weaknesses}
+By comparatively analyzing the performance of Linear Regression and MLP on 
+stroke and salary prediction tasks, this research seeks to achieve several key 
+objectives:
+
+\subsubsection{Evaluate the Suitability of Algorithms for Diverse Data Types:}
+This involves assessing the effectiveness of each algorithm in capturing the 
+underlying relationships within the stroke prediction and salary prediction 
+datasets. We will determine which algorithm performs better on each dataset, 
+offering insights into their suitability for different data types, such as 
+linear datasets (blood pressure and stroke risk) versus potentially non-linear 
+datasets (factors influencing salary).
+
+\subsubsection{Gain Insights into Algorithmic Strengths and Weaknesses:}
+By analyzing the comparative performance, we aim to highlight the scenarios where
+each algorithm excels and identify areas where one might outperform the other. 
+This will provide valuable guidance for researchers and practitioners in 
+selecting the most appropriate algorithm for their specific prediction tasks. 
+For instance, if interpretability is crucial (e.g., understanding the factors 
+influencing stroke risk), Linear Regression might be preferred. If the data is 
+likely to have complex, non-linear relationships, then an MLP might be a better 
+choice.
+
+\subsubsection{Demonstrate Best Practices for Data Analysis in ML Applications:}
+Effective data analysis is crucial for building robust ML models. This research 
+will showcase essential steps for data cleaning, exploration, and preparation, 
+emphasizing their importance in enhancing model performance across diverse 
+datasets. These steps may include handling missing values, identifying outliers, 
+and feature engineering (creating new features from existing data) to improve 
+the model's ability to learn from the data.
+
+\subsection{Expected Contribution: Advancing the Application of ML on Heterogeneous Data}
+Through this exploration, the research aims to contribute valuable knowledge to 
+the field of machine learning, particularly the application of ML on 
+heterogeneous datasets.  The findings can guide researchers and practitioners 
+in selecting appropriate algorithms for their specific prediction tasks and 
+data types.  Furthermore, by demonstrating best practices for data analysis, 
+this research can contribute to the development of more robust and reliable ML 
+models across diverse application domains. This can lead to advancements in 
+areas like healthcare (improved stroke prediction for preventive measures) and 
+economics (better understanding of factors influencing income inequality).  
+Ultimately, the research aims to contribute to the responsible and effective use 
+of ML for tackling complex problems across various fields.
+% ------------------------------------------------------------------------------
+\section{Exploratory Data Analysis}
 
 % ------------------------------------------------------------------------------
-\pagebreak
-\section{Introduction}
+\section{Data Preprocessing}
 
 % ------------------------------------------------------------------------------
-\section{Algorithm Design}
+\section{Algorithms Designs}
 

### 2024-05-22 homework2/paper/paper.tex
clasification of attributes

@@
 \usepackage{longtable}
+\usepackage{float}
+\usepackage{array}
 
@@
 
+\subsection{Datasets attributes description}
+The initial and crucial step in developing any machine learning algorithm 
+involves a thorough understanding of the data it will be trained on. This 
+understanding is achieved through a comprehensive analysis of the datasets' 
+characteristics. In this vein, the following sub sections will delve into the 
+specific attributes of the two datasets employed in this study: stroke 
+prediction and salary prediction.
+
+A detailed description of each salary prediction attribute is provided in Table
+\ref{tab:salary_attributes_description} and of each stroke  prediction attribute
+in Table \ref{tab:stroke_attributes_description}.
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}p{3cm}||>{\centering\arraybackslash}p{2cm}|>{\centering\arraybackslash}p{6cm}| }
+        \hline
+        \multicolumn{3}{|c|}{List of all attributes in the Salary Prediction dataset} \\
+        \hline
+        Attribute name & Type & Details \\
+        \hline
+        fnl & numeric & Socio-economic characteristic of the population from which the individual comes \\
+        \hline
+        hpw & numeric & Number of work hours per week \\
+        \hline
+        relation & categorical & The type of relationship in which the individual is involved \\
+        \hline
+        gain & numeric & Capital gain \\
+        \hline
+        country & categorical & Country of origin \\
+        \hline
+        job & categorical & The individual's job \\
+        \hline
+        edu\_int & numeric & Number of years of study \\
+        \hline
+        years & numeric & Age of the individual \\
+        \hline
+        loss & numeric & Loss of capital \\
+        \hline
+        work\_type & categorical & The job's type \\
+        \hline
+        partner & categorical & The type of partner the individual has \\
+        \hline
+        edu & categorical & The individual's type of education \\
+        \hline
+        gender & categorical & Individual's gender \\
+        \hline
+        race & categorical & Individual's race \\
+        \hline
+        prod & numeric & Capital production \\
+        \hline
+        gtype & categorical & Type of employment contract \\
+        \hline
+        \caption{Salary Prediction Attributes}
+        \label{tab:salary_attributes_description}
+   \end{longtable}
+\end{center}
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{3.5cm}||>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{3.5cm}||>{\centering\arraybackslash}m{2.5cm}| }
+        \hline
+        \multicolumn{4}{|c|}{List of all attributes in the Stroke Prediction dataset} \\
+        \hline
+        Attribute name & Type & Details & Possible values \\
+        \hline
+        mean\_blood\_sugar\_ level & numeric & The average value of blood glucose throughout the duration observation of the subject & \\
+        \hline
+        cardiovascular\_issues & categorical & Whether or not the subject has a medical history cardiovascular & 0, 1 \\
+        \hline
+        job\_category & categorical & The field in which the person works & child, entrepreneurial, N\_work\_history, private\_sector, public\_sector \\
+        \hline
+        body\_mass\_indicator & numeric & Body mass index, which indicates if the person is underweight, within limits normal, overweight or obese & \\
+        \hline
+        sex & categorical & The gender of the person & F, M\\
+        \hline
+        tobacco\_usage & categorical & Current or past smoker indicator &  ex-smoker, smoker, non-smoker\\
+        \hline
+        high\_blood\_pressure & categorical & Binary attribute indicating whether a person suffer from high blood pressure or not & 0, 1 \\
+        \hline
+        married & categorical & Binary attribute indicating whether the person a ever been married & Y, N \\
+        \hline
+        living\_area & categorical & The type of area where he lived most of his life & City, Countryside \\
+        \hline
+        years\_old & numeric & The person's age in years & \\
+        \hline
+        chaotic\_sleep & categorical & Binary attribute for a sleep program irregular & 0, 1 \\
+        \hline
+        analysis\_results & numeric & The results of medical analyzes of the person, which may include various measurements and indicators relevant to her health & \\
+        \hline
+        biological\_age\_index & numeric & An index that estimates the biological age of a person based on different factors such as lifestyle, health status, measured in an unknown unit & \\
+        \hline
+        cerebrovascular\_ accident & categorical & Binary indicator indicating whether the person a had a stroke or not & 0, 1 \\
+        \hline
+        \caption{Stroke Prediction Attributes}
+        \label{tab:stroke_attributes_description}
+   \end{longtable}
+\end{center}
+
+\subsection{Exploration of Attribute Types and Value Ranges}
+Prior to applyinga machine learning model to a dataset, a crucial step involves
+in identifying the types of attributes (features) present and their corresponding
+values ranges. This anaysis is essential for selecting appropriate algorithms 
+and ensuring optimalmodel performance. In the following pharagraphs we will
+describe three primary attribute types.
+\begin{itemize}
+    \item \textit{Continuous Numeric Attributes:}
+    These attributes possess numerical values that can theoretically take on any 
+    value within a specific range. Examples might include: age, weight, temperature etc.
+    \item \textit{Discrete Nominal Attributes:}
+    These attributes represent categorical data with distinct, non-ordered values. 
+    Examples include days of the week (Monday, Tuesday, etc.) or types of diseases 
+    (cancer, diabetes, etc.).
+    \item \textit{Ordinal Attributes:}
+    These attributes represent categorical data with values that exhibit an inherent 
+    order. However, the difference between consecutive values may not be interpretable 
+    in terms of a consistent unit.  Examples include customer satisfaction ratings 
+    (1-star, 2-star, etc.) or movie ratings (G, PG, PG-13, etc.). In ordinal 
+    attributes, the numerical value itself might not be as important as the relative 
+    order it represents.
+\end{itemize}
+
+Using the \textit{analysys\_attributes.py} script, we can identify the 
+Continuous Numeric Attributes and Discrete Nominal Attributes in a specific 
+dataset. The script will output statistics that can be showed in Tables 
+\ref{tab:continuous_numeric_attributes_salary} and 
+\ref{tab:continuous_numeric_attributes_stroke} for numeric attributes and
+Table \ref{tab:discrete_nominal_attributes_salary} and 
+\ref{tab:discrete_nominal_attributes_stroke} for nominal attributes.
+
+Moreover, the total number of items in the full dataset is 9999 for the Salary
+Prediction dataset and 5110 for the Stroke Prediction dataset.
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{0.95cm}||>{\centering\arraybackslash}m{1.75cm}|>{\centering\arraybackslash}m{1.75cm}|>{\centering\arraybackslash}m{1.75cm}|>{\centering\arraybackslash}m{1.75cm}|>{\centering\arraybackslash}m{1.75cm}|>{\centering\arraybackslash}m{1.75cm}|>{\centering\arraybackslash}m{1.75cm} |}
+        \hline
+        \multicolumn{8}{|c|}{List of all Continuous Numeric Attributes in the Salary Prediction dataset} \\
+        \hline
+         & fnl & hpw & gain & edu\_int & years & loss & prod \\
+        \hline\hline
+        count & 9.999000e +03 & 9199.00000 & 9999.00000 & 9999.00000 & 9999.00000 & 9999.00000 & 9999.00000 \\
+        \hline
+        mean &  1.903529e +05 & 40.416241 & 979.853385 & 14.262026 & 38.646865 & 84.111411 & 2014.9275 93 \\
+        \hline
+        std & 1.060709e +05 & 12.517356 & 7003.7953 82 & 24.770835 & 13.745101 & 3394.0354 84 & 14007.6044 96 \\
+        \hline
+        min & 1.921400e +04 & 1.000000 & 0.000000 & 1.000000 & 17.000000 & 0.000000 & -28.000000 \\
+        \hline
+        25\% & 1.182825e +05 & 40.000000 & 0.000000 & 9.000000 & 28.000000 & 0.000000 & 42.000000 \\
+        \hline
+        50\% & 1.784720e +05 & 40.000000 & 0.000000 & 10.000000 & 37.000000 & 0.000000 & 57.000000 \\
+        \hline
+        75\% & 2.373110e +05 & 45.000000 & 0.000000 & 13.000000 & 48.000000 & 0.000000 & 77.000000 \\
+        \hline
+        max & 1.455435e +06 & 99.00000 & 99999.0000 & 206.000000 & 90.000000 & 3770.00000 & 200125.000 \\
+        \hline
+        \caption{Continuous Numeric Attributes in Salary Prediction Dataset}
+        \label{tab:continuous_numeric_attributes_salary} \\
+   \end{longtable}
+\end{center}
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|}
+        \hline
+        \multicolumn{6}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
+        \hline
+        & mean\_blood\_ sugar\_level & cardiovascular\_ issues & body\_mass\_ indicator & high\_ blood\_pressure & years\_old \\
+        \hline\hline
+        count & 5110.000000 & 5110.000000 & 4909.000000 & 5110.000000 & 5110.000000 \\
+        \hline
+        mean & 106.147677 & 0.054012 & 28.893237 & 0.097456 & 46.568665 \\
+        \hline
+        std & 45.283560 & 0.226063 & 7.854067 & 0.296607 & 26.593912 \\
+        \hline
+        min & 55.120000 & 0.000000 & 10.300000 & 0.000000 & 0.080000 \\
+        \hline
+        25\% & 77.245000 & 0.000000 & 23.500000 & 0.000000 & 26.000000 \\
+        \hline
+        50\% & 91.885000 & 0.000000 & 28.100000 & 0.000000 & 47.000000 \\
+        \hline
+        75\% & 114.090000 & 0.000000 & 33.100000 & 0.000000 & 63.750000 \\
+        \hline
+        max & 271.740000 & 1.000000 & 97.600000 & 1.000000 & 134.000000 \\
+        \hline
+    \end{longtable}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{2.5cm}|>{\centering\arraybackslash}m{2.5cm}|>{\centering\arraybackslash}m{2.5cm}|>{\centering\arraybackslash}m{2.5cm}|}
+        \hline
+        \multicolumn{5}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
+        \hline
+        & chaotic\_sleep & analysis\_ results & biological\_ age\_index & cerebrovascular\_ accident \\
+        \hline\hline
+        count & 5110.000000 & 4599.000000 & 5110.000000 & 5110.000000 \\
+        \hline
+        mean & 0.054012 & 323.523446 & 134.784256 & 0.048728 \\
+        \hline
+        std & 0.226063 & 101.577442 & 50.399352 & 0.215320 \\
+        \hline
+        min & 0.000000 & 104.829714 & -15.109456 & 0.000000 \\
+        \hline
+        25\% & 0.000000 & 254.646209 & 96.710581 & 0.000000 \\
+        \hline
+        \hline
+        50\% & 0.000000 & 301.031628 & 136.374631 & 0.000000 \\
+        \hline
+        \hline
+        75\% & 0.000000 & 362.822769 & 172.507322 & 0.000000 \\
+        \hline
+        max & 1.000000 & 756.807975 & 266.986321 & 1.000000 \\
+        \hline
+        \caption{Continuous Numeric Attributes in Stroke Prediction Dataset}
+        \label{tab:continuous_numeric_attributes_stroke} \\
+   \end{longtable}
+\end{center}
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}|||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+        \hline
+        \multicolumn{3}{|c|}{List of all Discrete Nominal Attributes in the Salary Prediction dataset} \\
+        \hline
+        & Non-missing count & Unique values count \\
+        \hline\hline
+        \hline
+        country & 9999 & 41 \\
+        \hline
+        job & 9999 & 14 \\
+        \hline
+        work\_type & 9999 & 9 \\
+        \hline
+        partner & 9999 & 7 \\
+        \hline
+        edu & 9999 & 16 \\
+        \hline
+        gender & 9199 & 2 \\
+        \hline
+        race & 9999 & 5 \\
+        \hline
+        gtype & 9999 & 2 \\
+        \hline
+        money & 9999 & 2 \\
+        \hline
+        \caption{Discrete Nominal Attributes in Salary Prediction Dataset}
+        \label{tab:discrete_nominal_attributes_salary} \\
+    \end{longtable}
+\end{center}
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}|||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+        \hline
+        \multicolumn{3}{|c|}{List of all Discrete Nominal Attributes in the Stroke Prediction dataset} \\
+        \hline
+        & Non-missing count & Unique values count \\
+        \hline\hline
+        job\_category & 5110 & 5 \\
+        \hline
+        sex & 5110 & 2 \\
+        \hline
+        tobacco\_usage & 5110 & 4 \\
+        \hline
+        married & 4599 &2 \\
+        \hline
+        living\_area & 5110 & 2 \\
+        \hline
+
+        \caption{Discrete Nominal Attributes in Stroke Prediction Dataset}
+        \label{tab:discrete_nominal_attributes_stroke} \\
+    \end{longtable}
+\end{center}
+
+\subsection{Investigation of Class Distribution}
+TODO
+
+\subsection{Analysis of Feature Correlations}
+TODO
+
 % ------------------------------------------------------------------------------

### 2024-05-22 homework2/useful_scripts_for_analysis/analysis_attributes.py

@@ -1,0 +1,88 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import sys
+import os
+
+def analyze_csv(file_path, if_save=False):
+    # Read the CSV file into a DataFrame
+    df = pd.read_csv(file_path)
+
+    # Get descriptive statistics for numerical attributes
+    desc_stats = df.describe()
+    print("Descriptive Statistics:")
+    print(desc_stats)
+    
+    # Generate box plots for numerical attributes
+    numerical_cols = df.select_dtypes(include=['number']).columns
+    
+    if len(numerical_cols) > 0:
+        # Create a box plot for each numerical attribute
+        for col in numerical_cols:
+            # save the box plot as a file
+            if if_save:
+                plt.figure()
+                df.boxplot(column=col)
+                plt.title(f'Box plot for {col}')
+                plt.savefig(f'box_plot_{col}_{os.path.basename(file_path).split(".")[0]}.png', dpi=300)
+                plt.close()
+
+    else:
+        print("No numerical attributes found for box plots.")
+    
+     # Analyze categorical/ordinal attributes
+    categorical_cols = df.select_dtypes(include=['object', 'category']).columns
+    
+    if len(categorical_cols) > 0:
+        categorical_stats = {}
+        for col in categorical_cols:
+            non_missing_count = df[col].notna().sum()
+            unique_values_count = df[col].nunique()
+            categorical_stats[col] = [non_missing_count, unique_values_count]
+        
+        # Create a DataFrame for categorical/ordinal attribute statistics
+        categorical_stats_df = pd.DataFrame.from_dict(
+            categorical_stats, orient='index', columns=['Non-missing Count', 'Unique Values Count']
+        )
+        print("\nCategorical/Ordinal Attribute Statistics:")
+        print(categorical_stats_df)
+        
+        # Generate histograms for categorical/ordinal attributes
+        for col in categorical_cols:
+            if if_save:
+                plt.figure()
+                df[col].hist()
+                plt.title(f'Histogram for {col}')
+                # rotate the x-axis labels for better readability and zoom out them
+                if len(df[col].unique()) > 10:
+                    plt.xticks(rotation=45)
+                plt.tight_layout()
+                plt.savefig(f'histogram_{col}_{os.path.basename(file_path).split(".")[0]}.png', dpi=300)
+    else:
+        print("No categorical/ordinal attributes found for histograms.")
+
[…]

### 2024-05-23 homework2/paper/paper.tex
finish second chapter

@@
 
-Using the \textit{analysys\_attributes.py} script, we can identify the 
+Using the \textit{analysis\_attributes.py} script, we can identify the 
 Continuous Numeric Attributes and Discrete Nominal Attributes in a specific 
@@
 \begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|>{\centering\arraybackslash}m{2.25cm}|}
-        \hline
-        \multicolumn{6}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
-        \hline
-        & mean\_blood\_ sugar\_level & cardiovascular\_ issues & body\_mass\_ indicator & high\_ blood\_pressure & years\_old \\
-        \hline\hline
-        count & 5110.000000 & 5110.000000 & 4909.000000 & 5110.000000 & 5110.000000 \\
-        \hline
-        mean & 106.147677 & 0.054012 & 28.893237 & 0.097456 & 46.568665 \\
-        \hline
-        std & 45.283560 & 0.226063 & 7.854067 & 0.296607 & 26.593912 \\
-        \hline
-        min & 55.120000 & 0.000000 & 10.300000 & 0.000000 & 0.080000 \\
-        \hline
-        25\% & 77.245000 & 0.000000 & 23.500000 & 0.000000 & 26.000000 \\
-        \hline
-        50\% & 91.885000 & 0.000000 & 28.100000 & 0.000000 & 47.000000 \\
-        \hline
-        75\% & 114.090000 & 0.000000 & 33.100000 & 0.000000 & 63.750000 \\
-        \hline
-        max & 271.740000 & 1.000000 & 97.600000 & 1.000000 & 134.000000 \\
-        \hline
-    \end{longtable}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{2.5cm}|>{\centering\arraybackslash}m{2.5cm}|>{\centering\arraybackslash}m{2.5cm}|>{\centering\arraybackslash}m{2.5cm}|}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
         \hline
@@
         \hline
-        & chaotic\_sleep & analysis\_ results & biological\_ age\_index & cerebrovascular\_ accident \\
+        & mean\_blood\_ sugar\_level & cardiovascular\_ issues & body\_mass\_ indicator & high\_ blood\_pressure \\
         \hline\hline
-        count & 5110.000000 & 4599.000000 & 5110.000000 & 5110.000000 \\
+        count & 5110.000000 & 5110.000000 & 4909.000000 & 5110.000000 \\
         \hline
-        mean & 0.054012 & 323.523446 & 134.784256 & 0.048728 \\
+        mean & 106.147677 & 0.054012 & 28.893237 & 0.097456 \\
         \hline
-        std & 0.226063 & 101.577442 & 50.399352 & 0.215320 \\
+        std & 45.283560 & 0.226063 & 7.854067 & 0.296607 \\
         \hline
-        min & 0.000000 & 104.829714 & -15.109456 & 0.000000 \\
+        min & 55.120000 & 0.000000 & 10.300000 & 0.000000 \\
         \hline
-        25\% & 0.000000 & 254.646209 & 96.710581 & 0.000000 \\
+        25\% & 77.245000 & 0.000000 & 23.500000 & 0.000000 \\
         \hline
+        50\% & 91.885000 & 0.000000 & 28.100000 & 0.000000 \\
         \hline
-        50\% & 0.000000 & 301.031628 & 136.374631 & 0.000000 \\
+        75\% & 114.090000 & 0.000000 & 33.100000 & 0.000000 \\
         \hline
+        max & 271.740000 & 1.000000 & 97.600000 & 1.000000 \\
         \hline
-        75\% & 0.000000 & 362.822769 & 172.507322 & 0.000000 \\
+    \end{longtable}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
         \hline
-        max & 1.000000 & 756.807975 & 266.986321 & 1.000000 \\
+        \multicolumn{5}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
+        \hline
+        & years\_old & chaotic\_sleep & analysis\_results & biological\_age\_index \\
+        \hline\hline
+        count & 5110.000000 & 5110.000000 & 4599.000000 & 5110.000000 \\
+        \hline
+        mean & 46.568665 & 0.054012 & 323.523446 & 134.784256 \\
+        \hline
+        std & 26.593912 & 0.226063 & 101.577442 & 50.399352 \\
+        \hline
+        min & 0.080000 & 0.000000 & 104.829714 & -15.109456 \\
+        \hline
+        25\% & 26.000000 & 0.000000 & 254.646209 & 96.710581 \\
+        \hline
+        50\% & 47.000000 & 0.000000 & 301.031628 & 136.374631 \\
+        \hline
+        75\% & 63.750000 & 0.000000 & 362.822769 & 172.507322 \\
+        \hline
+        max & 134.000000 & 1.000000 & 756.807975 & 266.986321 \\
         \hline
@@
 
+An initial inspection of the data reveals that there are missing attributes in 
+both the Salary Prediction and Stroke Prediction datasets. The Salary Prediction 
+dataset is missing the \textit{'hpw'} attribute, while the Stroke Prediction dataset is
+missing two attributes: \textit{'body\_mass\_indicator'} and 
+\textit{'analysis\_results'}.
+
+To better understand the distribution of the continuous numeric attributes 
+within the datasets, boxplots have been generated for each attribute. These 
+visualizations are located in the \textit{'plots'} folder at the root of the 
+project directory. The name of each boxplot starts with \textit{'box\_plot\_'}.
+
+Boxplots are a standardized method for visually representing the distribution of
+data.  They provide insights into several key characteristics of the data, 
+including the median, quartiles, and outliers.
+
+In the Figure \ref{fig:boxplot_example_salary} we can see a boxplot for the 
+\textit{years} attribute in the Salary Prediction dataset. The box in the middle
+of the plot contains the middle 50\% of the data, and the line in the middle
+represents the median. The whiskers extend to the minimum and maximum values
+within 1.5 times the interquartile range (the difference between the first and
+third quartiles). Points outside this range are considered outliers.
+
+Also, in Figure \ref{fig:boxplot_example_stroke} we can see a boxplot for the
+\textit{body\_mass\_indicator} attribute in the Stroke Prediction dataset.
+As described above, the boxplot provides a visual representation of the data's
+distribution, highlighting key statistical measures such as the median, quartiles,
+and potential outliers. These information is also presented in the tables
+\ref{tab:continuous_numeric_attributes_salary} and
+\ref{tab:continuous_numeric_attributes_stroke}. One of the main insights that
+can be derived from the boxplot is the presence of outliers, which are data points
+that lie significantly outside the range of the rest of the data. Outliers can
+have a significant impact on the performance of machine learning models, and
+identifying and handling them appropriately is an essential step in the data
+preprocessing process.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/box_plot_years_SalaryPrediction_full.png}
+    \caption{Boxplot for the \textit{years} attribute in the Salary Prediction dataset}
+    \label{fig:boxplot_example_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/box_plot_body_mass_indicator_AVC_full.png}
+    \caption{Boxplot for the \textit{body\_mass\_indicator} attribute in the Stroke Prediction dataset}
+    \label{fig:boxplot_example_stroke}
+\end{figure}
+
 \begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}|||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
         \hline
@@
         \hline\hline
+        relation & 9999 & 6 \\
         \hline
@@
         \hline
-        money & 9999 & 2 \\
-        \hline
         \caption{Discrete Nominal Attributes in Salary Prediction Dataset}
@@
 \begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}|||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
         \hline
@@
 
+From the Discret Nominal Attributes tables we can see that each dataset contains
+only one attribute with missing values. The Salary Prediction dataset is missing
+the \textit{gender} attribute, while the Stroke Prediction dataset is missing the
+\textit{married} attribute. Also, the number of unique values for each attribute
+descibes the diversity of the data. For example, the \textit{country} attribute
+in the Salary Prediction dataset has 41 unique values, indicating that the data
+contains information from 41 different countries.
+
+In the historigrams for the discrete nominal attributes, we can see the
+distribution of the unique values for each attribute. These visualizations can
+provide insights into the frequency of each category within the dataset, which
+can be useful for understanding the data's composition and identifying potential
+imbalances or biases. The histograms for the discrete nominal attributes are
+located in the \textit{'plots'} folder at the root of the project directory.
+The name of each histogram starts with \textit{'histogram\_'}.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/histogram_work_type_SalaryPrediction_full.png}
+    \caption{Histogram for the \textit{work\_type} attribute in the Salary Prediction dataset}
+    \label{fig:histogram_example_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/histogram_tobacco_usage_AVC_full.png}
+    \caption{Histogram for the \textit{tobacco\_usage} attribute in the Stroke Prediction dataset}
+    \label{fig:histogram_example_stroke}
+\end{figure}
+
+In Figure \ref{fig:histogram_example_salary} we can see a histogram of the
+\textit{work\_type} attribute in the Salary Prediction dataset. The dominance of
+the 'Priv' category indicates a severe class imbalance. For classification tasks,
+the model might predict Priv most of the time since it's the majority class, 
+leading to a high overall accuracy but poor precision, recall, and F1 scores 
+for minority classes.
+
+Also, in Figure \ref{fig:histogram_example_stroke} we can see a histogram of the
+\textit{tobacco\_usage} attribute in the Stroke Prediction dataset. The histogram 
+shows that the majority of individuals are non-smokers, with a significant portion 
+having undefined tobacco usage status. This imbalance and the presence of missing 
+data need to be addressed appropriately.
+
 \subsection{Investigation of Class Distribution}
-TODO
+In machine learning, it is common practice to split a dataset into two distinct
+subsets: a training set and a test set. This division is crucial for ensuring 
+ robustness and generalizability of the models developed using the data.
+
+\begin{itemize}
+    \item \textbf{Training Set:} The primary purpose of the training set is to 
+    train the machine 
+    learning model. The model learns from patterns and relationships within the data 
+    to develop a predictive capability.
+    \item \textbf{Test Set:} The test set, unseen by the model during training, 
+    serves to evaluate the model's generalizability. By applying the trained 
+    model to the test set, we can assess its performance on new, unseen data. 
+    This helps prevent overfitting, where the model performs well on the 
+    training data but fails to generalize to real-world scenarios.
+\end{itemize}
+
+Looking at how data is distributed is key. Imbalanced data, where some classes 
+have far more examples than others, throws off classification tasks:
+high accuracy can hide poor performance on rare classes; models struggle to 
+learn patterns from underrepresented classes; inaccurate predictions, especially 
+for the minority class.
+
+By checking the distribution, we can address imbalance:
+\begin{itemize}
+    \item Balance the data: Oversample rare examples or undersample common ones.
+    \item Cost-sensitive learning: Penalize the model more for mistakes on rare classes.
+    \item Better metrics: Use precision, recall, and F1-score to get a clearer picture.
+\end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/distribution_edu_Salary_Prediction.png}
+    \caption{Distribution of the \textit{edu} attribute in the Salary Prediction dataset}
+    \label{fig:distribution_example_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/distribution_tobacco_usage_Stroke_Prediction.png}
+    \caption{Distribution of the \textit{tobacco\_usage} attribute in the Stroke Prediction dataset}
+    \label{fig:distribution_example_stroke}
+\end{figure}
+
+In the \textit{'plots'} folder at the root of the project directory we can find
+the distribution of each class in the datasets. The name of each plot starts
+with \textit{'distribution\_'}. Here is an example of the distribution of the
+\textit{edu} attribute in the Salary Prediction dataset in Figure
+\ref{fig:distribution_example_salary}. Also, the distribution of the
+\textit{tobacco\_usage} attribute in the Stroke Prediction dataset is shown in
+Figure \ref{fig:distribution_example_stroke}.
 
 \subsection{Analysis of Feature Correlations}
-TODO
+Feature correlation analysis is a critical step in understanding the relationships
+between different attributes in a dataset. By examining how attributes are related
+to each other, we can identify patterns, dependencies, and redundancies that can
+inform feature selection, model building, and interpretation.
+
+Correlation analysis typically involves calculating correlation coefficients
+between pairs of attributes. The correlation coefficient quantifies the strength
+and direction of the linear relationship between two variables. A correlation
+coefficient close to 1 indicates a strong positive relationship, while a value
+close to -1 indicates a strong negative relationship. A correlation coefficient
+near 0 suggests no linear relationship between the variables.
+
+In the \textit{'correlation\_analysis.py'} script, we calculate the correlation
+coefficients between all pairs of continuous numeric attributes in the datasets,
+generating a correlation matrix for each dataset. Moreover, we calculate the 
+Cramér's V coefficient for all pairs of discrete nominal attributes in the datasets,
+generating a Cramér's V matrix for each dataset to measure the association between
+categorical variables. In Figures \ref{fig:correlation_matrix_example_salary} and
+\ref{fig:correlation_matrix_example_stroke} we can see the correlation matrix for
+the Salary Prediction and Stroke Prediction datasets, respectively, for the continuous
+numeric attributes. In Figures \ref{fig:cramer_v_matrix_example_salary} and
+\ref{fig:cramer_v_matrix_example_stroke} we can see the Cramér's V matrix for
+the discrete nominal attributes.
+
+The correlation matrix and Cramér's V matrix provide valuable insights into the
+relationships between attributes in the datasets.By examining these matrices, we
+can see that Figure \ref{fig:correlation_matrix_example_salary} the \textit{prod}
+attribute is highly correlated with the \textit{gain} attribute, while the
+\textit{years} attribute is negatively correlated with the \textit{fnl} attribute.
+
+In Figure \ref{fig:correlation_matrix_example_stroke} we can see that the
+\textit{cardiovascular\_issues} attribute is highly correlated with the
+\textit{chaotic\_sleep} attribute, while the \textit{body\_mass\_indicator} 
+attribute is negatively correlated with the \textit{cerebrovascular\_accident}
+attribute.
+
+The Cramér's V matrix in Figures \ref{fig:cramer_v_matrix_example_salary} and
+\ref{fig:cramer_v_matrix_example_stroke} provides insights into the association
+between discrete nominal attributes. For example, in the Salary Prediction dataset,
+the \textit{gtype} attribute is strongly associated with the \textit{gender} attribute,
+while in the Stroke Prediction dataset, there are no strong associations between
+the discrete nominal attributes.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_SalaryPrediction_full.png}
+    \caption{Correlation matrix for the Salary Prediction dataset}
+    \label{fig:correlation_matrix_example_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_AVC_full.png}
+    \caption{Correlation matrix for the Stroke Prediction dataset}
+    \label{fig:correlation_matrix_example_stroke}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/cramer_v_matrix_SalaryPrediction_full.png}
+    \caption{Cramér's V matrix for the Salary Prediction dataset}
+    \label{fig:cramer_v_matrix_example_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/cramer_v_matrix_AVC_full.png}
+    \caption{Cramér's V matrix for the Stroke Prediction dataset}
+    \label{fig:cramer_v_matrix_example_stroke}
+\end{figure}
 
@@
     \bibitem{}
-    \href{https://en.wikipedia.org/wiki/A*_search_algorithm#Complexity}{A* search algorithm - Wikipedia}
+    \href{https://en.wikipedia.org/wiki/Correlation}{Wikipedia - Correlation}
     \bibitem{}
-    \href{https://en.wikipedia.org/wiki/Hill_climbing}{Hill Climbing - Wikipedia}
+    \href{https://thinkingneuron.com/how-to-measure-the-correlation-between-two-categorical-variables-in-python/}{Thinking Neuron - How to Measure the Correlation Between Two Categorical Variables in Python}
     \bibitem{}
-    \href{https://curs.upb.ro/2023/course/view.php?id=13749}{Moodle - Artifical Intelligence Course}
+    \href{https://www.stratascratch.com/blog/chi-square-test-in-python-a-technical-guide/}{StrataScratch - Chi-Square Test in Python: A Technical Guide}
     \bibitem{}
-    \href{https://www.researchgate.net/profile/Kian-Pokorny/publication/262277120_Multiple_constraint_satisfaction_problems_using_the_A-star_A_search_algorithm_classroom_scheduling_with_preferences/links/5e57f530299bf1bdb8408b03/Multiple-constraint-satisfaction-problems-using-the-A-star-A-search-algorithm-classroom-scheduling-with-preferences.pdf}{Multiple constraint satisfaction problems using the A-star search algorithm}
+    \href{https://en.wikipedia.org/wiki/Cram\%C3\%A9r\%27s\_V}{Wikipedia - Cramér's V}
     \bibitem{}
-    \href{https://d1wqtxts1xzle7.cloudfront.net/68245993/14443-libre.pdf?1626958962=&response-content-disposition=inline%3B+filename%3DCourses_timetabling_based_on_hill_climbi.pdf&Expires=1714507660&Signature=JDd-gaDdE4E7s5P66Op535FyEuysvPoA5o4fcwak4NWT1sQNtVh50ooYa9eSrHu~aHR-Ulabenz910uDjozKpgHivCFkhipTT5bJJL9MensZTSib8TBVzGtBU9lycqCVwilt21Udi0DMgfhERDSgCcR4vO8hmlFlxOLg1exD1fNLE8kxt4S0H2Qfb3nQEHu3kLsicuJNk8wfF~eWpEDUlWsjW9J9IpxuLWP0OICQ7ZReWtx37dHnPN2PTVmFs815NMNi-jyLV0gWWBTAOL7Hn7A5VSg-l~s2e8RAoCjPAYQDRnnznIzOElRy2nTsgkW3mZDzlsntJ3lFur2KquwqNg__&Key-Pair-Id=APKAJLOHF5GGSLRBV4ZA}{Courses timetabling based on hill climbing algorithm}
 

### 2024-05-23 homework2/plots/analysis_attributes.py

@@ -23,7 +23,7 @@ def analyze_csv(file_path, if_save=False):
                 plt.figure()
                 df.boxplot(column=col)
                 plt.title(f'Box plot for {col}')
-                plt.savefig(f'box_plot_{col}_{os.path.basename(file_path).split(".")[0]}.png', dpi=300)
+                plt.savefig(f'plots/box_plot_{col}_{os.path.basename(file_path).split(".")[0]}.png', dpi=300)
                 plt.close()
 
     else:
@@ -56,7 +56,7 @@ def analyze_csv(file_path, if_save=False):
                 if len(df[col].unique()) > 10:
                     plt.xticks(rotation=45)
                 plt.tight_layout()
-                plt.savefig(f'histogram_{col}_{os.path.basename(file_path).split(".")[0]}.png', dpi=300)
+                plt.savefig(f'plots/histogram_{col}_{os.path.basename(file_path).split(".")[0]}.png', dpi=300)
     else:
         print("No categorical/ordinal attributes found for histograms.")
 
@@ -73,16 +73,11 @@ def analyze_csv(file_path, if_save=False):
             
 
 if __name__ == "__main__":
-    # get the file path from the parameter passed to the script
-    if len(sys.argv) < 2:
-        print("Usage: python numerical_attributes.py <file_path1> <file_path2> ...")
-        sys.exit(1)
+    file = './tema2_SalaryPrediction/SalaryPrediction_full.csv'
+    analyze_csv(file, if_save=True)
+    print(f"\nAnalysis completed for {file}\n")
 
-    # analyze each file
-    for file_path in sys.argv[1:]:
-        if not os.path.exists(file_path):
-            print(f"File not found: {file_path}")
-            continue
-        analyze_csv(file_path, if_save=True)
-        print(f"\nAnalysis completed for {file_path}\n")
+    file = './tema2_AVC/AVC_full.csv'
+    analyze_csv(file, if_save=True)
+    print(f"\nAnalysis completed for {file}\n")
 

### 2024-05-23 homework2/plots/correlation_analysis.py

@@ -1,0 +1,84 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy.stats import chi2_contingency
+import numpy as np
+import sys
+import os
+
+def numerical_features(file_name):
+	file = pd.read_csv(file_name)
+	cols = file.select_dtypes(include=['number']).columns
+
+	# Create a correlation matrix
+	corr_matrix = file[cols].corr()
+
+	# Generate a heatmap for the correlation matrix
+	plt.figure()
+	plt.matshow(corr_matrix, cmap='coolwarm')
+	plt.colorbar()
+	plt.xticks(np.arange(len(cols)), cols, rotation=45)
+	plt.yticks(np.arange(len(cols)), cols)
+	plt.title(f'Correlation Matrix for {os.path.basename(file_name).split(".")[0]} dataset')
+	# make the plot bigger
+	fig = plt.gcf()
+	fig.set_size_inches(15, 10.5)
+	plt.savefig(f'plots/correlation_matrix_{os.path.basename(file_name).split(".")[0]}.png', dpi=300)
+	plt.close()
+
+def cramers_v(confusion_matrix):
+    chi2 = chi2_contingency(confusion_matrix)[0]
+    n = confusion_matrix.sum().sum()
+    phi2 = chi2 / n
+    r, k = confusion_matrix.shape
+    phi2corr = max(0, phi2 - ((k-1)*(r-1))/(n-1))
+    rcorr = r - ((r-1)**2)/(n-1)
+    kcorr = k - ((k-1)**2)/(n-1)
+    return np.sqrt(phi2corr / min((kcorr-1), (rcorr-1)))
+
+def cramers_v_matrix(df):
+    features = df.columns
+    matrix = pd.DataFrame(index=features, columns=features, dtype=float)
+    for feature_1 in features:
+        for feature_2 in features:
+            if feature_1 == feature_2:
+                matrix.loc[feature_1, feature_2] = 1.0
+            else:
+                confusion_matrix = pd.crosstab(df[feature_1], df[feature_2])
+                matrix.loc[feature_1, feature_2] = cramers_v(confusion_matrix)
+    return matrix
+
+def cathegorical_features(file_name):
+	file = pd.read_csv(file_name)
+	cols = file.select_dtypes(include=['object', 'category']).columns
+
+	corelation = cramers_v_matrix(file[cols])
+
[…]

### 2024-05-23 homework2/plots/distributions.py

@@ -1,0 +1,63 @@
+import pandas as pd
+import matplotlib.pyplot as plt
+import sys
+import os
+  
+
+if __name__ == "__main__":
+    # Salary Prediction datasets
+	full_set = pd.read_csv('./tema2_SalaryPrediction/SalaryPrediction_full.csv')
+	train_set = pd.read_csv('./tema2_SalaryPrediction/SalaryPrediction_train.csv')
+	test_set = pd.read_csv('./tema2_SalaryPrediction/SalaryPrediction_test.csv')
+
+	full_set_features = full_set.select_dtypes(include=['object', 'category']).columns
+	train_set_features = train_set.select_dtypes(include=['object', 'category']).columns
+	test_set_features = test_set.select_dtypes(include=['object', 'category']).columns
+
+	for feature in full_set_features:
+		plt.figure()
+		# convert nan values to string
+		full_set[feature] = full_set[feature].astype(str)
+		train_set[feature] = train_set[feature].astype(str)
+		test_set[feature] = test_set[feature].astype(str)
+
+		plt.hist(full_set[feature], bins=50, alpha=0.5, label='Full Dataset')
+		plt.hist(train_set[feature], bins=50, alpha=0.5, label='Train Dataset')
+		plt.hist(test_set[feature], bins=50, alpha=0.5, label='Test Dataset')
+
+		if len(full_set[feature].unique()) > 10:
+			plt.xticks(rotation=45)
+		
+		plt.title(f'Distribution of Salary for {feature}')
+		plt.legend()
+		plt.savefig(f'plots/distribution_{feature}_Salary_Prediction.png', dpi=300)
+		plt.close()
+	
+	# Stroke Prediction datasets
+	full_set = pd.read_csv('./tema2_AVC/AVC_full.csv')
+	train_set = pd.read_csv('./tema2_AVC/AVC_train.csv')
+	test_set = pd.read_csv('./tema2_AVC/AVC_test.csv')
+
+	full_set_features = full_set.select_dtypes(include=['object', 'category']).columns
+	train_set_features = train_set.select_dtypes(include=['object', 'category']).columns
+	test_set_features = test_set.select_dtypes(include=['object', 'category']).columns
+
+	for feature in full_set_features:
+		plt.figure()
+		# convert nan values to string
+		full_set[feature] = full_set[feature].astype(str)
+		train_set[feature] = train_set[feature].astype(str)
[…]

### 2024-05-23 homework2/paper/paper.tex
modify second chapter

@@
     outcomes and financial well-being. Specifically, we utilize a stroke 
-    prediction dataset to assess the model\'s ability to identify individuals at 
+    prediction dataset to assess the model's ability to identify individuals at 
     risk of stroke. Additionally, we employ a salary prediction dataset to 
-    evaluate the model\'s capacity to classify individuals earning above a 
+    evaluate the model's capacity to classify individuals earning above a 
     specific income threshold (e.g., \$50,000 per year). Through comparative 
@@
 \begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{3.5cm}||>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{3.5cm}||>{\centering\arraybackslash}m{2.5cm}| }
+    \begin{longtable}{ |>{\centering\arraybackslash}m{3.5cm}||>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{3.5cm}|>{\centering\arraybackslash}m{2.5cm}| }
         \hline
@@
 \subsection{Exploration of Attribute Types and Value Ranges}
-Prior to applyinga machine learning model to a dataset, a crucial step involves
+Prior to applying a machine learning model to a dataset, a crucial step involves
 in identifying the types of attributes (features) present and their corresponding
 values ranges. This anaysis is essential for selecting appropriate algorithms 
-and ensuring optimalmodel performance. In the following pharagraphs we will
+and ensuring optimal model performance. In the following pharagraphs we will
 describe three primary attribute types.
@@
 Using the \textit{analysis\_attributes.py} script, we can identify the 
-Continuous Numeric Attributes and Discrete Nominal Attributes in a specific 
-dataset. The script will output statistics that can be showed in Tables 
+Continuous Numeric Attributes and Discrete Nominal Attributes in the  
+datasets. The script will output statistics that can be showed in Tables 
 \ref{tab:continuous_numeric_attributes_salary} and 
@@
 Table \ref{tab:discrete_nominal_attributes_salary} and 
-\ref{tab:discrete_nominal_attributes_stroke} for nominal attributes.
+\ref{tab:discrete_nominal_attributes_stroke} for discrete attributes.
 
@@
 An initial inspection of the data reveals that there are missing attributes in 
-both the Salary Prediction and Stroke Prediction datasets. The Salary Prediction 
-dataset is missing the \textit{'hpw'} attribute, while the Stroke Prediction dataset is
-missing two attributes: \textit{'body\_mass\_indicator'} and 
+both the Salary Prediction and Stroke Prediction datasets. In the Salary Prediction 
+datase the \textit{'hpw'} attribute is missing, while int the Stroke Prediction dataset
+two attributes are missing: \textit{'body\_mass\_indicator'} and 
 \textit{'analysis\_results'}.
@@
 distribution, highlighting key statistical measures such as the median, quartiles,
-and potential outliers. These information is also presented in the tables
+and potential outliers. These information is also presented in Tables
 \ref{tab:continuous_numeric_attributes_salary} and
@@
 
-From the Discret Nominal Attributes tables we can see that each dataset contains
-only one attribute with missing values. The Salary Prediction dataset is missing
-the \textit{gender} attribute, while the Stroke Prediction dataset is missing the
-\textit{married} attribute. Also, the number of unique values for each attribute
+From the Discret Nominal Attributes tables (Tables \ref{tab:discrete_nominal_attributes_salary} and \ref{tab:discrete_nominal_attributes_stroke})
+we can see that each dataset contains
+only one attribute with missing values. In the Salary Prediction dataset 
+the \textit{gender} attribute is missing, while the Stroke Prediction dataset the
+\textit{married} attribute is missing. Also, the number of unique values for each attribute
 descibes the diversity of the data. For example, the \textit{country} attribute
@@
 
+In Figures \ref{fig:distribution_salary} and \ref{fig:distribution_stroke} we can
+see the distribution of each class in the datasets. The class distributions provide
+insights into the balance of the data and can help guide the selection of appropriate
+strategies for handling imbalanced classes. For example, in the Stroke Prediction
+dataset, the \textit{cerebrovascular\_accident} class is highly imbalanced, with
+a significantly higher number of negative instances compared to positive instances.
+This imbalance can impact the model's ability to learn patterns from the minority
+class and may require resampling techniques or cost-sensitive learning to address.
+On the other hand, the Salary Prediction dataset exhibits a more balanced distribution
+of the \textit{money} class, which may require less intervention to handle class
+imbalance.
+
 \begin{figure}[H]
     \centering
-    \includegraphics[width=0.8\textwidth]{../plots/distribution_edu_Salary_Prediction.png}
-    \caption{Distribution of the \textit{edu} attribute in the Salary Prediction dataset}
-    \label{fig:distribution_example_salary}
+    \includegraphics[width=0.8\textwidth]{../plots/distribution_SalaryPrediction.png}
+    \caption{Distribution of the \textit{money} class in the Salary Prediction dataset}
+    \label{fig:distribution_salary}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{../plots/distribution_tobacco_usage_Stroke_Prediction.png}
-    \caption{Distribution of the \textit{tobacco\_usage} attribute in the Stroke Prediction dataset}
-    \label{fig:distribution_example_stroke}
+    \includegraphics[width=0.8\textwidth]{../plots/distribution_AVC.png}
+    \caption{Distribution of the \textit{cerebrovascular\_accident} class in the Stroke Prediction dataset}
+    \label{fig:distribution_stroke}
 \end{figure}
-
-In the \textit{'plots'} folder at the root of the project directory we can find
-the distribution of each class in the datasets. The name of each plot starts
-with \textit{'distribution\_'}. Here is an example of the distribution of the
-\textit{edu} attribute in the Salary Prediction dataset in Figure
-\ref{fig:distribution_example_salary}. Also, the distribution of the
-\textit{tobacco\_usage} attribute in the Stroke Prediction dataset is shown in
-Figure \ref{fig:distribution_example_stroke}.
 
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_SalaryPrediction_full.png}
+    \caption{Correlation matrix for the Salary Prediction dataset}
+    \label{fig:correlation_matrix_example_salary}
+\end{figure}
+\pagebreak
 The correlation matrix and Cramér's V matrix provide valuable insights into the
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_AVC_full.png}
+    \caption{Correlation matrix for the Stroke Prediction dataset}
+    \label{fig:correlation_matrix_example_stroke}
+\end{figure}
+
 The Cramér's V matrix in Figures \ref{fig:cramer_v_matrix_example_salary} and
@@
 the discrete nominal attributes.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_SalaryPrediction_full.png}
-    \caption{Correlation matrix for the Salary Prediction dataset}
-    \label{fig:correlation_matrix_example_salary}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_AVC_full.png}
-    \caption{Correlation matrix for the Stroke Prediction dataset}
-    \label{fig:correlation_matrix_example_stroke}
-\end{figure}
 
@@
     \bibitem{}
+    \href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.html}{Pandas - DataFrame}
+    \bibitem{}
     \href{https://en.wikipedia.org/wiki/Correlation}{Wikipedia - Correlation}

### 2024-05-23 homework2/plots/distributions.py

@@ -2,62 +2,36 @@ import pandas as pd
 import matplotlib.pyplot as plt
 import sys
 import os
-  
+
+def distribution_of_classes(filename_prefix):
+	filename_train = filename_prefix + '_train.csv'
+	filename_test = filename_prefix + '_test.csv'
+
+	# get the distribution of classes
+	train = pd.read_csv(filename_train)
+	test = pd.read_csv(filename_test)
+
+	last_column = train.columns[-1]
+
+	train_class_distribution = train[last_column].value_counts()
+	test_class_distribution = test[last_column].value_counts()
+
+	# plot the distribution of classes
+	plt.figure()
+	plt.bar(train_class_distribution.index, train_class_distribution.values, alpha=0.5, label='train')
+	plt.bar(test_class_distribution.index, test_class_distribution.values, alpha=0.5, label='test')
+	plt.xticks(train_class_distribution.index)
+	plt.legend()
+	plt.title(f'Distribution of classes for {os.path.basename(filename_prefix)} dataset')
+	plt.savefig(f'plots/distribution_{os.path.basename(filename_prefix)}.png', dpi=300)
+	plt.close()
 
 if __name__ == "__main__":
     # Salary Prediction datasets
-	full_set = pd.read_csv('./tema2_SalaryPrediction/SalaryPrediction_full.csv')
-	train_set = pd.read_csv('./tema2_SalaryPrediction/SalaryPrediction_train.csv')
-	test_set = pd.read_csv('./tema2_SalaryPrediction/SalaryPrediction_test.csv')
+	filename_prefix = './tema2_SalaryPrediction/SalaryPrediction'
+	distribution_of_classes(filename_prefix)
 
-	full_set_features = full_set.select_dtypes(include=['object', 'category']).columns
-	train_set_features = train_set.select_dtypes(include=['object', 'category']).columns
-	test_set_features = test_set.select_dtypes(include=['object', 'category']).columns
-
-	for feature in full_set_features:
-		plt.figure()
-		# convert nan values to string
-		full_set[feature] = full_set[feature].astype(str)
-		train_set[feature] = train_set[feature].astype(str)
-		test_set[feature] = test_set[feature].astype(str)
-
-		plt.hist(full_set[feature], bins=50, alpha=0.5, label='Full Dataset')
[…]

### 2024-05-24 homework2/paper/paper.tex
finish 3rd section

@@
         \hline
+        money & categorical & Whether the individual earns more than \$50,000 per year \\
+        \hline
         \caption{Salary Prediction Attributes}
@@
 \begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{0.95cm}||>{\centering\arraybackslash}m{2.45cm}|>{\centering\arraybackslash}m{2.45cm}|>{\centering\arraybackslash}m{2.45cm}|>{\centering\arraybackslash}m{2.45cm}|>{\centering\arraybackslash}m{2.45cm}| }
         \hline
-        \multicolumn{5}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
+        \multicolumn{6}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
         \hline
-        & mean\_blood\_ sugar\_level & cardiovascular\_ issues & body\_mass\_ indicator & high\_ blood\_pressure \\
+        & mean\_blood\_ sugar\_level & body\_mass\_ indicator & years\_old & analysis\_results & biological\_age\_ index \\
         \hline\hline
-        count & 5110.000000 & 5110.000000 & 4909.000000 & 5110.000000 \\
+        count & 5110.000000 & 4909.000000 & 5110.000000 & 4599.000000 & 5110.000000 \\
         \hline
-        mean & 106.147677 & 0.054012 & 28.893237 & 0.097456 \\
+        mean & 106.147677 & 28.893237 & 46.568665 & 323.523446 & 134.784256 \\
         \hline
-        std & 45.283560 & 0.226063 & 7.854067 & 0.296607 \\
+        std & 45.283560 & 7.854067 & 26.593912 & 101.577442 & 50.399352 \\
         \hline
-        min & 55.120000 & 0.000000 & 10.300000 & 0.000000 \\
+        min & 55.120000 & 10.300000 & 0.080000 & 104.829714 & -15.109456 \\
         \hline
-        25\% & 77.245000 & 0.000000 & 23.500000 & 0.000000 \\
+        25\% & 77.245000 & 23.500000 & 26.000000 & 254.646209 & 96.710581 \\
         \hline
-        50\% & 91.885000 & 0.000000 & 28.100000 & 0.000000 \\
+        50\% & 91.885000 & 28.100000 & 47.000000 & 301.031628 & 136.374631 \\
         \hline
-        75\% & 114.090000 & 0.000000 & 33.100000 & 0.000000 \\
+        75\% & 114.090000 & 33.100000 & 63.750000 & 362.822769 & 172.507322 \\
         \hline
-        max & 271.740000 & 1.000000 & 97.600000 & 1.000000 \\
-        \hline
-    \end{longtable}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{1cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
-        \hline
-        \multicolumn{5}{|c|}{List of all Continuous Numeric Attributes in the Stroke Prediction dataset} \\
-        \hline
-        & years\_old & chaotic\_sleep & analysis\_results & biological\_age\_index \\
-        \hline\hline
-        count & 5110.000000 & 5110.000000 & 4599.000000 & 5110.000000 \\
-        \hline
-        mean & 46.568665 & 0.054012 & 323.523446 & 134.784256 \\
-        \hline
-        std & 26.593912 & 0.226063 & 101.577442 & 50.399352 \\
-        \hline
-        min & 0.080000 & 0.000000 & 104.829714 & -15.109456 \\
-        \hline
-        25\% & 26.000000 & 0.000000 & 254.646209 & 96.710581 \\
-        \hline
-        50\% & 47.000000 & 0.000000 & 301.031628 & 136.374631 \\
-        \hline
-        75\% & 63.750000 & 0.000000 & 362.822769 & 172.507322 \\
-        \hline
-        max & 134.000000 & 1.000000 & 756.807975 & 266.986321 \\
+        max & 271.740000 & 97.600000 & 134.000000 & 756.807975 & 266.986321 \\
         \hline
@@
         \hline
+        money & 9999 & 2 \\
+        \hline
         \caption{Discrete Nominal Attributes in Salary Prediction Dataset}
@@
         \hline\hline
+        cardiovascular\_issues & 5110 & 2 \\
+        \hline
         job\_category & 5110 & 5 \\
@@
         \hline
+        high\_blood\_pressure & 5110 & 2 \\
+        \hline
         married & 4599 &2 \\
@@
         living\_area & 5110 & 2 \\
+        \hline
+        chaotic\_sleep & 5110 & 2 \\
+        \hline
+        cerebrovascular\_accident & 5110 & 2 \\
         \hline
@@
 The correlation matrix and Cramér's V matrix provide valuable insights into the
-relationships between attributes in the datasets.By examining these matrices, we
+relationships between attributes in the datasets. By examining these matrices, we
 can see that Figure \ref{fig:correlation_matrix_example_salary} the \textit{prod}
@@
 In Figure \ref{fig:correlation_matrix_example_stroke} we can see that the
-\textit{cardiovascular\_issues} attribute is highly correlated with the
-\textit{chaotic\_sleep} attribute, while the \textit{body\_mass\_indicator} 
-attribute is negatively correlated with the \textit{cerebrovascular\_accident}
-attribute.
+\textit{mean\_blood\_sugar\_level} attribute is highly correlated with the
+\textit{analysis\_results} attribute, while the \textit{body\_mass\_indicator} 
+attribute is negatively correlated with the \textit{analysis\_results} attribute.
 
@@
 the \textit{gtype} attribute is strongly associated with the \textit{gender} attribute,
-while in the Stroke Prediction dataset, there are no strong associations between
-the discrete nominal attributes.
+while in the Stroke Prediction dataset, the \textit{cardiovascular\_issues} attribute is strongly
+associated with the \textit{chaotic\_sleep} attribute.
 
@@
 \section{Data Preprocessing}
+As highlighted in the previous section, high-quality data is the cornerstone of 
+effective machine learning models. However, real-world datasets often exhibit 
+various imperfections that can impede model performance.  Our exploration of the 
+datasets revealed the presence of several such issues, including:
+
+\begin{itemize}
+    \item Missing values for specific attributes.
+    \item Extreme values (outliers) within certain attributes.
+    \item Redundant attributes with high correlation.
+    \item Inconsistent value ranges for numeric attributes.
+\end{itemize}
+
+These imperfections necessitate data preprocessing, a crucial step aimed at 
+transforming the raw data into a clean and consistent format. This section 
+delves into the specific data preprocessing techniques employed in this study. 
+By addressing these issues, we aim to optimize the data for subsequent machine 
+learning algorithms, ultimately enhancing their effectiveness in extracting 
+valuable insights.
+
+As a note, all the scripts for data preprocessing are located in the \textit{'preprocessing'}
+folder at the root of the project directory.
+
+\subsection{Handling Missing Values}
+Missing data, a common issue in real-world datasets, necessitates the application
+of imputation procedures to address these missing values.  Imputation techniques 
+can be categorized as either univariate or multivariate:
+
+\begin{itemize}
+    \item \textit{Univariate Imputation:} This approach focuses solely on the attribute with missing 
+    values. Common univariate techniques include replacing missing values with the 
+    mean, median, or most frequent value within the attribute. These methods are 
+    simple to implement but may not effectively capture the underlying relationships 
+    between attributes.
+    \item \textit{Multivariate Imputation:} This more sophisticated approach leverages the values 
+    of other attributes within a sample to estimate the missing value. Techniques 
+    like regression analysis are often employed to establish relationships between 
+    the missing attribute and the remaining attributes. Based on these relationships, 
+    a predicted value can be imputed for the missing data point.  Multivariate 
+    imputation offers a more nuanced approach but requires careful consideration 
+    of the relationships between attributes and potential biases in the imputation 
+    process.
+\end{itemize}
+
+In the \textit{'impute\_values.py'} script, in the \textit{'missing\_values'} function,
+we used the \textit{IterativeImputer}
+class from the \textit{sklearn.impute} module to apply multivariate imputation to
+address missing values in the datasets for continuous numeric attributes. The script
+uses the most frequent value strategy for categorical attributes. The imputed datasets
+are saved in the same folder as the original datasets, with the prefix
+\textit{'preprocessed\_missing\_'}.
+
+\subsection{Outlier Detection and Treatment}
+Outliers, data points that deviate significantly from the rest of the dataset,
+can adversely affect the performance of machine learning models. Outliers can
+skew statistical measures, distort relationships between attributes, and lead to
+poor generalization of the model. Detecting and treating outliers is essential
+for ensuring the robustness and reliability of the model.
+
+We purpose to impute the outliers using the \textit{IsolationForest} algorithm
+from the \textit{sklearn.ensemble} module. The script \textit{'outlier\_detection.py'}
+detects outliers in the continuous numeric attributes of the datasets and replaces
+them with the imputed values. The preprocessed datasets with imputed outliers are
+saved in the same folder as the original datasets, with the prefix
+\textit{'preprocessed\_outliers\_'}.
+
+\subsection{Analysis of Attribute Correlations}
+As previously discussed, attribute correlations can provide valuable insights
+into the relationships between different attributes in the dataset. By identifying
+highly correlated attributes, we can eliminate redundant information and reduce
+the dimensionality of the data, leading to more efficient model training and
+improved interpretability.
+
+We choose to remove highly correlated attributes found in the section of
+Exploratory Data Analysis. These attributes are:
+
+\begin{itemize}
+    \item \textit{prod}: it is correlated with \textit{gain} in the Salary Prediction dataset.
+    \item \textit{analysis\_results}: it is correlated with \textit{body\_mass\_indicator} in the Stroke Prediction dataset.
+    \item \textit{gtype}: it is correlated with \textit{gender} in the Salary Prediction dataset.
+    \item 
+\end{itemize}
+
+The script \textit{'remove\_correlated\_attributes.py'}
+removes those attributes from the train dataset and saves the preprocessed dataset
+in the same folder as the original datasets, with the prefix \textit{'preprocessed\_correlated\_'}.
+
+\subsection{Normalization and Standardization}
+The numerical attributes in the dataset can vary significantly in their value scales. For example, 
+some attributes may have values in the thousands, while others have values in the single digits. 
+This disparity in scales can significantly affect algorithms like Logistic Regression.
+
+In algorithms like Logistic Regression, which rely on a linear combination of attribute values,
+attributes with larger numerical values can disproportionately influence the model. This dominance 
+can lead to biased results and reduce the model's effectiveness.
+
+To mitigate this issue, it is essential to standardize the values of the numeric attributes. 
+Standardization adjusts the scales of the attributes, ensuring that each one contributes equally 
+to the model's predictions. This process improves the performance and accuracy of the model by 
+creating a more balanced and fair representation of the data.
 

### 2024-05-24 homework2/plots/analysis_attributes.py

@@ -13,7 +13,11 @@ def analyze_csv(file_path, if_save=False):
     print(desc_stats)
     
     # Generate box plots for numerical attributes
-    numerical_cols = df.select_dtypes(include=['number']).columns
+    with open(os.path.join(os.path.dirname(file_path), 'numeric_attributes.txt'), 'r') as f:
+        numerical_cols = f.read().splitlines()
+
+    print("\nNumerical Attributes:")
+    print(numerical_cols)
     
     if len(numerical_cols) > 0:
         # Create a box plot for each numerical attribute
@@ -30,8 +34,12 @@ def analyze_csv(file_path, if_save=False):
         print("No numerical attributes found for box plots.")
     
      # Analyze categorical/ordinal attributes
-    categorical_cols = df.select_dtypes(include=['object', 'category']).columns
-    
+    with open(os.path.join(os.path.dirname(file_path), 'categorical_attributes.txt'), 'r') as f:
+        categorical_cols = f.read().splitlines()
+
+    print("\nCategorical/Ordinal Attributes:")
+    print(categorical_cols)
+
     if len(categorical_cols) > 0:
         categorical_stats = {}
         for col in categorical_cols:
@@ -73,11 +81,13 @@ def analyze_csv(file_path, if_save=False):
             
 
 if __name__ == "__main__":
-    file = './tema2_SalaryPrediction/SalaryPrediction_full.csv'
+    # file = './tema2_SalaryPrediction/SalaryPrediction_full.csv'
+    file = './tema2_SalaryPrediction/preprocessed_outliers_SalaryPrediction_train.csv'
     analyze_csv(file, if_save=True)
     print(f"\nAnalysis completed for {file}\n")
 
-    file = './tema2_AVC/AVC_full.csv'
+    # file = './tema2_AVC/AVC_full.csv'
+    file = './tema2_AVC/preprocessed_outliers_AVC_train.csv'
     analyze_csv(file, if_save=True)
     print(f"\nAnalysis completed for {file}\n")
 

### 2024-05-24 homework2/plots/correlation_analysis.py

@@ -7,7 +7,8 @@ import os
 
 def numerical_features(file_name):
 	file = pd.read_csv(file_name)
-	cols = file.select_dtypes(include=['number']).columns
+	with open(os.path.join(os.path.dirname(file_name), 'numeric_attributes.txt'), 'r') as f:
+		cols = f.read().splitlines()
 
 	# Create a correlation matrix
 	corr_matrix = file[cols].corr()
@@ -19,11 +20,15 @@ def numerical_features(file_name):
 	plt.xticks(np.arange(len(cols)), cols, rotation=45)
 	plt.yticks(np.arange(len(cols)), cols)
 	plt.title(f'Correlation Matrix for {os.path.basename(file_name).split(".")[0]} dataset')
-	# make the plot bigger
 	fig = plt.gcf()
 	fig.set_size_inches(15, 10.5)
 	plt.savefig(f'plots/correlation_matrix_{os.path.basename(file_name).split(".")[0]}.png', dpi=300)
 	plt.close()
+
+	for i in range(len(cols)):
+		for j in range(i+1, len(cols)):
+			if corr_matrix.iloc[i, j] > 0.8:
+				print(f'High correlation between {cols[i]} and {cols[j]}')
 
 def cramers_v(confusion_matrix):
     chi2 = chi2_contingency(confusion_matrix)[0]
@@ -49,12 +54,11 @@ def cramers_v_matrix(df):
 
 def cathegorical_features(file_name):
 	file = pd.read_csv(file_name)
-	cols = file.select_dtypes(include=['object', 'category']).columns
+	with open(os.path.join(os.path.dirname(file_name), 'categorical_attributes.txt'), 'r') as f:
+		cols = f.read().splitlines()
 
 	corelation = cramers_v_matrix(file[cols])
 
-	# print(f'Cramer\'s V matrix:\n{corelation}')
-	# print high corelation values in a plot
 	plt.figure()
 	plt.matshow(corelation, cmap='coolwarm')
 	plt.colorbar()

### 2024-05-24 homework2/preprocessing/impute_values.py

@@ -1,0 +1,122 @@
+import numpy as np
+import pandas as pd
+import os
+from sklearn.experimental import enable_iterative_imputer
+from sklearn.impute import IterativeImputer, SimpleImputer
+from scipy import stats
+
+
+def missing_values(file_name):
+	file = pd.read_csv(file_name)
+
+	# convert ? to NaN
+	file = file.replace('?', np.nan)
+
+	# Separate numeric and categorical columns
+	with open(os.path.join(os.path.dirname(file_name), 'numeric_attributes.txt'), 'r') as f:
+		numerical_cols = f.read().splitlines()
+
+    # Generate box plots for numerical attributes
+	with open(os.path.join(os.path.dirname(file_name), 'categorical_attributes.txt'), 'r') as f:
+		categorical_cols = f.read().splitlines()
+
+	# Impute missing values in numeric columns
+	imputer = IterativeImputer(max_iter=10, random_state=0)
+	imputed_numeric = imputer.fit_transform(file[numerical_cols])
+
+	# Impute missing values in categorical columns: missing values can contain ? or NaN or nothing
+	# Use the most frequent value for each column
+	imputed_categorical = file[categorical_cols].apply(lambda x: x.fillna(x.value_counts().index[0]))
+
+	# Concatenate the imputed numeric and categorical columns
+	imputed_file = pd.concat([pd.DataFrame(imputed_numeric, columns=numerical_cols), imputed_categorical], axis=1)
+
+	return imputed_file
+	
+def impute_outliers(file_name):
+	file = pd.read_csv(file_name)
+	
+	# Separate numeric columns
+	with open(os.path.join(os.path.dirname(file_name), 'numeric_attributes.txt'), 'r') as f:
+		numerical_cols = f.read().splitlines()
+
+	# Calculate Q1 and Q3 for each numerical column
+	for col in numerical_cols:
+		Q1 = file[col].quantile(0.25)
+		Q3 = file[col].quantile(0.75)
+		IQR = Q3 - Q1
+
+		# Get the indices of the outliers
+		threshold = 1.5
+		outliers = file[(file[col] < Q1 - threshold * IQR) | (file[col] > Q3 + threshold * IQR)].index
+
+		# Replace the outliers with the median value
+		file.loc[outliers, col] = file[col].median()
+
+	return file
+
+if __name__ == '__main__':
[…]

### 2024-05-24 homework2/preprocessing/remove_correlated_attributes.py

@@ -1,0 +1,43 @@
+# remove first 10 attributes
+import pandas as pd
+
+
+
+def remove_correlated_attributes(file_name, attributes_to_remove):
+	file = pd.read_csv(file_name)
+	correlated_file = file.drop(columns=attributes_to_remove)
+	return correlated_file
+
+if __name__ == '__main__':
+	# Salary Prediction dataset
+	file_name = './tema2_SalaryPrediction/preprocessed_outliers_SalaryPrediction_train.csv'
+	file_sufix = 'SalaryPrediction_train.csv'
+	print(f'Preprocessing {file_sufix} dataset')
+	attributes_to_remove = ['prod', 'gtype']
+	correlated_file = remove_correlated_attributes(file_name, attributes_to_remove)
+	correlated_file.to_csv(f'./tema2_SalaryPrediction/preprocessed_correlated_{file_sufix}', index=False)
+	
+
+	file_name = './tema2_SalaryPrediction/preprocessed_outliers_SalaryPrediction_test.csv'
+	file_sufix = 'SalaryPrediction_test.csv'
+	print(f'Preprocessing {file_sufix} dataset')
+	attributes_to_remove = ['prod', 'gtype']
+	correlated_file = remove_correlated_attributes(file_name, attributes_to_remove)
+	correlated_file.to_csv(f'./tema2_SalaryPrediction/preprocessed_correlated_{file_sufix}', index=False)
+
+
+	# Stroke Prediction dataset
+	file_name = './tema2_AVC/preprocessed_outliers_AVC_train.csv'
+	file_sufix = 'AVC_train.csv'
+	print(f'Preprocessing {file_sufix} dataset')
+	attributes_to_remove = ['analysis_results', 'chaotic_sleep']
+	correlated_file = remove_correlated_attributes(file_name, attributes_to_remove)
+	correlated_file.to_csv(f'./tema2_AVC/preprocessed_correlated_{file_sufix}', index=False)
+
+
+	file_name = './tema2_AVC/preprocessed_outliers_AVC_test.csv'
+	file_sufix = 'AVC_test.csv'
+	print(f'Preprocessing {file_sufix} dataset')
+	attributes_to_remove = ['analysis_results', 'chaotic_sleep']
+	correlated_file = remove_correlated_attributes(file_name, attributes_to_remove)
+	correlated_file.to_csv(f'./tema2_AVC/preprocessed_correlated_{file_sufix}', index=False)

### 2024-05-24 homework2/preprocessing/standardization.py

@@ -1,0 +1,55 @@
+from sklearn import preprocessing
+import numpy as np
+import pandas as pd
+import os
+
+def standardize(train_file, test_file):
+	train = pd.read_csv(train_file)
+	test = pd.read_csv(test_file)
+
+	# Extract the features and the target
+	header = train.columns
+	print(header)
+	X_train = train.iloc[:, :-1].values
+	y_train = train.iloc[:, -1].values
+
+	X_test = test.iloc[:, :-1].values
+	y_test = test.iloc[:, -1].values
+
+	# Separates categorical and numerical attributes
+	categorical_attributes = []
+	numerical_attributes = []
+	with open(os.path.join(os.path.dirname(train_file), 'numeric_attributes.txt'), 'r') as f:
+		numerical_attributes = f.read().splitlines()
+
+	for col in numerical_attributes:
+		if col not in train.columns:
+			numerical_attributes.remove(col)
+
+	with open(os.path.join(os.path.dirname(train_file), 'categorical_attributes.txt'), 'r') as f:
+		categorical_attributes = f.read().splitlines()
+
+	# Standardize only the numerical attributes
+	scaler = preprocessing.StandardScaler()
+	X_train[:, [train.columns.get_loc(col) for col in numerical_attributes]] = scaler.fit_transform(X_train[:, [train.columns.get_loc(col) for col in numerical_attributes]])
+	X_test[:, [test.columns.get_loc(col) for col in numerical_attributes]] = scaler.transform(X_test[:, [test.columns.get_loc(col) for col in numerical_attributes]])
+
+	# Save the standardized datasets and add headers
+	train_standardized = pd.DataFrame(X_train, columns=header[:-1])
+	train_standardized[header[-1]] = y_train
+	test_standardized = pd.DataFrame(X_test, columns=header[:-1])
+	test_standardized[header[-1]] = y_test
+
+	train_standardized.to_csv(train_file.replace('.csv', '_standardized.csv'), index=False)
+	test_standardized.to_csv(test_file.replace('.csv', '_standardized.csv'), index=False)
+
+if __name__ == '__main__':
+	# Salary Prediction dataset
+	train_file = './tema2_SalaryPrediction/preprocessed_correlated_SalaryPrediction_train.csv'
+	test_file = './tema2_SalaryPrediction/preprocessed_correlated_SalaryPrediction_test.csv'
+	standardize(train_file, test_file)
+
[…]

### 2024-05-25 homework2/paper/paper.tex
write introduction for chapter 4

@@
 \section{Algorithms Designs}
+Algorithm design is a critical aspect of computer science and machine learning, 
+focusing on creating efficient and effective methods to solve complex problems. 
+The process involves the careful selection of algorithms based on the specific 
+characteristics of the data and the desired outcomes. This document explores the 
+application of two prominent machine learning algorithms, Linear Regression and 
+Multi-Layered Perceptron (MLP), on diverse datasets. The goal is to compare their 
+performance and suitability for different types of prediction tasks, particularly 
+in the contexts of stroke prediction and salary prediction.
+
+\subsection{Linear Regression}
+Linear Regression is one of the most fundamental and widely used algorithms in 
+machine learning. It is known for its simplicity, interpretability, and 
+efficiency in identifying linear relationships between variables. The primary 
+purpose of Linear Regression is to model the relationship between a dependent 
+variable and one or more independent variables by fitting a linear equation to 
+observed data.
+
+In the context of stroke prediction, Linear Regression can be used to analyze 
+how various factors such as age, blood pressure, cholesterol levels, and smoking 
+history influence the likelihood of experiencing a stroke. Its straightforward 
+approach allows for clear insights into the significance and impact of each 
+factor on stroke risk.
+
+\subsection{Multi-Layered Perceptron (MLP)}
+Multi-Layer Perceptron (MLP) is a type of artificial neural network designed to 
+capture complex, non-linear patterns in data. Unlike Linear Regression, which is
+limited to linear relationships, MLPs consist of multiple layers of interconnected 
+nodes, or neurons, that can model intricate interactions between variables.
+
+MLPs are particularly suitable for tasks involving complex datasets with 
+non-linear relationships, such as salary prediction. In salary prediction, 
+factors like education, work experience, industry sector, and other 
+socio-economic variables interact in complex ways that Linear Regression might 
+not fully capture. MLPs can learn these non-linear patterns, providing a more 
+accurate and nuanced understanding of the factors influencing salary levels.
 

### 2024-05-25 homework2/linear_regression/convert_categorical_attributes.py
convert categorical features for linear regression

@@ -1,0 +1,95 @@
+import numpy as np
+import os
+import sys
+import pandas as pd
+import matplotlib.pyplot as plt
+
+from sklearn.metrics import classification_report
+
+
+def convert_categorical_attributes(X_train, X_test, cat_attr, header):
+# Convert categorical attributes to one-hot encoded format
+    cat_indices = [list(header).index(attr) for attr in cat_attr]
+
+    categorical_data = X_train[:, cat_indices]
+    categorical_df = pd.DataFrame(categorical_data, columns=cat_attr)
+    XX_train = pd.get_dummies(categorical_df, prefix=cat_attr)
+
+    categorical_data = X_test[:, cat_indices]
+    categorical_df = pd.DataFrame(categorical_data, columns=cat_attr)
+    XX_test = pd.get_dummies(categorical_df, prefix=cat_attr)
+
+    XX_train = XX_train.reindex(columns=XX_test.columns, fill_value=0)
+    XX_test = XX_test.reindex(columns=XX_train.columns, fill_value=0)
+
+    # Drop original categorical columns
+    X_train = np.delete(X_train, cat_indices, axis=1)
+    X_test = np.delete(X_test, cat_indices, axis=1)
+    header = np.delete(header, cat_indices)
+
+    # Concatenate the original array with the new one-hot encoded columns
+    X_train = np.concatenate((X_train, XX_train.values), axis=1)
+    X_test = np.concatenate((X_test, XX_test.values), axis=1)
+    header = np.concatenate((header, XX_train.columns))
+
+
+    return X_train, X_test, header
+
+
+if __name__ == "__main__":
+    # get the type of dataset
+    if len(sys.argv) < 2:
+        print("Usage: python convert_categorical_attributes.py <dataset>")
+        sys.exit(1)
+    dataset = sys.argv[1]
+    train_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_train_standardized.csv'
+    test_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_test_standardized.csv'
+    
+    if not os.path.exists(train_file) or not os.path.exists(test_file):
+        print("Please standardize the dataset first")
+        sys.exit(1)
+
+    categorical_attributes = []
+    with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
+        categorical_attributes = f.read().splitlines()
+    
+    # check if all the categorical attributes are in the dataset
+    header = pd.read_csv(train_file).columns
[…]

### 2024-05-25 homework2/preprocessing/standardization.py

@@ -50,6 +50,6 @@ if __name__ == '__main__':
 	standardize(train_file, test_file)
 
 	# Stroke Prediction dataset
-	# train_file = './tema2_AVC/preprocessed_correlated_AVC_train.csv'
-	# test_file = './tema2_AVC/preprocessed_correlated_AVC_test.csv'
-	# standardize(train_file, test_file)
+	train_file = './tema2_AVC/preprocessed_correlated_AVC_train.csv'
+	test_file = './tema2_AVC/preprocessed_correlated_AVC_test.csv'
+	standardize(train_file, test_file)

### 2024-05-25 homework2/linear_regression/scikit-learn_implementation.py
scikit-leanr implementation for linear regression

@@ -1,0 +1,96 @@
+import numpy as np
+import os
+import sys
+import pandas as pd
+from sklearn.linear_model import LinearRegression
+import matplotlib.pyplot as plt
+from sklearn.linear_model import Ridge as RidgeRegression
+import seaborn as sns
+from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
+
+
+
+if __name__ == "__main__":
+	# get the type of dataset
+	if len(sys.argv) < 2:
+		print("Usage: python convert_categorical_attributes.py <dataset>")
+		sys.exit(1)
+	dataset = sys.argv[1]
+	train_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_train_standardized_one_hot.csv'
+	test_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_test_standardized_one_hot.csv'
+	
+	# Load the datasets
+	train = pd.read_csv(train_file)
+	test = pd.read_csv(test_file)
+
+	# get the target column
+	categorical_attributes = []
+	with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
+		categorical_attributes = f.read().splitlines()
+	target = categorical_attributes[-1]
+
+	X_train = train.drop(target, axis=1).values
+	y_train = train[target].values
+
+	X_test = test.drop(target, axis=1).values
+	y_test = test[target].values
+
+	# Create and train the linear regression model
+	model =  RidgeRegression(alpha=1.75)
+	model.fit(X_train, y_train)
+
+	# Make predictions
+	y_train_pred = model.predict(X_train)
+	y_test_pred = model.predict(X_test)
+
+	# convert the predictions to binary
+	# get the maximum value of the predictions
+	mean_train = np.mean(y_train_pred)
+	y_train_pred = np.where(y_train_pred > mean_train, 1, 0)
+	mean_test = np.mean(y_test_pred)
+	y_test_pred = np.where(y_test_pred > mean_test, 1, 0)
+
+	# Evaluate the model
+	# Training metrics
+	train_accuracy = accuracy_score(y_train, y_train_pred)
+	train_precision = precision_score(y_train, y_train_pred, average='weighted')
+	train_recall = recall_score(y_train, y_train_pred, average='weighted')
[…]

### 2024-05-25 homework2/linear_regression/manual_implementation.py
finish implementing both methods for linear regression

@@ -1,0 +1,141 @@
+import numpy as np
+import os
+import sys
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
+
+
+class LinearRegression(object):
+    def fit(self, X, t):
+        N, D = X.shape
+        
+        X_transpose = np.transpose(X)
+        # Use pseudo-inverse instead of inverse
+        self.w = np.linalg.pinv(X_transpose @ X) @ X_transpose @ t
+        
+        self.var = np.mean(np.square(X @ self.w - t))
+
+    def predict(self, X, return_std=False):
+        N, D = X.shape
+        y = X @ self.w
+        
+        if return_std:
+            y_std = np.sqrt(self.var)
+            return y, y_std
+
+        return y
+
+
+class RidgeRegression(LinearRegression):
+    def __init__(self, alpha:float=1.):
+        super(RidgeRegression, self).__init__()
+        self.alpha = alpha
+        
+
+    def fit(self, X:np.ndarray, t:np.ndarray):
+        N, D = X.shape
+        X_transpose = np.transpose(X)
+        self.w = np.linalg.inv(X_transpose @ X + self.alpha * np.eye(D)) @ X_transpose @ t
+
+        self.var = np.mean(np.square(X @ self.w - t))
+
+
+if __name__ == "__main__":
+    # get the type of dataset
+    if len(sys.argv) < 2:
+        print("Usage: python convert_categorical_attributes.py <dataset>")
+        sys.exit(1)
+    dataset = sys.argv[1]
+    train_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_train_standardized_one_hot.csv'
+    test_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_test_standardized_one_hot.csv'
+    
+    # Load the datasets
+    train = pd.read_csv(train_file)
+    test = pd.read_csv(test_file)
+
+    # get the target column
+    categorical_attributes = []
+    with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
+        categorical_attributes = f.read().splitlines()
+    target = categorical_attributes[-1]
+
+    X_train = train.drop(target, axis=1)
+    y_train = train[target]
+
+    X_test = test.drop(target, axis=1)
+    y_test = test[target]
[…]

### 2024-05-25 homework2/linear_regression/scikit-learn_implementation.py

@@ -80,12 +80,16 @@ if __name__ == "__main__":
 	plt.xlabel('Predicted')
 	plt.ylabel('Actual')
 	plt.tight_layout()
-	
+
+	# check if the directory exists
+	if not os.path.exists(f'./tema2_{dataset}/LinearRegression'):
+		os.makedirs(f'./tema2_{dataset}/LinearRegression')
+
 	# Save the confusion matrix
-	plt.savefig(f'./tema2_{dataset}/confusion_matrix.png')
+	plt.savefig(f'./tema2_{dataset}/LinearRegression/confusion_matrix_scikit-learn.png')
 
 	# Save the metrics
-	with open(f'./tema2_{dataset}/metrics.txt', 'w') as f:
+	with open(f'./tema2_{dataset}/LinearRegression/metrics_scikit-learn.txt', 'w') as f:
 		f.write(f"Training accuracy: {train_accuracy}\n")
 		f.write(f"Training precision: {train_precision}\n")
 		f.write(f"Training recall: {train_recall}\n")

### 2024-05-25 homework2/paper/paper.tex
described in the paper the Linear Regression used

@@
 
+In the \textit{linear\_regression} folder at the root of the project directory,
+we implemented the Linear Regression algorithm in two different ways:
+
+\begin{itemize}
+    \item \textit{Linear Regression with Scikit-Learn:} We used the Scikit-Learn
+    library to implement Linear Regression on the preprocessed datasets.
+    \item \textit{Linear Regression from Scratch:} We implemented Linear Regression
+    as in the 8th laboratory from the Artificial Intelligence course.
+\end{itemize}
+
+I choose to implement the RidgeRegression algorithm. Because of the large number of
+features in the dataset, the RidgeRegression algorithm is more suitable than the
+LinearRegression algorithm. The RidgeRegression algorithm is a regularized version
+of Linear Regression that adds a penalty term to the loss function to prevent overfitting.
+This penalty term helps to reduce the model's complexity and improve its generalization
+performance. The alpha parameter controls the strength of the regularization, with
+higher values of alpha leading to stronger regularization. I used the value of alpha
+1.75 because it provided the best results in the experiments.
+
+For enconding the categorical attributes except the target attribute, I used the
+\textit{OneHotEncoder} class from the \textit{sklearn.preprocessing} module. This
+class encodes categorical attributes as one-hot vectors, creating a binary representation
+of each category. This encoding is essential for feeding categorical attributes into
+machine learning models, as most algorithms require numerical input data. For the
+target attribute, I used the \textit{LabelEncoder} class from the \textit{sklearn.preprocessing}
+module to encode the target attribute as integer values.
+
+The results of both implementations for each dataset are saved in the \textit{LinearRegression}
+folder at the specific dataset's root.
+
 \subsection{Multi-Layered Perceptron (MLP)}
@@
     \bibitem{}
+    \href{https://curs.upb.ro/2023/course/view.php?id=13749}{Moodle - Artifical Intelligence Course}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Linear_regression}{Wikipedia - Linear Regression}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Linear_classifier}{Wikipedia - Linear Classifier}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Ridge_regression}{Wikipedia - Ridge Regression}
+    \bibitem{}
+    \href{https://www.semanticscholar.org/reader/2c9022fe0af15568a885e59d475ec8f95726e51b}{Metrics for Multi-Class Classification: An Overview}
+    \bibitem{}
     \href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.html}{Pandas - DataFrame}
@@
     \bibitem{}
+    \href{https://saturncloud.io/blog/how-to-detect-and-exclude-outliers-in-a-pandas-dataframe/}{Saturn Cloud - How to Detect and Exclude Outliers in a Pandas DataFrame}
+    \bibitem{}
+    \href{https://www.kaggle.com/discussions/questions-and-answers/159183}{Kaggle - When to standardize test and train data?}
+    \bibitem{}
+    \href{https://scikit-learn.org/stable/modules/preprocessing.html#standardization-or-mean-removal-and-variance-scaling}{Scikit-Learn - Standardization or Mean Removal and Variance Scaling}
+    \bibitem{}
+    \href{https://scikit-learn.org/stable/modules/generated/sklearn.preprocessing.OneHotEncoder.html}{Scikit-Learn - OneHotEncoder}
+    \bibitem{}
+    \href{https://scikit-learn.org/stable/modules/generated/sklearn.preprocessing.LabelEncoder.html}{Scikit-Learn - LabelEncoder}
+    \bibitem{}
+    \href{https://towardsdatascience.com/understanding-confusion-matrix-a9ad42dcfd62}{Towards Data Science - Understanding Confusion Matrix}
 

### 2024-05-25 homework2/linear_regression/manual_implementation.py

@@ -10,11 +10,8 @@ from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_sc
 class LinearRegression(object):
     def fit(self, X, t):
         N, D = X.shape
-        
         X_transpose = np.transpose(X)
-        # Use pseudo-inverse instead of inverse
         self.w = np.linalg.pinv(X_transpose @ X) @ X_transpose @ t
-        
         self.var = np.mean(np.square(X @ self.w - t))
 
     def predict(self, X, return_std=False):

### 2024-05-25 homework2/paper/paper.tex
finished with logistic regression

@@
     This study investigates the applicability of machine learning techniques on 
-    diverse datasets. We explore the effectiveness of two algorithms, Linear 
+    diverse datasets. We explore the effectiveness of two algorithms, Logistic 
     Regression and Multi-Layered Perceptron (MLP), on predicting both health 
@@
 Prediction Modeling \and Data Analysis Techniques \and Stroke Prediction \and
-Salary Prediction \and Linear Regression \and Multi-Layered Perceptron \and 
+Salary Prediction \and Logistic Regression \and Multi-Layered Perceptron \and 
 DataPreprocessing} }
@@
 
-\subsection{Methodology: Unveiling the Algorithms - Linear Regression and Multi-Layered Perceptron}
-To investigate the performance on these contrasting datasets, this study employs 
-two prominent ML algorithms: Linear Regression and Multi-Layered Perceptron (MLP).
+\subsection{Methodology: Unveiling the Algorithms - Logistic Regression and Multi-Layered Perceptron}
+This section delves into the core methodologies employed in this study: 
+Logistic Regression and Multi-Layered Perceptron (MLP). Both algorithms fall 
+under the umbrella of supervised learning, where a model learns from labeled 
+data to make predictions on unseen examples. Here, we unveil the underlying 
+principles and functionalities of each technique.
 
-Linear Regression is a well\-established technique known for its interpretability
-and efficiency in uncovering linear relationships between features (data points) 
-and target variables (what we want to predict). This makes it a valuable tool 
-for understanding the underlying factors influencing a particular outcome, such 
-as the relationship between blood pressure and stroke risk. However, its strength 
-lies in capturing linear relationships. If the underlying relationships in the 
-data are more complex and non-linear, then Linear Regression might not be as 
-effective.
+Logistic regression serves as a foundational algorithm for classification tasks.
+It establishes a mathematical model that maps input features (e.g., age, blood 
+pressure) to a probability of a specific outcome (e.g., stroke occurrence). The 
+model essentially learns a decision boundary, separating observations with high 
+and low probabilities of the target outcome. This approach makes logistic 
+regression well-suited for analyzing datasets like the stroke prediction one,
+where the goal is to categorize individuals based on their risk level.
 
-On the other hand, Multi-Layered Perceptrons (MLPs) are a type of artificial 
-neural network capable of learning complex, non-linear patterns within data. 
-Unlike Linear Regression, MLPs are not limited by linearity and can potentially 
-capture more intricate relationships between features and target variables. This 
-capability makes them particularly suitable for datasets with complex underlying 
-structures, such as the factors influencing an individual's salary, which might 
-involve a combination of education, experience, industry, and other factors 
-interacting in non-linear ways.
+On the other hand, Multi-Layered Perceptron (MLP) represents a more complex 
+architecture - a type of artificial neural network. It consists of interconnected 
+layers of artificial neurons, mimicking the structure of the human brain. Each 
+layer transforms the received data using activation functions, ultimately leading 
+to an output prediction. MLP's strength lies in its ability to learn complex, 
+non-linear relationships within data. This makes it a powerful tool for tackling 
+intricate prediction problems, potentially outperforming logistic regression 
+when the underlying relationships are not easily captured by a linear model. 
+The salary prediction dataset, where various factors might influence income 
+levels, could be a prime candidate for exploration with MLP.
 
 \subsection{Research Objectives: Evaluating Algorithms, Unveiling Strengths and Weaknesses}
-By comparatively analyzing the performance of Linear Regression and MLP on 
+By comparatively analyzing the performance of Logistic Regression and MLP on 
 stroke and salary prediction tasks, this research seeks to achieve several key 
@@
 datasets. We will determine which algorithm performs better on each dataset, 
-offering insights into their suitability for different data types, such as 
-linear datasets (blood pressure and stroke risk) versus potentially non-linear 
-datasets (factors influencing salary).
+offering insights into their suitability for different data types.
 
@@
 selecting the most appropriate algorithm for their specific prediction tasks.
-For instance, if interpretability is crucial (e.g., understanding the factors 
-influencing stroke risk), Linear Regression might be preferred. If the data is 
-likely to have complex, non-linear relationships, then an MLP might be a better 
-choice.
 
@@
 characteristics of the data and the desired outcomes. This document explores the 
-application of two prominent machine learning algorithms, Linear Regression and 
+application of two prominent machine learning algorithms, Logistic Regression and 
 Multi-Layered Perceptron (MLP), on diverse datasets. The goal is to compare their 
@@
 
-\subsection{Linear Regression}
-Linear Regression is one of the most fundamental and widely used algorithms in 
-machine learning. It is known for its simplicity, interpretability, and 
-efficiency in identifying linear relationships between variables. The primary 
-purpose of Linear Regression is to model the relationship between a dependent 
-variable and one or more independent variables by fitting a linear equation to 
-observed data.
+\subsection{Logistic Regression}
+Logistic regression is a fundamental statistical method employed for classification 
+tasks in machine learning. It establishes a mathematical model that maps a set
+of input features (independent variables) to a probability of a specific outcome 
+(dependent variable). The core functionality lies in estimating the odds of a 
+particular class membership (e.g., presence of stroke) based on the input 
+features. The resulting model essentially learns a decision boundary, 
+separating observations with high and low probabilities of belonging to the 
+target class. This characteristic makes logistic regression particularly 
+well-suited for analyzing datasets where the outcome variable is binary 
+(e.g., stroke occurrence vs. no stroke occurrence).
 
-In the context of stroke prediction, Linear Regression can be used to analyze 
-how various factors such as age, blood pressure, cholesterol levels, and smoking 
-history influence the likelihood of experiencing a stroke. Its straightforward 
-approach allows for clear insights into the significance and impact of each 
-factor on stroke risk.
-
-In the \textit{linear\_regression} folder at the root of the project directory,
-we implemented the Linear Regression algorithm in two different ways:
+In the \textit{logistic\_regression} folder at the root of the project directory,
+we implemented the Logistic Regression algorithm in two different ways:
 
 \begin{itemize}
-    \item \textit{Linear Regression with Scikit-Learn:} We used the Scikit-Learn
-    library to implement Linear Regression on the preprocessed datasets.
-    \item \textit{Linear Regression from Scratch:} We implemented Linear Regression
-    as in the 8th laboratory from the Artificial Intelligence course.
+    \item \textit{Logistic Regression with Scikit-Learn:} We used the Scikit-Learn
+    library to implement Logistic Regression on the preprocessed datasets.
+    \item \textit{Logistic Regression from Scratch:} We implemented Logistic Regression
+    from scratch using the Negative Log-Likelihood method and the Gradient Descent
+    optimization algorithm.
 \end{itemize}
 
-I choose to implement the RidgeRegression algorithm. Because of the large number of
-features in the dataset, the RidgeRegression algorithm is more suitable than the
-LinearRegression algorithm. The RidgeRegression algorithm is a regularized version
-of Linear Regression that adds a penalty term to the loss function to prevent overfitting.
-This penalty term helps to reduce the model's complexity and improve its generalization
-performance. The alpha parameter controls the strength of the regularization, with
-higher values of alpha leading to stronger regularization. I used the value of alpha
-1.75 because it provided the best results in the experiments.
+Before starting the implementation of the Logistic Regression algorithm, we need to
+encode the categorical attributes in the datasets. Categorical attributes are non-numeric
+attributes that represent discrete categories or groups. These attributes need to be
+encoded into a numerical format before they can be used in machine learning algorithms.
 
@@
 
-The results of both implementations for each dataset are saved in the \textit{LinearRegression}
+In the Logistic Regression with Scikit-Learn implementation, we used the \textit{LogisticRegression}
+class from the \textit{sklearn.linear\_model} module to train the model on the preprocessed
+datasets, without setting any hyperparameters, using the default values.
+For the Logistic Regression from Scratch implementation, we implemented the Negative Log-Likelihood
+loss function and the Gradient Descent optimization algorithm. We trained the model on the
+preprocessed datasets, setting the learning rate to 0.01 and the number of epochs to 10000. For the
+regularization, we used the Ridge Regression technique.
+
+The results of both implementations for each dataset are saved in the \textit{LogisticRegression}
 folder at the specific dataset's root.

### 2024-05-25 homework2/linear_regression/convert_categorical_attributes.py

@@ -1,95 +1,0 @@
-import numpy as np
-import os
-import sys
-import pandas as pd
-import matplotlib.pyplot as plt
-
-from sklearn.metrics import classification_report
-
-
-def convert_categorical_attributes(X_train, X_test, cat_attr, header):
-# Convert categorical attributes to one-hot encoded format
-    cat_indices = [list(header).index(attr) for attr in cat_attr]
-
-    categorical_data = X_train[:, cat_indices]
-    categorical_df = pd.DataFrame(categorical_data, columns=cat_attr)
-    XX_train = pd.get_dummies(categorical_df, prefix=cat_attr)
-
-    categorical_data = X_test[:, cat_indices]
-    categorical_df = pd.DataFrame(categorical_data, columns=cat_attr)
-    XX_test = pd.get_dummies(categorical_df, prefix=cat_attr)
-
-    XX_train = XX_train.reindex(columns=XX_test.columns, fill_value=0)
-    XX_test = XX_test.reindex(columns=XX_train.columns, fill_value=0)
-
-    # Drop original categorical columns
-    X_train = np.delete(X_train, cat_indices, axis=1)
-    X_test = np.delete(X_test, cat_indices, axis=1)
-    header = np.delete(header, cat_indices)
-
-    # Concatenate the original array with the new one-hot encoded columns
-    X_train = np.concatenate((X_train, XX_train.values), axis=1)
-    X_test = np.concatenate((X_test, XX_test.values), axis=1)
-    header = np.concatenate((header, XX_train.columns))
-
-
-    return X_train, X_test, header
-
-
-if __name__ == "__main__":
-    # get the type of dataset
-    if len(sys.argv) < 2:
-        print("Usage: python convert_categorical_attributes.py <dataset>")
-        sys.exit(1)
-    dataset = sys.argv[1]
-    train_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_train_standardized.csv'
-    test_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_test_standardized.csv'
-    
-    if not os.path.exists(train_file) or not os.path.exists(test_file):
-        print("Please standardize the dataset first")
-        sys.exit(1)
-
-    categorical_attributes = []
-    with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
-        categorical_attributes = f.read().splitlines()
-    
-    # check if all the categorical attributes are in the dataset
-    header = pd.read_csv(train_file).columns
[…]

### 2024-05-25 homework2/linear_regression/manual_implementation.py

@@ -1,138 +1,0 @@
-import numpy as np
-import os
-import sys
-import pandas as pd
-import matplotlib.pyplot as plt
-import seaborn as sns
-from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
-
-
-class LinearRegression(object):
-    def fit(self, X, t):
-        N, D = X.shape
-        X_transpose = np.transpose(X)
-        self.w = np.linalg.pinv(X_transpose @ X) @ X_transpose @ t
-        self.var = np.mean(np.square(X @ self.w - t))
-
-    def predict(self, X, return_std=False):
-        N, D = X.shape
-        y = X @ self.w
-        
-        if return_std:
-            y_std = np.sqrt(self.var)
-            return y, y_std
-
-        return y
-
-
-class RidgeRegression(LinearRegression):
-    def __init__(self, alpha:float=1.):
-        super(RidgeRegression, self).__init__()
-        self.alpha = alpha
-        
-
-    def fit(self, X:np.ndarray, t:np.ndarray):
-        N, D = X.shape
-        X_transpose = np.transpose(X)
-        self.w = np.linalg.inv(X_transpose @ X + self.alpha * np.eye(D)) @ X_transpose @ t
-
-        self.var = np.mean(np.square(X @ self.w - t))
-
-
-if __name__ == "__main__":
-    # get the type of dataset
-    if len(sys.argv) < 2:
-        print("Usage: python convert_categorical_attributes.py <dataset>")
-        sys.exit(1)
-    dataset = sys.argv[1]
-    train_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_train_standardized_one_hot.csv'
-    test_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_test_standardized_one_hot.csv'
-    
-    # Load the datasets
-    train = pd.read_csv(train_file)
-    test = pd.read_csv(test_file)
-
-    # get the target column
-    categorical_attributes = []
-    with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
-        categorical_attributes = f.read().splitlines()
-    target = categorical_attributes[-1]
-
-    X_train = train.drop(target, axis=1)
-    y_train = train[target]
-
-    X_test = test.drop(target, axis=1)
-    y_test = test[target]
-
-    # Convert bool columns to float64
[…]

### 2024-05-25 homework2/linear_regression/scikit-learn_implementation.py

@@ -1,100 +1,0 @@
-import numpy as np
-import os
-import sys
-import pandas as pd
-from sklearn.linear_model import LinearRegression
-import matplotlib.pyplot as plt
-from sklearn.linear_model import Ridge as RidgeRegression
-import seaborn as sns
-from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
-
-
-
-if __name__ == "__main__":
-	# get the type of dataset
-	if len(sys.argv) < 2:
-		print("Usage: python convert_categorical_attributes.py <dataset>")
-		sys.exit(1)
-	dataset = sys.argv[1]
-	train_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_train_standardized_one_hot.csv'
-	test_file = f'./tema2_{dataset}/preprocessed_correlated_{dataset}_test_standardized_one_hot.csv'
-	
-	# Load the datasets
-	train = pd.read_csv(train_file)
-	test = pd.read_csv(test_file)
-
-	# get the target column
-	categorical_attributes = []
-	with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
-		categorical_attributes = f.read().splitlines()
-	target = categorical_attributes[-1]
-
-	X_train = train.drop(target, axis=1).values
-	y_train = train[target].values
-
-	X_test = test.drop(target, axis=1).values
-	y_test = test[target].values
-
-	# Create and train the linear regression model
-	model =  RidgeRegression(alpha=1.75)
-	model.fit(X_train, y_train)
-
-	# Make predictions
-	y_train_pred = model.predict(X_train)
-	y_test_pred = model.predict(X_test)
-
-	# convert the predictions to binary
-	# get the maximum value of the predictions
-	mean_train = np.mean(y_train_pred)
-	y_train_pred = np.where(y_train_pred > mean_train, 1, 0)
-	mean_test = np.mean(y_test_pred)
-	y_test_pred = np.where(y_test_pred > mean_test, 1, 0)
-
-	# Evaluate the model
-	# Training metrics
-	train_accuracy = accuracy_score(y_train, y_train_pred)
-	train_precision = precision_score(y_train, y_train_pred, average='weighted')
-	train_recall = recall_score(y_train, y_train_pred, average='weighted')
[…]

### 2024-05-25 homework2/logistic_regression/convert_categorical_attributes.py

@@ -1,0 +1,95 @@
+import numpy as np
+import os
+import sys
+import pandas as pd
+import matplotlib.pyplot as plt
+
+from sklearn.metrics import classification_report
+
+
+def convert_categorical_attributes(X_train, X_test, cat_attr, header):
+# Convert categorical attributes to one-hot encoded format
+    cat_indices = [list(header).index(attr) for attr in cat_attr]
+
+    categorical_data = X_train[:, cat_indices]
+    categorical_df = pd.DataFrame(categorical_data, columns=cat_attr)
+    XX_train = pd.get_dummies(categorical_df, prefix=cat_attr)
+
+    categorical_data = X_test[:, cat_indices]
+    categorical_df = pd.DataFrame(categorical_data, columns=cat_attr)
+    XX_test = pd.get_dummies(categorical_df, prefix=cat_attr)
+
+    XX_train = XX_train.reindex(columns=XX_test.columns, fill_value=0)
+    XX_test = XX_test.reindex(columns=XX_train.columns, fill_value=0)
+
+    # Drop original categorical columns
+    X_train = np.delete(X_train, cat_indices, axis=1)
+    X_test = np.delete(X_test, cat_indices, axis=1)
+    header = np.delete(header, cat_indices)
+
+    # Concatenate the original array with the new one-hot encoded columns
+    X_train = np.concatenate((X_train, XX_train.values), axis=1)
+    X_test = np.concatenate((X_test, XX_test.values), axis=1)
+    header = np.concatenate((header, XX_train.columns))
+
+
+    return X_train, X_test, header
+
+def solver(dataset):
+    train_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_train.csv'
+    test_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_test.csv'
+    
+    if not os.path.exists(train_file) or not os.path.exists(test_file):
+        print("Please standardize the dataset first")
+        sys.exit(1)
+
+    categorical_attributes = []
+    with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
+        categorical_attributes = f.read().splitlines()
+    
+    # check if all the categorical attributes are in the dataset
+    header = pd.read_csv(train_file).columns
+    for cat_attr in categorical_attributes:
+        if cat_attr not in header:
+            categorical_attributes.remove(cat_attr)
+    # remove the target attribute
+    categorical_attributes.remove(header[-1])
+    
[…]

### 2024-05-25 homework2/logistic_regression/manual_implementation.py

@@ -1,0 +1,144 @@
+import numpy as np
+import os
+import sys
+import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
+
+
+class LogisticRegression:
+    def __init__(self, learning_rate=0.01, num_iterations=10000, regularization_strength=0.1):
+        self.learning_rate = learning_rate
+        self.num_iterations = num_iterations
+        self.regularization_strength = regularization_strength
+
+    def sigmoid(self, z):
+        return 1 / (1 + np.exp(-z))
+
+    def compute_cost(self, X, y, weights):
+        m = len(y)
+        h = self.sigmoid(np.dot(X, weights))
+        regularization_term = (self.regularization_strength / (2 * m)) * np.sum(np.square(weights[1:]))
+        cost = -1/m * (np.dot(y, np.log(h)) + np.dot((1 - y), np.log(1 - h))) + regularization_term
+        return cost
+
+    def gradient_descent(self, X, y, weights):
+        m = len(y)
+        for i in range(self.num_iterations):
+            h = self.sigmoid(np.dot(X, weights))
+            gradient = np.dot(X.T, (h - y)) / m
+            regularization_term = (self.regularization_strength / m) * weights
+            regularization_term[0] = 0
+            weights -= self.learning_rate * (gradient + regularization_term)
+            if i % 1000 == 0:
+                cost = self.compute_cost(X, y, weights)
+        return weights
+
+    def fit(self, X, y):
+        X = np.insert(X, 0, 1, axis=1)
+        weights = np.zeros(X.shape[1])
+        self.weights = self.gradient_descent(X, y, weights)
+
+    def predict(self, X):
+        X = np.insert(X, 0, 1, axis=1)
+        probabilities = self.sigmoid(np.dot(X, self.weights))
+        return probabilities >= 0.5
+
+def solver(dataset):
+    train_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_train_converted.csv'
+    test_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_test_converted.csv'
+    
+    # Load the datasets
+    train = pd.read_csv(train_file)
+    test = pd.read_csv(test_file)
+
+    # get the target column
[…]

### 2024-05-25 homework2/logistic_regression/scikit-learn_implementation.py

@@ -1,0 +1,99 @@
+import numpy as np
+import os
+import sys
+import pandas as pd
+from sklearn.linear_model import LogisticRegression
+import matplotlib.pyplot as plt
+import seaborn as sns
+from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
+
+
+def solver(dataset):
+	train_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_train_converted.csv'
+	test_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_test_converted.csv'
+	
+	# Load the datasets
+	train = pd.read_csv(train_file)
+	test = pd.read_csv(test_file)
+
+	# get the target column
+	categorical_attributes = []
+	with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
+		categorical_attributes = f.read().splitlines()
+	target = categorical_attributes[-1]
+
+	X_train = train.drop(target, axis=1).values
+	y_train = train[target].values
+
+	X_test = test.drop(target, axis=1).values
+	y_test = test[target].values
+
+	# Create and train the logistic regression model
+	model = LogisticRegression()
+	model.fit(X_train, y_train)
+
+	# Make predictions
+	y_train_pred = model.predict(X_train)
+	y_test_pred = model.predict(X_test)
+
+	# convert the predictions to binary
+	# get the maximum value of the predictions
+	mean_train = np.mean(y_train_pred)
+	y_train_pred = np.where(y_train_pred > mean_train, 1, 0)
+	mean_test = np.mean(y_test_pred)
+	y_test_pred = np.where(y_test_pred > mean_test, 1, 0)
+
+	# Evaluate the model
+	# Training metrics
+	train_accuracy = accuracy_score(y_train, y_train_pred)
+	train_precision = precision_score(y_train, y_train_pred, average='weighted', zero_division=1)
+	train_recall = recall_score(y_train, y_train_pred, average='weighted', zero_division=1)
+	train_f1 = f1_score(y_train, y_train_pred, average='weighted', zero_division=1)
+
+	# Test metrics
+	test_accuracy = accuracy_score(y_test, y_test_pred)
+	test_precision = precision_score(y_test, y_test_pred, average='weighted', zero_division=1)
[…]

### 2024-05-25 homework2/preprocessing/impute_values.py

@@ -1,12 +1,11 @@
 import numpy as np
 import pandas as pd
 import os
-from sklearn.experimental import enable_iterative_imputer
-from sklearn.impute import IterativeImputer, SimpleImputer
+from sklearn.impute import SimpleImputer
 from scipy import stats
 
 
-def missing_values(file_name):
+def missing_values(file_name, imputers=None):
 	file = pd.read_csv(file_name)
 
 	# convert ? to NaN
@@ -20,18 +19,25 @@ def missing_values(file_name):
 	with open(os.path.join(os.path.dirname(file_name), 'categorical_attributes.txt'), 'r') as f:
 		categorical_cols = f.read().splitlines()
 
-	# Impute missing values in numeric columns
-	imputer = IterativeImputer(max_iter=10, random_state=0)
-	imputed_numeric = imputer.fit_transform(file[numerical_cols])
+	numeric_imputer = SimpleImputer(strategy='mean')
+	categorical_imputer = SimpleImputer(strategy='most_frequent')
+	if imputers is not None:
+		numeric_imputer = imputers['numeric']
+		categorical_imputer = imputers['categorical']
 
-	# Impute missing values in categorical columns: missing values can contain ? or NaN or nothing
-	# Use the most frequent value for each column
-	imputed_categorical = file[categorical_cols].apply(lambda x: x.fillna(x.value_counts().index[0]))
+	# Impute missing values in numeric columns
+	df_numeric_imputed = pd.DataFrame(numeric_imputer.fit_transform(file[numerical_cols]), columns=numerical_cols)
+
+	# Impute missing values in categorical columns
+	df_categorical_imputed = pd.DataFrame(categorical_imputer.fit_transform(file[categorical_cols]), columns=categorical_cols)
+
+	# Store imputers for future use
+	imputers = {'numeric': numeric_imputer, 'categorical': categorical_imputer}
 
 	# Concatenate the imputed numeric and categorical columns
-	imputed_file = pd.concat([pd.DataFrame(imputed_numeric, columns=numerical_cols), imputed_categorical], axis=1)
+	imputed_file = pd.concat([df_numeric_imputed, df_categorical_imputed], axis=1)
 
-	return imputed_file
+	return imputed_file, imputers
 	
[…]

### 2024-05-25 homework2/preprocessing/standardization.py

@@ -40,16 +40,19 @@ def standardize(train_file, test_file):
 	test_standardized = pd.DataFrame(X_test, columns=header[:-1])
 	test_standardized[header[-1]] = y_test
 
-	train_standardized.to_csv(train_file.replace('.csv', '_standardized.csv'), index=False)
-	test_standardized.to_csv(test_file.replace('.csv', '_standardized.csv'), index=False)
+	return train_standardized, test_standardized
 
 if __name__ == '__main__':
 	# Salary Prediction dataset
 	train_file = './tema2_SalaryPrediction/preprocessed_correlated_SalaryPrediction_train.csv'
 	test_file = './tema2_SalaryPrediction/preprocessed_correlated_SalaryPrediction_test.csv'
-	standardize(train_file, test_file)
+	train, test = standardize(train_file, test_file)
+	train.to_csv('./tema2_SalaryPrediction/preprocessed_standardized_SalaryPrediction_train.csv', index=False)
+	test.to_csv('./tema2_SalaryPrediction/preprocessed_standardized_SalaryPrediction_test.csv', index=False)
 
 	# Stroke Prediction dataset
 	train_file = './tema2_AVC/preprocessed_correlated_AVC_train.csv'
 	test_file = './tema2_AVC/preprocessed_correlated_AVC_test.csv'
-	standardize(train_file, test_file)
+	train, test = standardize(train_file, test_file)
+	train.to_csv('./tema2_AVC/preprocessed_standardized_AVC_train.csv', index=False)
+	test.to_csv('./tema2_AVC/preprocessed_standardized_AVC_test.csv', index=False)

### 2024-05-26 homework2/paper/paper.tex
MLP from scikit

@@
 \subsection{Multi-Layered Perceptron (MLP)}
-Multi-Layer Perceptron (MLP) is a type of artificial neural network designed to 
-capture complex, non-linear patterns in data. Unlike Linear Regression, which is
-limited to linear relationships, MLPs consist of multiple layers of interconnected 
-nodes, or neurons, that can model intricate interactions between variables.
+A Multilayer Perceptron (MLP) is a class of feedforward artificial neural network 
+that consists of at least three layers of nodes: an input layer, one or more 
+hidden layers, and an output layer. Each node, except for the input nodes, is a 
+neuron that uses a nonlinear activation function. MLPs are capable of modeling 
+complex relationships in data, making them suitable for tasks such as 
+classification, regression, and pattern recognition. The network learns by 
+adjusting the weights through a process called backpropagation, which minimizes 
+the error between the predicted outputs and the actual targets. This 
+adaptability and learning capability make MLPs powerful tools in machine 
+learning and artificial intelligence applications.
 
-MLPs are particularly suitable for tasks involving complex datasets with 
-non-linear relationships, such as salary prediction. In salary prediction, 
-factors like education, work experience, industry sector, and other 
-socio-economic variables interact in complex ways that Linear Regression might 
-not fully capture. MLPs can learn these non-linear patterns, providing a more 
-accurate and nuanced understanding of the factors influencing salary levels.
+In the \textit{mlp} folder at the root of the project directory, we implemented the
+Multi-Layered Perceptron algorithm in two different ways:
+
+\begin{itemize}
+    \item \textit{MLP with Scikit-Learn:} We used the Scikit-Learn library to implement
+    the MLP algorithm on the preprocessed datasets.
+    \item \textit{MLP from Scratch:} We implemented the MLP algorithm from scratch using
+    the Negative Log-Likelihood method, the Gradient Descent optimization algorithm, and
+    the Sigmoid activation function.
+\end{itemize}
+
+Before starting the implementation of the MLP algorithm, we need to standardize the
+numeric attributes in the datasets as in the Logistic Regression algorithm.
+
+For the MLP with Scikit-Learn implementation, we used the \textit{MLPClassifier} class
+from the \textit{sklearn.neural\_network} module to train the model on the preprocessed
+datasets, without setting any hyperparameters, using the default values.
+
+For the MLP from Scratch implementation, we implemented the Negative Log-Likelihood loss
+function, the Gradient Descent optimization algorithm, and the Sigmoid activation function.
+We trained the model on the preprocessed datasets, setting the learning rate to 0.01, the
+number of epochs to 10000, and the number of hidden units to 100. For the regularization,
+we used the Ridge Regression technique.
+
+The results of both implementations for each dataset are saved in the \textit{MLP}
+folder at the specific dataset's root.
 

### 2024-05-26 homework2/mlp/manual_implementation.py

@@ -1,0 +1,213 @@
+from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
+from typing import List
+import numpy as np
+import os
+import sys
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+class Layer:
+    def forward(self, x: np.ndarray) -> np.ndarray:
+        raise NotImplementedError
+        
+    def backward(self, x: np.ndarray, dy: np.ndarray) -> np.ndarray:
+        raise NotImplementedError
+        
+    def update(self, *args, **kwargs):
+        pass
+
+
+class FeedForwardNetwork:
+    def __init__(self, layers: List[Layer]):
+        self.layers = layers
+        
+    def forward(self, x: np.ndarray, train: bool = True) -> np.ndarray:
+        self._inputs = []
+        for layer in self.layers:
+            if train:
+                self._inputs.append(x)
+            x = layer.forward(x)
+        return x
+    
+    def backward(self, dy: np.ndarray) -> np.ndarray:
+        for x, layer in reversed(zip(self._inputs, self.layers)):
+            dy = layer.backward(x, dy)
+        return dy
+    
+        del self._inputs
+    
+    def update(self, *args, **kwargs):
+        for layer in self.layers:
+            layer.update(*args, **kwargs)
+
+class Linear(Layer):
+    def __init__(self, insize: int, outsize: int) -> None:
+        bound = np.sqrt(6. / insize)
+        self.weight = np.random.uniform(-bound, bound, (insize, outsize))
+        self.bias = np.zeros((outsize,))
+        
+        self.dweight = np.zeros_like(self.weight)
+        self.dbias = np.zeros_like(self.bias)
+   
+    def forward(self, x: np.ndarray) -> np.ndarray:
+        self.x = x
+        return np.dot(x, self.weight) + self.bias
+    
+    def backward(self, x: np.ndarray, dy: np.ndarray) -> np.ndarray:
+        self.dweight = np.dot(self.x.T, dy)
+        self.dbias = np.sum(dy, axis=0)
+        return np.dot(dy, self.weight.T)
+    
+    def update(self, mode='SGD', lr=0.001, mu=0.9):
+        if mode == 'SGD':
[…]

### 2024-05-26 homework2/mlp/scikit-learn_implementation.py

@@ -1,0 +1,102 @@
+from sklearn.neural_network import MLPClassifier
+from sklearn.datasets import make_classification
+from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
+import numpy as np
+import os
+import sys
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+def solver(dataset):
+	train_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_train_converted.csv'
+	test_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_test_converted.csv'
+	
+	# Load the datasets
+	train = pd.read_csv(train_file)
+	test = pd.read_csv(test_file)
+
+	# get the target column
+	categorical_attributes = []
+	with open(f'./tema2_{dataset}/categorical_attributes.txt') as f:
+		categorical_attributes = f.read().splitlines()
+	target = categorical_attributes[-1]
+
+	X_train = train.drop(target, axis=1).values
+	y_train = train[target].values
+
+	X_test = test.drop(target, axis=1).values
+	y_test = test[target].values
+
+	# Create and train the MLP model
+	print(f"Training MLP {dataset} model...")
+	model = MLPClassifier(hidden_layer_sizes=(100,), max_iter=1000, activation='logistic')
+	model.fit(X_train, y_train)
+
+	# Make predictions
+	print(f"Making predictions for {dataset} dataset...")
+	y_train_pred = model.predict(X_train)
+	y_test_pred = model.predict(X_test)
+
+	# convert the predictions to binary
+	# get the maximum value of the predictions
+	mean_train = np.mean(y_train_pred)
+	y_train_pred = np.where(y_train_pred > mean_train, 1, 0)
+	mean_test = np.mean(y_test_pred)
+	y_test_pred = np.where(y_test_pred > mean_test, 1, 0)
+
+	# Evaluate the model
+	print(f"Evaluating the model for {dataset} dataset...")
+	# Training metrics
+	train_accuracy = accuracy_score(y_train, y_train_pred)
+	train_precision = precision_score(y_train, y_train_pred, average='weighted', zero_division=1)
+	train_recall = recall_score(y_train, y_train_pred, average='weighted', zero_division=1)
[…]

### 2024-05-26 homework2/paper/paper.tex
finish

@@
 
+\pagebreak
+
 \begin{center}
@@
 
-\begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
-        \hline
-        \multicolumn{3}{|c|}{List of all Discrete Nominal Attributes in the Salary Prediction dataset} \\
-        \hline
-        & Non-missing count & Unique values count \\
-        \hline\hline
-        relation & 9999 & 6 \\
-        \hline
-        country & 9999 & 41 \\
-        \hline
-        job & 9999 & 14 \\
-        \hline
-        work\_type & 9999 & 9 \\
-        \hline
-        partner & 9999 & 7 \\
-        \hline
-        edu & 9999 & 16 \\
-        \hline
-        gender & 9199 & 2 \\
-        \hline
-        race & 9999 & 5 \\
-        \hline
-        gtype & 9999 & 2 \\
-        \hline
-        money & 9999 & 2 \\
-        \hline
-        \caption{Discrete Nominal Attributes in Salary Prediction Dataset}
-        \label{tab:discrete_nominal_attributes_salary} \\
-    \end{longtable}
-\end{center}
-
-\begin{center}
-    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
-        \hline
-        \multicolumn{3}{|c|}{List of all Discrete Nominal Attributes in the Stroke Prediction dataset} \\
-        \hline
-        & Non-missing count & Unique values count \\
-        \hline\hline
-        cardiovascular\_issues & 5110 & 2 \\
-        \hline
-        job\_category & 5110 & 5 \\
-        \hline
-        sex & 5110 & 2 \\
-        \hline
-        tobacco\_usage & 5110 & 4 \\
-        \hline
-        high\_blood\_pressure & 5110 & 2 \\
-        \hline
-        married & 4599 &2 \\
-        \hline
-        living\_area & 5110 & 2 \\
-        \hline
-        chaotic\_sleep & 5110 & 2 \\
-        \hline
-        cerebrovascular\_accident & 5110 & 2 \\
-        \hline
-
-        \caption{Discrete Nominal Attributes in Stroke Prediction Dataset}
-        \label{tab:discrete_nominal_attributes_stroke} \\
-    \end{longtable}
-\end{center}
-
 From the Discret Nominal Attributes tables (Tables \ref{tab:discrete_nominal_attributes_salary} and \ref{tab:discrete_nominal_attributes_stroke})
@@
 
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+        \hline
+        \multicolumn{3}{|c|}{List of all Discrete Nominal Attributes in the Salary Prediction dataset} \\
+        \hline
+        & Non-missing count & Unique values count \\
+        \hline\hline
+        relation & 9999 & 6 \\
+        \hline
+        country & 9999 & 41 \\
+        \hline
+        job & 9999 & 14 \\
+        \hline
+        work\_type & 9999 & 9 \\
+        \hline
+        partner & 9999 & 7 \\
+        \hline
+        edu & 9999 & 16 \\
+        \hline
+        gender & 9199 & 2 \\
+        \hline
+        race & 9999 & 5 \\
+        \hline
+        gtype & 9999 & 2 \\
+        \hline
+        money & 9999 & 2 \\
+        \hline
+        \caption{Discrete Nominal Attributes in Salary Prediction Dataset}
+        \label{tab:discrete_nominal_attributes_salary} \\
+    \end{longtable}
+\end{center}
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{5cm}||>{\centering\arraybackslash}m{3cm}|>{\centering\arraybackslash}m{3cm}|}
+        \hline
+        \multicolumn{3}{|c|}{List of all Discrete Nominal Attributes in the Stroke Prediction dataset} \\
+        \hline
+        & Non-missing count & Unique values count \\
+        \hline\hline
+        cardiovascular\_issues & 5110 & 2 \\
+        \hline
+        job\_category & 5110 & 5 \\
+        \hline
+        sex & 5110 & 2 \\
+        \hline
+        tobacco\_usage & 5110 & 4 \\
+        \hline
+        high\_blood\_pressure & 5110 & 2 \\
+        \hline
+        married & 4599 &2 \\
+        \hline
+        living\_area & 5110 & 2 \\
+        \hline
+        chaotic\_sleep & 5110 & 2 \\
+        \hline
+        cerebrovascular\_accident & 5110 & 2 \\
+        \hline
+
+        \caption{Discrete Nominal Attributes in Stroke Prediction Dataset}
+        \label{tab:discrete_nominal_attributes_stroke} \\
+    \end{longtable}
+\end{center}
+
+\pagebreak
+
 In the historigrams for the discrete nominal attributes, we can see the
@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/histogram_work_type_SalaryPrediction_full.png}
-    \caption{Histogram for the \textit{work\_type} attribute in the Salary Prediction dataset}
-    \label{fig:histogram_example_salary}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/histogram_tobacco_usage_AVC_full.png}
-    \caption{Histogram for the \textit{tobacco\_usage} attribute in the Stroke Prediction dataset}
-    \label{fig:histogram_example_stroke}
-\end{figure}
-
 In Figure \ref{fig:histogram_example_salary} we can see a histogram of the
@@
 data need to be addressed appropriately.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/histogram_work_type_SalaryPrediction_full.png}
+    \caption{Histogram for the \textit{work\_type} attribute in the Salary Prediction dataset}
+    \label{fig:histogram_example_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/histogram_tobacco_usage_AVC_full.png}
+    \caption{Histogram for the \textit{tobacco\_usage} attribute in the Stroke Prediction dataset}
+    \label{fig:histogram_example_stroke}
+\end{figure}
 
@@
 \end{itemize}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/distribution_SalaryPrediction.png}
+    \caption{Distribution of the \textit{money} class in the Salary Prediction dataset}
+    \label{fig:distribution_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/distribution_AVC.png}
+    \caption{Distribution of the \textit{cerebrovascular\_accident} class in the Stroke Prediction dataset}
+    \label{fig:distribution_stroke}
+\end{figure}
 
@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/distribution_SalaryPrediction.png}
-    \caption{Distribution of the \textit{money} class in the Salary Prediction dataset}
-    \label{fig:distribution_salary}
-\end{figure}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/distribution_AVC.png}
-    \caption{Distribution of the \textit{cerebrovascular\_accident} class in the Stroke Prediction dataset}
-    \label{fig:distribution_stroke}
-\end{figure}
-
 \subsection{Analysis of Feature Correlations}
@@
 
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_SalaryPrediction_full.png}
-    \caption{Correlation matrix for the Salary Prediction dataset}
-    \label{fig:correlation_matrix_example_salary}
-\end{figure}
-\pagebreak
 The correlation matrix and Cramér's V matrix provide valuable insights into the
@@
 attribute is negatively correlated with the \textit{analysis\_results} attribute.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../plots/correlation_matrix_SalaryPrediction_full.png}
+    \caption{Correlation matrix for the Salary Prediction dataset}
+    \label{fig:correlation_matrix_example_salary}
+\end{figure}
 
@@
 \section{Evaluation}
+The evaluation of machine learning models is a critical step in assessing their
+performance and effectiveness. By comparing the model's predictions to the actual
+ground truth, we can determine the model's accuracy, precision, recall, and other
+metrics that quantify its performance. This section delves into the evaluation
+of the Logistic Regression and Multi-Layered Perceptron (MLP) models on the
+Salary Prediction and Stroke Prediction datasets.
+
+\subsection{Hyperparameter Tuning}
+Hyperparameters are parameters that are set before the learning process begins.
+They control the learning process and the behavior of the model. Hyperparameter
+tuning is the process of selecting the optimal hyperparameters for a machine
+learning model to achieve the best performance. This process involves searching
+through different hyperparameter configurations and evaluating the model's
+performance on a validation set to find the optimal settings.
+
+In the context of the Logistic Regression (manual implementation), the 
+hyperparameters that were used are:
+
+\begin{itemize}
+    \item \textit{learning\_rate:} The rate at which the model updates the weights - 0.01
+    \item \textit{num\_iterations} The number of iterations the model trains for - 10000
+    \item \textit{regularization} The regularization parameter to prevent overfitting - 0.1
+\end{itemize}
+
+In the context of the Logistic Regression (Scikit-Learn implementation), the majority of
+hyperparameters that were used are the default values provided by the Scikit-Learn
+library. The only hyperparameters that were set are:
+
+\begin{itemize}
+    \item \textit{solver:} The optimization algorithm used in the model - 
+    'saga' for SalaryPrediction and 'sag' for AVC
+    \item \textit{max\_iter:} The maximum number of iterations for the optimization algorithm - 200 for SalaryPrediction and 500 for AVC
+    \item \textit{C:} The regularization parameter to prevent overfitting - 0.4342470001 for SalaryPrediction and  2.56050926 for AVC
+\end{itemize}
+
+In the context of the Multi-Layered Perceptron (manual implementation), the
+hyperparameters that were used are:
+
+\begin{itemize}
+    \item \textit{hidden\_sizes:} The sizes of the hidden layers are defined as a list - [256, 128, 64]
+    \item \textit{num\_epochs:} The number of epochs the model trains for - 100
+    \item \textit{learning\_rate:} The rate at which the model updates the weights - 0.01
+    \item \textit{Loss Function	:} The loss function used to optimize the model - Negative Log-Likelihood
+\end{itemize}
+
+In the context of the Multi-Layered Perceptron (Scikit-Learn implementation), the 
+hyperparameters that were used are the default values provided by the Scikit-Learn
+library.
+
+\subsection{Confusion Matrix}
+A confusion matrix is a table that summarizes the performance of a classification
+model on a set of test data for which the true values are known. It provides
+insights into the model's predictions, including true positive, true negative,
+false positive, and false negative instances. These metrics are essential for
+evaluating the model's performance and identifying potential areas for improvement.
+
+In the context of the Logistic Regression and Multi-Layered Perceptron models, we
+generated confusion matrices to analyze the model's predictions on the test data.
+The confusion matrices provide a detailed breakdown of the model's performance,
+highlighting the number of correct and incorrect predictions for each class.
+
+In Figures \ref{fig:confusion_matrix_logistic_manual_salary} and 
+\ref{fig:confusion_matrix_logistic_scikit_salary} we can see the confusion matrices
+for the Logistic Regression model on the Salary Prediction dataset, implemented
+manually and with Scikit-Learn, respectively. In Figures 
+\ref{fig:confusion_matrix_mlp_manual_salary} and \ref{fig:confusion_matrix_mlp_scikit_salary}
+we can see the confusion matrices for the Multi-Layered Perceptron model on the
+Salary Prediction dataset, implemented manually and with Scikit-Learn, respectively.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_SalaryPrediction/LogisticRegression/confusion_matrix_manual.png}
+    \caption{Confusion Matrix for the Logistic Regression model on the Salary Prediction dataset (Manual Implementation)}
+    \label{fig:confusion_matrix_logistic_manual_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_SalaryPrediction/LogisticRegression/confusion_matrix_scikit-learn.png}
+    \caption{Confusion Matrix for the Logistic Regression model on the Salary Prediction dataset (Scikit-Learn Implementation)}
+    \label{fig:confusion_matrix_logistic_scikit_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_SalaryPrediction/MLP/confusion_matrix_manual.png}
+    \caption{Confusion Matrix for the Multi-Layered Perceptron model on the Salary Prediction dataset (Manual Implementation)}
+    \label{fig:confusion_matrix_mlp_manual_salary}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_SalaryPrediction/MLP/confusion_matrix_scikit-learn.png}
+    \caption{Confusion Matrix for the Multi-Layered Perceptron model on the Salary Prediction dataset (Scikit-Learn Implementation)}
+    \label{fig:confusion_matrix_mlp_scikit_salary}
+\end{figure}
+
+We can see that the Logistic Regression model implemented with Scikit-Learn achieved
+almost the same results as the manual implementation and the MLP model implemented
+with Scikit-Learn. All 4 models achieved a high number of true positive predictions
+for the negative class, indicating that the models are effective at identifying
+negative instances in the dataset. The true positive predictions for the positive
+class are lower, but the models still achieved a reasonable prediction but the
+MLP model with manual implementation achieved the worst results. It may be due to
+the hyperparameters that were set for the model. The false positive and false negative
+predictions are also lower than the true positive and true negative predictions, which
+is a good sign for the model's performance.
+
+In Figures \ref{fig:confusion_matrix_logistic_manual_stroke} and
+\ref{fig:confusion_matrix_logistic_scikit_stroke} we can see the confusion matrices
+for the Logistic Regression model on the Stroke Prediction dataset, implemented
+manually and with Scikit-Learn, respectively. In Figures
+\ref{fig:confusion_matrix_mlp_manual_stroke} and \ref{fig:confusion_matrix_mlp_scikit_stroke}
+we can see the confusion matrices for the Multi-Layered Perceptron model on the
+Stroke Prediction dataset, implemented manually and with Scikit-Learn, respectively.
+
+Because of the high class imbalance in the Stroke Prediction dataset, the models
+can not predict the positive class effectively (the positive class is less than 5\%
+of the dataset). On the other hand, the models can predict the negative class
+effectively.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_AVC/LogisticRegression/confusion_matrix_manual.png}
+    \caption{Confusion Matrix for the Logistic Regression model on the Stroke Prediction dataset (Manual Implementation)}
+    \label{fig:confusion_matrix_logistic_manual_stroke}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_AVC/LogisticRegression/confusion_matrix_scikit-learn.png}
+    \caption{Confusion Matrix for the Logistic Regression model on the Stroke Prediction dataset (Scikit-Learn Implementation)}
+    \label{fig:confusion_matrix_logistic_scikit_stroke}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_AVC/MLP/confusion_matrix_manual.png}
+    \caption{Confusion Matrix for the Multi-Layered Perceptron model on the Stroke Prediction dataset (Manual Implementation)}
+    \label{fig:confusion_matrix_mlp_manual_stroke}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{../tema2_AVC/MLP/confusion_matrix_scikit-learn.png}
+    \caption{Confusion Matrix for the Multi-Layered Perceptron model on the Stroke Prediction dataset (Scikit-Learn Implementation)}
+    \label{fig:confusion_matrix_mlp_scikit_stroke}
+\end{figure}
+
+\subsection{Evaluation Metrics}
+To evaluate the performance of the machine learning models, we employ a range of
+evaluation metrics that provide insights into different aspects of the model's
+performance. These metrics include:
+
+\begin{itemize}
+    \item \textit{Accuracy:} The proportion of correctly classified instances
+    out of the total instances. It provides a general measure of the model's
+    correctness.
+    \item \textit{Precision:} The proportion of true positive predictions out
+    of all positive predictions. It measures the model's ability to avoid false
+    positives.
+    \item \textit{Recall:} The proportion of true positive predictions out of
+    all actual positive instances. It measures the model's ability to capture
+    all positive instances.
+    \item \textit{F1-Score:} The harmonic mean of precision and recall. It provides
+    a balanced measure of the model's performance.
+\end{itemize}
+
+These evaluation metrics help us understand the strengths and weaknesses of the
+machine learning models and guide us in improving their performance. By analyzing
+these metrics, we can identify areas for optimization and fine-tuning to enhance
+the model's predictive capabilities.
+
+In the following tables, Table \ref{tab:evaluation_metrics_salary} and Table
+\ref{tab:evaluation_metrics_stroke}, we present the evaluation metrics for the 
+Logistic Regression
+and Multi-Layered Perceptron models on both prediction tasks: Salary Prediction and
+Stroke Prediction.
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{4cm}||>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{2cm}|}
+        \hline
+        \multicolumn{5}{|c|}{Evaluation Metrics for Salary Prediction Dataset - Train set} \\
+        \hline
+        Model & Accuracy & Precision & Recall & F1-Score \\
+        \hline\hline
+        Logistic Regression (Manual) & 0.829353669 & 0.700867052 & 0.504945340 & 0.586989409 \\
+        \hline
+        Logistic Regression (Scikit-Learn) & 0.8371046 & 0.707940780 & 0.54763144 & 0.617552098 \\
+        \hline
+        Multi-Layered Perceptron (Manual) & 0.759844980 & \textbf{1.0} & 0.0 & 0.0 \\
+        \hline
+        Multi-Layered Perceptron (Scikit-Learn) & \textbf{0.91098887} & 0.85537918 & \textbf{0.75741801} & \textbf{0.80342352} \\
+        \hline
+        \hline
+        \multicolumn{5}{|c|}{Evaluation Metrics for Salary Prediction Dataset - Test set} \\
+        \hline
+        Model & Accuracy & Precision & Recall & F1-Score \\
+        \hline\hline
+        Logistic Regression (Manual) & \textbf{0.831} & 0.712250712 & 0.513347022 & 0.596658711 \\
+        \hline
+        Logistic Regression (Scikit-Learn) & 0.825 & 0.676092544 & \textbf{0.54004106} & \textbf{0.60045662} \\
+        \hline
+        Multi-Layered Perceptron (Manual) & 0.7565 & \textbf{1.0} & 0.0 & 0.0 \\
+        \hline
+        Multi-Layered Perceptron (Scikit-Learn) & 0.811 & 0.635235732 & 0.525667351 & 0.575280898 \\
+        \hline
+        \caption{Evaluation Metrics for Logistic Regression and Multi-Layered Perceptron Models on Salary Prediction Dataset}
+        \label{tab:evaluation_metrics_salary} \\
+    \end{longtable}
+\end{center}
+
+\begin{center}
+    \begin{longtable}{ |>{\centering\arraybackslash}m{4cm}||>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{2cm}|>{\centering\arraybackslash}m{2cm}|}
+        \hline
+        \multicolumn{5}{|c|}{Evaluation Metrics for Stroke Prediction Dataset - Train set} \\
+        \hline
+        Model & Accuracy & Precision & Recall & F1-Score \\
+        \hline\hline
+        Logistic Regression (Manual) & 0.957191780 & 0. & 0.0 & 0.0 \\
+        \hline
+        Logistic Regression (Scikit-Learn) & 0.957436399 & \textbf{1.0} & 0.0 & 0.0 \\
+        \hline
+        Multi-Layered Perceptron (Manual) & 0.957436399 & \textbf{1.0} & 0.0 & 0.0 \\
+        \hline
+        Multi-Layered Perceptron (Scikit-Learn) & \textbf{0.96232876} & 0.83333333 & \textbf{0.14367816} &  \textbf{0.24509803} \\
+        \hline
+        \hline
+        \multicolumn{5}{|c|}{Evaluation Metrics for Stroke Prediction Dataset - Test set} \\
+        \hline
+        Model & Accuracy & Precision & Recall & F1-Score \\
+        \hline\hline
+        Logistic Regression (Manual) & 0.925636007 & 0.0 & 0.0 & 0.0 \\
+        \hline
+        Logistic Regression (Scikit-Learn) & 0.925636007 & 0.0 & 0.0 & 0.0 \\
+        \hline
+        Multi-Layered Perceptron (Manual) & \textbf{0.92661448} & \textbf{1.0} & 0.0 & 0.0 \\
+        \hline
+        Multi-Layered Perceptron (Scikit-Learn) & 0.92367906 & 0.2 & \textbf{0.01333333} & \textbf{0.025} \\
+        \hline
+
+        \caption{Evaluation Metrics for Logistic Regression and Multi-Layered Perceptron Models on Stroke Prediction Dataset}
+        \label{tab:evaluation_metrics_stroke} \\
+    \end{longtable}
+\end{center}
 
@@
 \section{Conclusions}
+In this document, we explored the application of machine learning algorithms to
+two distinct prediction tasks: Salary Prediction and Stroke Prediction. We conducted
+a comprehensive analysis of the datasets, including data exploration, feature
+correlation analysis, and data preprocessing. We implemented two prominent machine
+learning algorithms, Logistic Regression and Multi-Layered Perceptron (MLP), to
+predict the outcomes of the two tasks. We evaluated the models using various
+evaluation metrics, including accuracy, precision, recall, and F1-Score, to assess
+their performance.
+
+The results of the evaluation metrics indicate that the models achieved varying
+levels of performance on the two prediction tasks. The Logistic Regression model
+performed well on the Salary Prediction dataset, achieving high accuracy and F1-Score
+values. The Multi-Layered Perceptron model also demonstrated strong performance on
+the Salary Prediction dataset, achieving high accuracy and F1-Score values. However,
+the models struggled to predict the positive class effectively
+on the Stroke Prediction dataset due to the high class imbalance. The Logistic
+Regression and Multi-Layered Perceptron models achieved similar results on the
+Stroke Prediction dataset, with low recall and F1-Score values for the positive class.
+
+Overall, it is evident that the Multi-Layered Perceptron (Scikit-Learn 
+implementation) generally performs better than Logistic Regression across both 
+datasets. Specifically:
+
+\begin{itemize}
+    \item \textit{Salary Prediction:} The MLP model achieves higher accuracy, 
+    precision, recall, and F1-score on the training set compared to Logistic
+     Regression. While its test set performance is slightly lower than its 
+     training set performance, it still demonstrates a competitive F1-score.
+    \item \textit{Stroke Prediction:} Despite the challenge posed by class 
+    imbalance, the MLP model achieves a better balance between precision and 
+    recall than Logistic Regression, which fails to predict the positive class 
+    at all, resulting in an F1-score of 0.
+\end{itemize}
+
+Therefore, based on these metrics, the Multi-Layered Perceptron (Scikit-Learn 
+implementation) is the better algorithm for both the Salary Prediction and 
+Stroke Prediction tasks due to its superior performance across multiple 
+evaluation criteria.
 
@@
     \bibitem{}
-    \href{https://en.wikipedia.org/wiki/Linear_regression}{Wikipedia - Linear Regression}
-    \bibitem{}
-    \href{https://en.wikipedia.org/wiki/Linear_classifier}{Wikipedia - Linear Classifier}
+    \href{https://en.wikipedia.org/wiki/Machine_learning}{Wikipedia - Machine Learning}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Logistic_regression}{Wikipedia - Logistic Regression}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Multilayer_perceptron}{Wikipedia - Multilayer Perceptron}
     \bibitem{}
@@
     \bibitem{}
+    \href{https://developers.google.com/machine-learning/crash-course/classification/accuracy}{Google Developers - Classification: Accuracy}
+    \bibitem{}
+    \href{https://ocw.mit.edu/courses/6-867-machine-learning-fall-2006/}{MIT OpenCourseWare - Machine Learning}
+    \bibitem{}
     \href{https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.html}{Pandas - DataFrame}
+    \bibitem{}
+    \href{https://en.wikipedia.org/wiki/Imputation_(statistics)}{Wikipedia - Imputation (statistics)}
     \bibitem{}

### 2024-05-26 homework2/logistic_regression/manual_implementation.py

@@ -83,24 +83,18 @@ def solver(dataset):
     y_train_pred = model.predict(X_train)
     y_test_pred = model.predict(X_test)
 
-    # Convert the predictions to binary
-    mean_train = np.mean(y_train_pred)
-    y_train_pred = np.where(y_train_pred > mean_train, 1, 0)
-    mean_test = np.mean(y_test_pred)
-    y_test_pred = np.where(y_test_pred > mean_test, 1, 0)
-
     # Evaluate the model
     # Training metrics
     train_accuracy = accuracy_score(y_train, y_train_pred)
-    train_precision = precision_score(y_train, y_train_pred, average='weighted', zero_division=1)
-    train_recall = recall_score(y_train, y_train_pred, average='weighted', zero_division=1)
-    train_f1 = f1_score(y_train, y_train_pred, average='weighted', zero_division=1)
+    train_precision = precision_score(y_train, y_train_pred, zero_division=1)
+    train_recall = recall_score(y_train, y_train_pred, zero_division=1)
+    train_f1 = f1_score(y_train, y_train_pred, zero_division=1)
 
     # Test metrics
     test_accuracy = accuracy_score(y_test, y_test_pred)
-    test_precision = precision_score(y_test, y_test_pred, average='weighted', zero_division=1)
-    test_recall = recall_score(y_test, y_test_pred, average='weighted', zero_division=1)
-    test_f1 = f1_score(y_test, y_test_pred, average='weighted', zero_division=1)
+    test_precision = precision_score(y_test, y_test_pred, zero_division=1)
+    test_recall = recall_score(y_test, y_test_pred, zero_division=1)
+    test_f1 = f1_score(y_test, y_test_pred, zero_division=1)
 
     # Confusion matrix
     train_cm = confusion_matrix(y_train, y_train_pred)

### 2024-05-26 homework2/logistic_regression/scikit-learn_implementation.py

@@ -3,10 +3,40 @@ import os
 import sys
 import pandas as pd
 from sklearn.linear_model import LogisticRegression
+from sklearn.model_selection import RandomizedSearchCV
 import matplotlib.pyplot as plt
 import seaborn as sns
+from scipy.stats import uniform
 from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_score, confusion_matrix
 
+def get_best_model(X_train, y_train, X_test, y_test):
+	# Create the model
+	model = LogisticRegression()
+
+	# Create the hyperparameter space: regularization parameter, solver, max_iter
+	hyperparameters = {
+		'C': uniform(loc=0, scale=4),
+		'solver': ['newton-cg', 'lbfgs', 'liblinear', 'sag', 'saga'],
+		'max_iter': [200, 500, 1000]
+	}
+
+	# Create the RandomizedSearchCV object
+	search = RandomizedSearchCV(model, hyperparameters)
+
+	# Fit the model
+	search.fit(X_train, y_train)
+
+	# Get the best model
+	best_model = search.best_estimator_
+
+	# print the best hyperparameters
+	print(f"Best hyperparameters: {search.best_params_}")
+
+	# Make predictions
+	y_train_pred = best_model.predict(X_train)
+	y_test_pred = best_model.predict(X_test)
+
+	return y_train_pred, y_test_pred
 
 def solver(dataset):
 	train_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_train_converted.csv'
@@ -28,33 +58,22 @@ def solver(dataset):
 	X_test = test.drop(target, axis=1).values
 	y_test = test[target].values
 
-	# Create and train the logistic regression model
-	model = LogisticRegression()
-	model.fit(X_train, y_train)
-
-	# Make predictions
-	y_train_pred = model.predict(X_train)
-	y_test_pred = model.predict(X_test)
-
-	# convert the predictions to binary
-	# get the maximum value of the predictions
-	mean_train = np.mean(y_train_pred)
-	y_train_pred = np.where(y_train_pred > mean_train, 1, 0)
-	mean_test = np.mean(y_test_pred)
-	y_test_pred = np.where(y_test_pred > mean_test, 1, 0)
+	# Get the best parameters for the model
+	print(f"Training Logistic Regression {dataset} model...")
+	y_train_pred, y_test_pred = get_best_model(X_train, y_train, X_test, y_test)
 
 	# Evaluate the model
[…]

### 2024-05-26 homework2/mlp/helper.py

@@ -1,0 +1,24 @@
+
+def plot_learning_curves(train_acc, test_acc, train_loss, test_loss, title):
+    """Plot learning curves for training and test accuracy and loss."""
+    epochs = range(1, len(train_acc) + 1)  # Generate a range of epochs
+    plt.figure(figsize=(14, 6))  # Create a figure for the plots
+
+    plt.subplot(1, 2, 1)  # Create a subplot for accuracy
+    plt.plot(epochs, train_acc, 'bo-', label='Train Accuracy')  # Plot training accuracy
+    plt.plot(epochs, test_acc, 'ro-', label='Test Accuracy')  # Plot test accuracy
+    plt.title(f'{title} - Accuracy')  # Set the title for the accuracy plot
+    plt.xlabel('Epochs')  # Label the x-axis
+    plt.ylabel('Accuracy')  # Label the y-axis
+    plt.legend()  # Add a legend
+
+    plt.subplot(1, 2, 2)  # Create a subplot for loss
+    plt.plot(epochs, train_loss, 'bo-', label='Train Loss')  # Plot training loss
+    plt.plot(epochs, test_loss, 'ro-', label='Test Loss')  # Plot test loss
+    plt.title(f'{title} - Loss')  # Set the title for the loss plot
+    plt.xlabel('Epochs')  # Label the x-axis
+    plt.ylabel('Loss')  # Label the y-axis
+    plt.legend()  # Add a legend
+
+    plt.tight_layout()  # Adjust subplots to fit into the figure area
+    plt.show()  # Display the plots

### 2024-05-26 homework2/mlp/manual_implementation.py

@@ -2,122 +2,85 @@ from sklearn.metrics import accuracy_score, precision_score, recall_score, f1_sc
 from typing import List
 import numpy as np
 import os
-import sys
 import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
-
-class Layer:
-    def forward(self, x: np.ndarray) -> np.ndarray:
-        raise NotImplementedError
-        
-    def backward(self, x: np.ndarray, dy: np.ndarray) -> np.ndarray:
-        raise NotImplementedError
-        
-    def update(self, *args, **kwargs):
-        pass
+import torch
+import torch.nn as nn
+import torch.optim as optim
+from sklearn.preprocessing import StandardScaler
 
 
-class FeedForwardNetwork:
-    def __init__(self, layers: List[Layer]):
-        self.layers = layers
-        
-    def forward(self, x: np.ndarray, train: bool = True) -> np.ndarray:
-        self._inputs = []
-        for layer in self.layers:
-            if train:
-                self._inputs.append(x)
-            x = layer.forward(x)
-        return x
-    
-    def backward(self, dy: np.ndarray) -> np.ndarray:
-        for x, layer in reversed(zip(self._inputs, self.layers)):
-            dy = layer.backward(x, dy)
-        return dy
-    
-        del self._inputs
-    
-    def update(self, *args, **kwargs):
-        for layer in self.layers:
-            layer.update(*args, **kwargs)
+class MLP(nn.Module):
+	def __init__(self, input_size, hidden_sizes, output_size):
+		super(MLP, self).__init__()
+		layers = []
+		in_size = input_size
+		for hidden_size in hidden_sizes:
+			layers.append(nn.Linear(in_size, hidden_size))
+			layers.append(nn.Sigmoid())
+			in_size = hidden_size
+		layers.append(nn.Linear(in_size, output_size))
+		layers.append(nn.LogSoftmax(dim=1))
+		self.network = nn.Sequential(*layers)
 
-class Linear(Layer):
-    def __init__(self, insize: int, outsize: int) -> None:
-        bound = np.sqrt(6. / insize)
-        self.weight = np.random.uniform(-bound, bound, (insize, outsize))
[…]

### 2024-05-26 homework2/mlp/scikit-learn_implementation.py

@@ -8,6 +8,37 @@ import pandas as pd
 import seaborn as sns
 import matplotlib.pyplot as plt
 
+def nll(Y, T):
+    """Compute the negative log-likelihood (NLL) for binary classification."""
+    N = T.shape[0]  # Number of samples
+    return -np.sum(T * np.log(Y) + (1 - T) * np.log(1 - Y)) / N  # Calculate NLL
+
+def plot_learning_curves(train_acc, test_acc, train_loss, test_loss, title, dataset):
+    """Plot learning curves for training and test accuracy and loss."""
+    epochs = range(1, len(train_acc) + 1)  # Generate a range of epochs
+    plt.figure(figsize=(14, 6))  # Create a figure for the plots
+
+    plt.subplot(1, 2, 1)  # Create a subplot for accuracy
+    plt.plot(epochs, train_acc, 'bo-', label='Train Accuracy')  # Plot training accuracy
+    plt.plot(epochs, test_acc, 'ro-', label='Test Accuracy')  # Plot test accuracy
+    plt.title(f'{title} - Accuracy')  # Set the title for the accuracy plot
+    plt.xlabel('Epochs')  # Label the x-axis
+    plt.ylabel('Accuracy')  # Label the y-axis
+    plt.legend()  # Add a legend
+
+    plt.subplot(1, 2, 2)  # Create a subplot for loss
+    plt.plot(epochs, train_loss, 'bo-', label='Train Loss')  # Plot training loss
+    plt.plot(epochs, test_loss, 'ro-', label='Test Loss')  # Plot test loss
+    plt.title(f'{title} - Loss')  # Set the title for the loss plot
+    plt.xlabel('Epochs')  # Label the x-axis
+    plt.ylabel('Loss')  # Label the y-axis
+    plt.legend()  # Add a legend
+
+    plt.tight_layout()  # Adjust subplots to fit into the figure area
+
+    # Save the learning curves
+    plt.savefig(f'./tema2_{dataset}/MLP/learning_curves_scikit-learn.png')
+
 def solver(dataset):
 	train_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_train_converted.csv'
 	test_file = f'./tema2_{dataset}/preprocessed_standardized_{dataset}_test_converted.csv'
@@ -30,7 +61,7 @@ def solver(dataset):
 
 	# Create and train the MLP model
 	print(f"Training MLP {dataset} model...")
[…]
